FACTIONS of the UTCS Time Period


FIXME

§	Confed IntelSec 
o	Confederation Navy 
o	Exploratory Service 
o	Concerned Confederation Citizens Against the War 
o	Confed Pleasure Planet Travel Consortium (CPPTC) 

\section{Aeran Ascendancy}

The Aeran Ascendency governs the Aerans and Bzbr. The Aeran Ascendancy
is, in practice, a body wholly controlled by entirely Aeran will of
the Aeran Oligarchy. All distinct Aeran subgroups, such as the Aeran
Merchant Marines are subject to the edicts of the Aeran Oligarchy.

\subsection{Aera}
Faction data 
Aera 
Species 	Aera 
Homeworld (Origin) 	Aeneth 
Capital 	Aeneth 


\subsubsection{A Brief History of the Aera}

The Aera homeworld was home to extremely competitive ecosystems. It
was not a supportive environment. If one considers Earth the "mother"
of the human race, then, in comparison, Aeneth was an abusive
parent. The Aera evolved in an environment where a slew of things
really did want to kill/eat/infest them. Beginning with the early
harnessing of fire, and ending with industrial might, the Aera
remedied this problem by destroying the vast jungles that bore
them. However, their outlook on the universe was fundamentally shaped
by their beginnings in a direction that humans would consider
paranoid, or at least profoundly pessimistic and somewhat untrusting,
though the Area merely see it as prudent recognition of how the
universe works.

This outlook was greatly reinforced by the Aera experience with the
nano-plague. The precocious Aera, unlike humans or Rlaan, developed
jump-based FTL before having otherwise left their solar system. The
resultant reactivation of the nano-plague and the devastation it
wrought on their population, served to gel their concept of an
inherently antagonistic universe. Aeran society greatly increased its
militarization, heretofore on the wane since global unification, so as
to better prepare for potential conflicts in what, evidenced by the
nano-plague, they presumed was an inhabited galaxy.

The Aera are the youngest of the three currently dominant space-faring
major groups, but are also the most expansionist, fastest breeding,
shortest lived, and devote the highest fraction of their economy to
military and military related R\&D spending. They are not evil, they
are not delusional, they are not irrational, but they have some
fundamentally different assumptions they are working from that make
them somewhat difficult to get along with. Finding out that their
section of the jump network had them pinned by the Humans and Rlaan,
they first attempted to negotiate passage, and were rebuffed. They
then attempted to sneak a colony convoy through Rlaan space, but this
turned into an utter debacle after Aeran escorts killed Rlaan
Civilians, sparking a war lasting several years, that churned the
Rlaan-Aeran border into an abattoir. Although formal peace has never
been brokered between the two, a cease-fire has been in effect for
several years. The Aera have now turned their sights toward Human
space, invading Forsaken territory, hoping to push through toward the
less defended Forsaken/Confederation border crossings, carve a
corridor through to the other side of humans space, and keep it open
long enough so that they can send enough colonization fleets through
to the other side to make the venture worthwhile.

\subsubsection{Development}
Compared to contemporary 33rd century humans, the Aera are comparable
or somewhat more advanced in some of the physical sciences and their
applications, notably so with respect weaponizations of certain
technologies. They are noticeably behind in life sciences and AI.

Having been wandering the jump network for the least amount of time
(among the Rlaan/Humans/Aera), the Aera, though occupying the same
order of magnitude of systems as the Humans or the Rlaan
(Rlaan/Lmpl/Nuhln/Saahasayaay have most, followed by
Humans/Klk'k/Dgn/Purth/Mishtali, followed by Aera/Bzbr, then the much
smaller Uln, Shmrn) have not occupied many of them for nearly as
long. The Aera expanded in territory faster than that territory could
be developed up until running into the Uln. After one last push then
brought them to the Human and Rlaan borders as well, the Aera have
been racing to build up their newly settled colonies nearer the
borders, but the bulk of their industrial potential remains
concentrated in systems closer to their homeworld than to alien
space. This difference was especially clear during the Rlaan-Aera
conflict, wherein many newly settled Aeran colonies along the border
fell to the Rlaan assault, but the same Rlaan fleets were badly
bloodied when they tried to push into Aeran systems with more matured
defenses. Due to the war, Aeran military spending and infrastructure
development has been extravagant in comparison to human budgets, but
the Aera also had to cope with sizeable losses in personnel and
materiel.

While there are key differences between the level of population and
industrial development between the core colonies and the newer
colonies, due to the very strong central organizing forces in Aeran
governance and economy, this is not a deep political divide, nor an
economic one - it is merely a matter of the more fringe planets
growing as fast as they can into states undifferentiable from the more
core worlds. This centralization should not be taken as evidence that
Aerans are a selfless society of collectivists. Rather, the Aera have
a strong natural ability and desire to sublimate personal interests to
higher authority, a trait left over from their more pack-like
origins. Success, however, is still judged at individual granularity,
and Aera are entirely opportunistic about personal advancement when
the opportunity either does not come at the expense of the dictates of
higher authority, or places them into a position of higher authority.

\subsubsection{Culture}
The Aera are a bit culturally dour, although they do engage in
organizational events, such as rallies, sporting contests, and
Military parades. Entertainment pursuits, such as music, for personal
pleasure are not a significant thread in Aeran culture. Such pursuits
are seen as necessary avenues of release, but to devote oneself to
pursuing purely entertainment oriented activities merely because they
are pleasant is seen as wasteful, wantonly hedonistic, and a reckless
abandonment of one's duties. Entertainment with physical components,
such as sporting competitions, are viewed in a more favorable light.

In Aeran culture, mortality is to be pondered and meditated upon in
its inevitability. It is worth noting that it is not so much those
that came before them that the Aera cherish as the accomplishments of
those who came before them. The Aera perspective is that it is only
though accomplishment on behalf of the Aera that the Aera can
continue, and the dead, thereby, can continue to live in the memories
of the Aera.

\subsubsection{Organization}
Aera culture is highly organized and decidedly hierarchical, but in
the form of a meritocracy rather than an aristocracy. While what has
constituted merit has morphed over the millennia since the first Aera
tribes selected work crews to cut back the encroachments of the jungle
upon their early settlements, given the relative position of the
pre-technological Aera in their local food chain, there has long been
a favoring of cleverness and determination over raw strength. The
current social and vocational position of any Aera is immediately
indicated by the color and pattern of an individual's coverall. The
Aera are ruled by a subset of the highest caste, with membership in
the oligarchy changing whenever either an individual steps down, or a
third of the other members call for a member's replacement. New
members must be confirmed by two thirds of the current oligarchy. It
is much more common for members to voluntarily remove themselves from
power, believing themselves more useful elsewhere in society, than to
be cast out. An average stay in the oligarchy lasts a few Aera years.

Need to more usefully carve up faction vs. species information. Will
do so later.FIXME

\subsection{Aeran Merchant Marine}

Faction data 
Merchant Marines 
Species 	Aera 
Homeworld (Origin) 	Aeneth 
Capital 	Aeneth 

The Aera don't really have a civilian sector. All Aera resources are
controlled by the state, and ruled over by the Oligarchy. The Aera do,
however differentiate between combatant and non-combatant forces, and
the Merchant Marine, while armed, are designed to transport goods
through dangerous space rather than conquer said space. (It may be
duly noted that the Aera are predisposed to consider all space to be
dangerous).

\subsection{Bzbr}

Faction data 
Bzbr 
Species 	Bzbr 
Homeworld (Origin) 	Fixme? 
Capital 	Fixme? 
	
The Bzbr are a client race uplifted by the Aera. They are reasonable
laborers, but poor conversationalists, in part due to their lack of
vocal cords.  The Aera uplifted the Bzbr out of pity for the Bzbr due
to similarities the Bzbr homeworld had with the Aera homeworld.

The Bzbr engage in significant hero worship of the Aera, although
indications are that the Aera are not appreciative of this
tendency. The Bzbr are wholly integrated into the Aeran Ascendency,
and given extremely little self-control as a group. Aeran officials
permeate their entire governmental structure, itself in parts borrowed
from and imposed by the Aerans upon the existing stone-age Bzbr. The
Bzbr, however, remain near universally accepting of their Aeran
overlords due to a combination of religious reverence and vast
improvements in their standards of living. Experiences during the
Rlaan/Aera conflict found Bzbr to be even less accepting than Aerans
of non-Aeran governance. Their fanatical level of religious devotion
to the Aera found them to be prime subjects for guerilla actions
involving the likely demise of the participants.


\section{Confederation of Inhabited Worlds}

Faction data 
Confed 
Species 	Humans, Klk'k, Purth, Dgn and Mishtali 
Homeworld (Origin) 	Various 
Capital 	Mars 

The Confederation of Inhabited Worlds (Confed) is made up of
partially-autonomous member states, and was created with a primary
purpose being to arbitrate disputes between them in a more structured
and less violent way than the Andolian-Lightbearer dispute was
handled. In addition to this, with the Klk'k representing a reasonably
advanced alien species and the Andolian-Lightbearer war pushing home
notions of the potential for interstellar war, a re-evaluation of the
need for maintaining a unified defensive force was seen to be in
order.

Each constituent state is allowed to maintain its own military forces,
but it must submit some portion of its resources to maintaining the
Confederation fleet, which is distinct from any state's military.

The confederation's rights over its constituent states are those and
only those so ceded by the constituent states in the confederation
constitution. Admittedly, exactly what rights were ceded is always a
debatable issue. Likewise, the confederation's unicameral legislature
has oft attempted to pass dubiously constitutional edicts. However,
the constitutionality of the Homeland Security Forces is actually
unimpeachable, as they were specifically called for in the
constitution to bypass the previously tangled web of extradition
policies between the constituent states and to force some measure of
legal compatibility between said states such that outright and overt
support for groups violently antagonistic to other member states would
not be tolerated. While none of the member states is particularly
ecstatic about the confederation's nature and its imposition on their
internal affairs, all are, at least publicly, thankful for the unified
front that allows humanity to operate alongside alien entities much
larger than any single member state.

The civilian center of the confederation as well as the fleet
headquarters is located on Mars since the Sol location is symbolic of
humanity's common roots. Earth came with too much political baggage,
and so Mars was chosen to be the center of the confederation. A
hierarchy of civilian and military installations then extends out from
Mars to the furthest reaches of confederation space. While sufficient
function is distributed redundantly to prevent a catastrophic
beheading, only the confed installations on Mars can be said to have
been designed with splendor in mind, and thus hold no rivals among the
other installations.




\subsection{Andolian Protectorate}
\index{Andolian!Andolian Protectorate}

Faction data 
Andolian Protectorate 
Species 	Human, Klk'k and Purth 
Homeworld (Origin) 	Various 
Capital 	Kubernan 

The Andolian Protectorate refers specifically to a government. 

\subsubsection{Members}

All Andolians are members of the Andolian Protectorate, not all
members of the Andolian Protectorate are Andolians. In particular,
there are Klk'k, Purth, and non-Andolian human citizens of the
Protectorate (the non-Andolian humans mostly being non-integrationist
members of annexed entities or yet to be integrated immigrants to the
Protectorate from other polities). Andolians, being the dominant
faction within the Protectorate, are good indicators of how the
Protectorate will act.

\subsubsection{History}

The Andolian Protectorate was created by the Andolians as the
political entity controlling both Andolian and Klk'k interests. During
the formation of the Confederation, there was some consolidation of
minor factions, bringing an influx of non-Andolian humans into the
Protectorate (an extremely small minority of the human population born
into the Protectorate remains non-Andolian - less than 1\% of the total
human population). Many Klk'k are more accurately considered Andolian
at present (in ideological adherence), but the integrated Klk'k are a
distinct minority of the total Klk'k population, which is
ideologically more diverse.


\subsubsection{Andolians}

Faction data 
Andolians 
Species 	Human, primarily of the Pluralis variant. 
Homeworld (Origin) 	Earth 
Capital 	Kubernan 

Andolians are a human faction primarily consisting of the Pluralis
variant well developed in both information and physical resources. All
Andolians are members of the Andolian Protectorate but maintain a
subordinate government concerned with Andolian affairs.

Many of the major Human governments could be considered "ideocracies"
wherein a supermajority, often a near totality, of the population
governed by an entity are adherents of a given ideological
perspective. The Andolians were such a group, with a clear
correspondance between the Andolians (the adherents) and the Andolians
(the political entity controlled by the adherents) until the Andolians
came to have a more diverse political landscape when they became
Protectors of the Klk'k population, and later, the Purth. Many Klk'k
have since become integrated with the Andolian mainstream, but the
Protectorate remains more ideologically diverse than the Andolians
proper.

Needs more FIXME 

\subsubsection{Klk'k}

Faction data 
Klk'k 
Species 	Klk'k 
Homeworld (Origin) 	Ktah 
Capital 	Ktah 
Needs a better desc FIXME 

The Klk'k are a client race of the Andolians, but enjoy exceptional
freedom within this position. Indeed, while fiercely loyal to their
Andolian benefactors, there exist a non-trivial number of
independently operated Klk'k vessels. The Klk'k are perhaps best known
for their odd brand of twisted humor.

\subsubsection{Purth}
\index{Andolian@Purth}
\index{Purth@\textit{Purth}}

Faction data 
Purth 
Species 	Purth 
Homeworld (Origin) 	fixme 
Capital 	fixme 

The Purth are politically subserviant members of the Andolian
Protectorate, heavily reliant in most facets of their existence upon
the assistance of the Klk'k and Human members of the Protectorate. The
Purth do not operate independently, instead being heavily integrated
into the Protectorate military forces, as might be expected given
their origins as a cybernetics research project.  FIXME

\subsubsection{Spaceborn}

Faction data 
Spaceborn 
Species 	Spaceborn 
Homeworld (Origin) 	Cradle
Capital 	Kubernan 

Those Spaceborn who chose to remain as such rather than attempt to
reintegrate into more mainline human gene-pools live almost
exclusively under the auspices of their liberators, the Andolian
Protectorate. The Spaceborn make up a miniscule fraction of the total
Protectorate population, but are heavily involved in many of the
Protectorate's free-space construction and maintenance projects.


\subsection{Highborn}

Faction data 
Highborn 
Species 	Humans 
Homeworld (Origin) 	Earth 
Capital 	Fixme? 

The Highborn are a human faction better known for its disregard for
life than its art or achievement.

They are one of the oldest colonial groups and had significant
resources on Earth, lending them the cream of the initial colony
locations. They were the main defenders of the 1st Confed Party
Reform, which had forbidden the ISO.  Primary manufacturers of dueling
weaponry.  Motto: "There is no substitute for a superior human being."

\subsection{Homeland Security}

Faction data 
Homeland Security 
Species 	Human 
Homeworld (Origin) 	Earth 
Capital 	Mars 

The Homeland Security forces are primarily composed of Purists but
contain members from most human groups. The Homeland Security forces
exist for internal policing and control, especially in cases crossing
between the jurisdictional boundaries of member states of the
Confederation, as they are an arm of the Confederation rather an
amalgamation of forces from member states. While distinctly limited by
both practical and political concerns in how much pressure they can
bring to bear if a member state actively obstructs the pursuit of
their duties, they counterbalance by being more authoritarian then
generally deemed necessary within what is clearly their own
domain. This is known to be especially true of the IntelSec wing of
the Homeland Security Forces.

\subsection{Hunters}

Faction data 
Hunter 
Species 	Humans 
Homeworld (Origin) 	Earth 
Capital 	Non-governmental group. 

Arising in the profit opportunities of the lawlessness of the Diamond
Dust period, various mercenary groups underwent iterations of
consolidation due to expanding polities, increasing regulation, and a
decline in per-capita bounty-hunting opportunities. While still
comprised of numerous independent groups of guns-for-hire, the
Hunter's guild provides a unified interface for those seeking their
still-needed services and a well-financed legal wing to make sure that
bounty hunting remains as legal as it is profitable.

\subsection{League of Independent Human Worlds}

Faction data 
LIHW 
Species 	Human 
Homeworld (Origin) 	Earth 
Capital 	FIXME

Scattered across the frontiers of known space are many worlds housing
minor subsets of humanity which hold no place among the larger
meme-group entities.

Humanity has, throughout the past, been an oft balkanized lot. While
the bulk of human power and population has, for various reasons,
aligned itself with one of the major or minor meme-groups there are
many colonies, that, for reasons of either intense pluralism or
adherence to a tertiary meme-group have remained independent (examples
of such being the inhabitants of Vegan-ville, or the citizens of the
"Brotherhood of Militant Agnostics" (motto: "I don't know, and neither
do you!") ).

But in the wake of the Mankind's first notable interstellar fraternal
conflict--the demolishing of the Lightbearers by the Andolians--and
the efforts which followed in the founding of the Confed, whose
membership consists of the major meme-groups, the bulk of the lesser
subsets realized that--while they may not get along all that well with
each other--if they did not in some way present a united front of
resistance, they would likely be consumed by the major meme-groups. As
such, they joined the Confed united as the League of Independent Human
Worlds (LIHW), and, counter-intuitively, gained, through their
co-operation, a guarantee of protection of their individual and often
separatist modes of life.

Those that did not band together or exit en mass to Forsaken space,
are now only records in history books, having been overrun, subverted,
co-opted, or in other ways gobbled up by the major meme-groups.

\subsection{Mechanists}

Faction data 
Mechanist 
Species 	Mechanist 
Homeworld (Origin) 	Earth 
Capital 	Plato
FIXME Stub entry - taken from dynamic\_news\_content.py 
·	Full faction name: Mandate for Corporeal Perfection via the Abandonment of Flesh 
·	Government name: Concordance of Enlightened Ones 

\subsection{Merchant's Guild}

Faction data 
Interstellar Shipping and Mercantile Guild 
Species 	Humans 
Homeworld (Origin) 	Earth 
Capital 	Bantam 

A latecomer to the great sowing of Earth's seeds, a too-small planet,
previously passed over, became the new home for a group of colonists
united more by a desire to leave behind their varied existences in Sol
than by any common ethos. The colonists struck out from their
settlements in the deep canyons and valleys where the air was thicker,
clearing out the simple layers of native life, ever so slowly etching
humanity's presence indelibly into the planet's crust. However, the
future of their efforts did not lie on the surface of a world, but in
the cold of space surrounding it.

No too-small world can hope to passively hold on to the sort of
atmosphere that gives rise to sunny, green meadows alight with
frolicking schoolchildren. The release from the domes, from the deep
places, would be a Herculean task, but the insights bestowed by modern
science had given mere mortal men more abilities than the authors of
the ancient demigod had foreseen. In a common goal, the residents of
Bantam found new unity and identity. They labored together tirelessly
to construct the centerpiece of their terraforming effort: a colossal
station that would serve as both the shipyard producing and
maintaining the resource gathering fleet and the processing center
that would amalgamate the offerings of the entire system into
sustenance for Bantam's development.

It would likely have worked. 

The nano-plague struck without warning, long before the first human
FTL ship would visit the system. Only the near-paranoid levels of
over-engineering with which Rainbow station had been constructed saved
it from the fate of many of Bantam's residents, too reliant on cheap
and plentiful nanites. It was far too early in the terraforming effort
to live in the open, and suddenly overstrained environmental systems
caused entire cities to suffocate. Bantam was devastated, losing all
vibrancy and vitality in a matter of weeks. What before had been a
growing planet-wide civilization was now a collection of frightened
outposts of inhabitation. Thus were Rainbow station and its infant
fleet of resource gatherers nearly orphaned, the mother planet an
invalid. With the station barely able to sustain itself, the
stationers were left with the unenviable role of deciding whom amongst
Bantam's survivors they could afford to save.

As the Diamond Dust Age progressed, and Bantam continued to wither
away to a pale shadow of its brief heyday, Rainbow station endured,
growing throughout the period of isolation, if at a glacial
pace. Warning messages, lazily arriving at the speed of light from
their neighbors, came in one after the other, followed sometimes by
descriptions of chaos, and sometimes by silence. There was no official
expectation of assistance, but humans are renowned for their
counter-empirical optimism.

Thus, it was with great surprise and anticipation that the arrival of
the first FTL ship was greeted, and numbing disappointment the feeling
that swept the stationers when they learned that it was not a portend
of any mission of mercy. All, however, did not look grim. The visitors
were High-Born explorers, but from a family of lower standing. The
Marcos family realized that they had much to gain from control of
Rainbow station once it began producing jump capable vessels. With FTL
technology and the silent threat of a second abandonment for leverage,
the Marcos family gained a large stake in Rainbow station.

Luck had indeed changed for the better for the people of Bantam and
Rainbow. Not only were much needed resources flowing in from High-Born
space, payments for an ever increasing list of ship construction
orders, but Rainbow soon proved itself to occupy a prime location in
the jump network, a hub on what would remain for many decades the only
practical routes connecting several populous systems to each other -
including Sol.

Some colonies had fallen far and some had not had far to fall. The
lone merchant was a valuable target in a starved land. Neither were
the financial benefits of collusion upon profit margins lost upon the
increasing number of traders operating out of or, less commonly at the
time, passing through Rainbow. The first of the great mercantile
cartels was formed, again under the leadership of the Marcos clan, a
silent, nameless cartel that worked to establish a near monopoly on
interstellar trade in much of human space during the Reconstruction
period. Coercion, cooption, and even the occasional assassination were
all valid tools to grow the blossoming financial empire centered on
Rainbow and the slowly awakening Bantam. The Rainbow station cartel
was one of a few groups to capitalize upon the realization that, in
the absence of any overarching authorities, whomsoever dominated the
trade lanes now could well be able to do so for centuries to come,
barring fantastic advances in interstellar travel.

Generations passed. Competitors fell as they came, though some
struggled valiantly, and only the governments of the larger
meme-groups were sufficiently imposing to hold their own at the
bargaining tables. Shares of control of Rainbow station changed hands,
and it became wholly an instrument of the cartel, although Bantam and
Stationer descendants still held a plurality of control of the cartel,
a sizeable majority if one included the increasingly integrated Marcos
clan. The name of the station changed then as well, and the cartel
gained a name, rechristened Cherryh station and the Cherryh Mercantile
Trust (CMT), respectively.

Though flagrantly in bed with High-Born interests, the CMT maintained
a resolute official neutrality, even going so far as to make a point
of hiring other parties for anti-piracy protection rather than
building their own combat fleets. While many at the time took this to
be a purely political maneuver, those more informed knew that it was
actually a necessity. Cherryh station had been a vital trump card for
the CMT in the early portions of the Reconstruction period, but
mankind had recovered somewhat, and was once more advancing, and the
age of the station was beginning to show. Bulk freighters are in many
ways much easier to build than the tiny craft that protect them and
avoid obsolescence much longer, and Cherryh station was only up to the
task of constructing the first, and not the latter.

More modern shipyards, though not on the scale of the CMT's aging
beauty, were being constructed all over, and the competitive advantage
the CMT had enjoyed due to their production capacity was waning. There
was, however, little feared by the CMT at this point, as there was no
longer any competition of note to gain upon them. All other transport
of goods was in-system, governmental, or doomed to obscurity or
absorption. The CMT used this time to increase its wealth, strengthen
and modernize its infrastructure, and deepen its influence on various
governments. The resettlement of Bantam was in full swing.

The CMT's opportunity to rest did not last as long as they would have
preferred. The development of the SPEC drive dramatically changed the
landscape of trade routes, altered latencies, and made runs by smaller
vessels much more practical. Rapidly expanding borders made for new
regions for local competitors to exploit more rapidly than the CMT
might be able to react. However, like the age itself, the CMT had
matured. There would be no massive wave of assassinations or stream of
engineered third-party blockades. Rather, there would be the ultimate
buy-out - the largest recruitment effort ever mounted, a clarion call
summoning all traders to a common banner with the offer of common
profits. What could have been the death knell for the CMT merely
announced the birth of the Interstellar Shipping and Mercantile Guild,
which if it still had the CMT at its core, was, by the nature of its
creation, intended to be a beast with a very different public face.

The emergence of the Confederation only served to strengthen the
position of the Guild, being included in the Confederation Senate's
Committees despite technically only owning a single star system (the
plethora of commerce stations and trading outposts in the star systems
of other powers not counting as seats of population) - an important
technicality, as it would otherwise be difficult to pretend they were
a governmental entity in the same sense as the other Senate
members. The newly formed Confederation Navy and Homeland Security
forces served to decrease piracy along standard trade routes, even if
the decline of several independent powers in the face of enforcing new
Confederation regulations gave rise to a whole new breed of
paramilitary forces. While in matters internal to the Confederation
the Guild was still maintaining an increasingly less fictitious
neutrality on matters not of self-interest, there was no love lost
between the Forsaken and the Guild, the much lower profit margins
available on return trips never having enticed much Guild investment,
and Guild price fixing long having raised the ire of the
Forsaken. Guild presence in Forsaken space is thus limited to border
stations, much like the trade relations with the Uln and Rlaan.

Contact with alien trading partners was a great boon to the Guild, as
Guild political clout ensured that all major alien traffic would pass
through Guild trading stations, that Guild shipping would be awarded
nearly all lucrative inter-government contracts calling for human
merchants, and so forth. Moreover, no independent trader could so
easily call upon the shared resources of the Guild in obtaining
appropriate permissions from the Uln and Rlaan governments to conduct
business within their borders, or the expertise of a cultural
specialist in determining the likely intentions of an alien customer.

Though the upper echelons of the Guild are still plentifully stocked
with CMT personnel, the Guild favored success over nepotism, and many
rose through the ranks to positions of power - unlike the larger
meme-groups, no grand conversion of belief system is required for
acceptance, merely the conviction that profit conquers all, and timely
payment of portions thereof. The lower tiers of the Guild are full of
all sort and manner of small time traders who were willing to trade a
fraction of their profits for protection, access, and
opportunity. Unlike the pre-SPEC period, the smaller independent
traders are not actively squashed, and those who achieve success are
always initially courted rather than destroyed outright, but this is
largely because any emerging independent group which appeared to offer
a real threat would be rapidly legislated out of existence by calling
in favors in the Confederation Senate. The current Guild is rivaled
only by the logistical corps of some major meme-groups that did not
desire to give up the independence of their internal supply chains,
but even these entities are dwarfish by comparison. Though still not a
military power, nor aspiring to become one, the Guild's numerous
shipyards, kept modernized under lucrative Confederation contracts now
churn out capital and sub-capital military vessels for the
Confederation Navy alongside their freight bearing kin. With the
renovation of Cherryh station over the last few decades further
boosting production, the Guild shipyards maintain a distinct lead, by
tonnage, as the largest suppliers of Confederation military vessels.

\subsection{Purist}

Faction data 
Purist 
Species 	Human 
Homeworld (Origin) 	Earth 
Capital 	Earth 

A surprisingly large memegroup, and the dominant one in Sol, the
Purists seek a humanity free of changes to the species. They do not
stand for non-therapudic genetic modification, nor for unnecessary
cybernetic implants. They are, however, not opposed to technology
which is not used to alter humans. Motto: In being what we are, become
all that we can be.






\subsection{Shapers}

Faction data 
Shaper 
Species 	Human 
Homeworld (Origin) 	Earth 
Capital 	Bifröst 

The Shaper government, such as one exists, was originally constructed
along 'libertarian' lines (albeit with some elements that would seem
deeply foreign or even perhaps contradictory to any person harking to
the same 'libertarian' title who is not also a devotee of the Shaper
ideology), and is a generally minimalist entity, delegating many of
the tasks performed by the other human governments to private
interests within the Shaper populace. Confoundingly, the Shaper views
on children are remarkably collectivist, and closely tied to the
particular Shaper notions of what constitutes an actual human
existence. Government social programs are unheard of for adults, but
heavily invested for children, and while there are a number of
government standards committees, they almost exclusively produce
guidelines, rather than regulations. Government interference into the
economy is utterly untolerated. Individual civil liberties are
fiercely protected, and one of the more significantly funded programs,
aside from defense spending, is the publicly funded "Authority
Surveillance Guardianship" or ASGuard, the funding for which and chief
administrators of being determined directly by the Shaper populace.

Shaper society is egalitarian in opportunity, but extremely
competitive and indifferent, when not outright unforgiving, of
failure, with notable emphasis being put into ensuring that the
failings of members of one generation are not allowed to be magnified
through their offspring. Notably, procreation in any form (be it
cloning, or gene-mixing, or more traditional conception) is not
considered a right within Shaper society. The furtherance of one's
genetic material is deemed a privilege of personal success. Attempting
to raise Shaper children without appropriate resources or accepted
competence in child-rearing is considered criminally abusive of the
children. Non-Shaper children have a decidedly different legal status,
are protected only in a sense of general 'human' liberties, and are
under no circumstances the recipients of collective support.

It should be noted that childhood is a rather abbreviated thing in
Shaper society, due to genetic engineering and common practices
greatly affecting both pre-natal and post-natal growth and
learning. Natural childbirth is seen as exceptionally eccentric within
Shaper society and is tolerated in much the same way that artists in
the 20th century who deigned to use their own bodily fluids in their
artwork were tolerated, but far from welcomed. Normal Shaper children
are born in child producing factories, more often than not, with the
exception of vanity children, being very carefully designed from mixes
of parental DNA with sequences from currently popular designs. The
gestation period in such factories tends to runs toward 14 months, a
length second only to Andolian full-growth gestation-immersion tanks -
unlike the other cultures using the tanks to replace normal childbirth
where the delay of from conception to birth is often seen as an
inconvenience to the prospective parents and time-to-delivery is
generally minimized, Shaper society considers the physically
ineffectual infancy of normal humanity far more inconvenient (and
repugnantly archaic - an evolutionary solution from before our own
evolution was ours to control) than the delay. Shaper children leave
the tanks already able to walk, run, and metabolize most foods - which
is important, as they have a voracious metabolism that will lead them
to grow to their adult mass within another 7 years. Brain development
in Shaper children is also accelerated. The mind is extremely plastic
for the first two years post-natal, and allows for socialization and
other acclimating patterns to form via processes still analogous to
those of a human child, if a somewhat precocious and extremely
intelligent one. The brain then rapidly proceeds to develop fully all
of its cognitive regions, at the expense of much of the childlike
plasticity for reorganization, and the Shaper child begins their
"adolescence", a state lasting from age two through seven, during
which they will leave the direct care of their parents, and begin to
compete against one another to be able to progress through an
assortment of privately run educational facilities (albeit the
students themselves are still collectively supported at this
point). The general educational period lasts an intense five years,
ending at the same time as Shaper children achieve physical and legal
adulthood and are no longer supported in any collective fashion. This
academic feat is made possible by the extremely low sleep needs of the
Shaper body, the lack of any need to exercise or endure sun exposure
in order to maintain health and muscle tone, and the
genetically-engineered capacity for extreme focus, easy memorization,
and ravenous intellect that every Shaper is possessed of.

For those not at the top of the competition, they remain free to
choose their path, but even the dullest Shaper is more than bright
enough, and by the turn of adulthood, sufficiently scrubbed of initial
ignorance, to understand with fair clarity where they are likeliest to
find others willing to choose them, and the importance of their own
below-the-billing role in Shaper society. The Shaper population is not
very densely arrayed on their planets, and the most menial tasks are
handled by Dgn, engineered lifeforms, and machines. While the Shapers
utilize AIs, the bulk of these AIs, for the past several hundred
years, have generally been of foreign construction, and the Shapers
are somewhat wary of the allegiances of the AIs in their employ - both
to other AI entities and to their constructors. A number of Shaper
specialists are therefore employed to monitor the AIs, and some Shaper
research has gone into developing minds better suited to such
tasks. Social unrest remains extremely low among the Shaper population
- Shaper society is not utopian, but it is without noticeable friction
- despite the intense competition and uneven distribution of
resources. This is largely due to the firmness of belief in the Shaper
ideology ingrained into the population - the underlying rightness of
the system of struggle is accepted, the rightness of the goals of the
society unquestioned, and those involved are possessed of minds
capable of understanding how they have come to occupy their place in
society, their role in that society and its continued existence, and,
just as importantly, how other individuals in Shaper society fit into
the greater picture. They operate with self-interest, but informed,
exquisitely reasoned, and long-termed self-interest that includes a
notion of Shaper society that has been made difficult to cast as
counter to personal interests.

For those who succeed, there is more than time for art - to practice
The Art is success. In the Shaper ethos, design is king, and flesh is
the favored palette. Thus, Shaper spacecraft, if not aesthetically
neglected as with the Mechanists, and, to an arguably lesser extent,
the Unadorned, are not anywhere near as profound as their work
groundside. While there is some artistry at work in their inanimate
constructions, working with metal and stone is as cold as either to
the Shaper mind: pale when compared to the calling to bend the path of
life to one's will, expressed for as long as the creations may
continue to survive their competitors and environs. Those tools, such
as spacecraft, of which great functionality is demanded, suffer
greatly in relative aesthetics to architectures that may be bent to
support the aesthetic demands of myriad Shaper minds, writ in living
spaces filled with living things, but only those living things chosen
to be there. While this leads to jaw-dropping wonderlands of living
material populating their art-oriented enclosures and an enjoyable
spread of magnificent creations adorning most living areas, the
exercise in control over what life lives where leaves Shaper planets
as devoid of real wilderness as an industrial farming world, if far
prettier for the same detriment. There is a sense that even a vast
uninhabited stretch on a Shaper world, possessed of some paradise-like
qualities in its safety and provision as it may be, is just a
carefully tended garden laying fallow for future inhabitants to
paint. There is chaos allowed to taint the waiting land. Nature is not
allowed to sully any canvas destined for Shaper hands.

The outside observer rarely tends to have their eyes drawn to the
Shaper underachievers, both because of the higher standards of living
(relative to Purist worlds, for instance), even on the low end, that
the Shaper economy has (made possible, in large part, by limiting the
population density on Shaper worlds), and the astounding beauty that
the high end of the Shaper culture produces. There is a Shaper saying
that, "It is no accident that the path to perfection travels through
beauty." The functional and aesthetic are merged whenever tenable in
Shaper society, and there is a significant place reserved for the
aesthetic, even absent, though rarely in lieu of, the functional. To
create more perfectly is the pervasive goal, and beauty is often
aligned with the perfect for anything that must be seen. It should be
noted that Shaper artists have produced works considered masterpieces
in a plethora of genres, as is not unexpected given a population
wherein even the dullest of healthy individuals would have been
considered quite mentally gifted in millennia past - however, within
Shaper society itself, only the construction of life-art is held in
meaningful esteem. Other venues for expression are welcomed, but
reserved for hobbyists and amateurs, not professionals. To construct
epic compositions in and of the human form is the highest art in
Shaper society. As full of science and practical consideration as
their genepool is, it is also their most holy of pieces of art, and
those who have risen to operate upon it their idols.

There are three common Shaper forms. In decreasing order of
population, these are females, males, neuters. A fourth form, the
'hulk', is only debatably a "Shaper form" and even more divergent from
the human strain than the Shapers themselves - however as they are
non-reproducing, few in number, have limited interaction with the rest
of Shaper society, and are distinct in construction based upon their
version and intended purpose, it is easier to consider them apart from
the Shaper populace and group them with other designer human life
forms, despite being of Shaper origin and mentality. The first
females, males, and neuters are of the same height, and all beautiful
by most standards, providing the lack of hair is not a deterrent to
such a judgment being rendered, nor the lack of nipples on Shaper
males and neuters found distrubing. Shapers are not muscle-bound, but
neither are they in any way waifish, with toned and sleek muscles
making their presence subtly known underneath taut skin - their
physique sometimes evokes images of acrobats, swimmers, and dancers,
depending upon which portions thereof were being used for the
metaphor. All Shapers are in prime physical condition both for
strength and endurance, although there exist plenty of custom bio-mods
who are stronger or have higher endurance. Shapers remain in prime
condition without need for exercise. Neuters carry both male and
female genetic code, but do not express either set of reproductive
organs or secondary sexual characteristics, earning them nicknames
associated with old plastic dolls. Instead of physical features to
distinguish Shapers from one another, with their near-singular skin
and eye color and lack of hair to style, Shapers tend to use clothing
and other symbiotic adornments when wishing to express their
individuality. Even then, a non-Shaper would find it extremely
difficult to identify which Shaper is which. Although much diverted
genetically from most humans, they are still capable of producing
fertile offspring with other human subspecies. Though a misnomer, as
many of the genes require either a Shaper female's womb or a
Shaper-compatible gestation tank to activate, "half-Shapers" arising
from Shaper males fertilizing non-Shaper females are not uncommon
among most human factions, as restrictions on Shaper procreation
privileges do not extend to the production of non-Shaper children
("half-Shapers" are not eligible to become Shaper citizens and are
subject to the same stringent residency requirements as any other
non-Shaper wishing to settle on a Shaper planet). Far more rare are
"incomplete-Shapers" - those born to a Shaper mother by a non-Shaper
father, or otherwise gestated in a Shaper compatible tank. As the
Shaper genes, when activated, tend to be dominant,
"incomplete-Shapers" tend to much more strongly express their Shaper
heritage, although they are still not legally considered
Shapers. Those with Shaper ancestors are commonly recognizable for
their generally gray to blue-black skin, often violet irises set in
gray eyes, and sparse body hair, although it is difficult, if not
sometimes nearly impossible, to distinguish between those with Shaper
ancestors and those who are the product or progeny of the custom
bio-modified.

Shaper clothing tends to fall into one of two very distinct
categories. Shapers have exceptionally minimal body modesty, and any
wearing of clothing whose primary purpose is to keep portions of the
body hidden from sight is done so on account of the modesty of
non-Shapers with whom they must interact - after all, the Shaper body
is a work of art and science they are proud to display. Shaper
work/functional clothes, therefore, act primarily to protect more
sensitive surfaces, to keep softer tissues in place during
motion/exertion, and to provide convenient places to store objects
being used. Such clothes are considered artful to the degree that they
do not interfere with the natural Shaper form and movement thereof. On
the other hand, as Shaper mastery both of their own body and their
environment is very advanced, with clothing for them somewhat
superfluous save for statement anyway in most climes and many
conditions, Shaper art-wear is something else entirely. High-end
Shaper art-wear often takes the form a living, generally translucent
expanse of vegetation designed at the protein and genome level by the
wearer - anything from a thin film, to a sheet-like wrap, to a
crawling expanse of vines, flowering if watered. Those less talented,
but financially equipped, will pay top credits to be similarly
designed for and decorated. Those in neither situation may opt to
obtain a license to perform minor customizations on existing designs,
or, for the even lower social rungs, to simply wear something mass
marketed - the Shaper equivalent of a cheap suit. Less common, but far
from rare, are non-vegetable incarnations of the above. Also common in
Shaper fashion are full-body markings and bioluminescent symbiotes
worn in lieu of jewelery. Full-body markings are normally applied by
developing pore patterns in body-sized engineered leaves that will
leave temporary colorations when wrapped around the body. Shaper
art-wear, though itself only exported as a luxury item, has become
very influential to Highborn fashion. Luxury items such as art-wear
and living rugs account for a noteworthy portion of Shaper exports, in
keeping with the unmatched tangible beauty of Shaper art. This
tangible and enduring nature, and it's profitable export, stands in
stark contrast to the only other major society with a comparable
number of artists, the Andolians, whose art the link-blind cannot even
sense, and whose works come and vanish on whim.

The Shapers are quietly contemptuous of most other meme-groups. They
find the Purists shameful, the Unadorned defective, the Highborn
arrogant and outmoded, the Merchants merely an unfortunate necessity,
and the Mechanists downright repulsive. Though they disagree
vehemently with the Andolians on a number of fundamental premises and
future directions, they respect them for their accomplishments. The
Shapers find the Rlaan interesting, the Uln pitiful, and the Aera
worthy, if unsympathetic, adversaries, if hopeless materialist
aesthetes incapable of appreciating life in any way valuable to the
Shapers.

The Shapers, with their superhuman aims, were a pre-Diaspora offshoot
of what later became the larger, supra-human aiming Lightbearer
meme-group. The two groups had wildly diverging goals and methodology,
despite both seeking a sort of embodied human perfection. Expedience,
however, makes strange bedfellows. Through location and shared
technological expertise, each group had something to offer the other,
even if each found the other ideologically perverse. Thus, what began
as a minor Shaper investment into a Lightbearer mission that would
bear closer to Shaper than Lightbearer space blossomed into a wary
partnership when the discovery of the ancestral Dgn made the Shaper
experience with non-human genetics immediately more valuable. However,
the partnership with the Lightbearers was not particularly long-lived,
as ideological differences made it exceptionally difficult for either
side to continue to operate in good faith absent obvious mutual
advantage, and by the time the aptly named Beckett's Murky Venture had
been exploited, there was little obvious advantage to continued
pretense of cooperation. Initial Shaper political, financial, and
resource support for the Lightbearers during the Fraternal War was
linked not to political or ideological support of the Lightbearers,
but a firm opposition to what they felt to be Andolian Imperialism and
the desecration of the natural sovereignty of the Lightbearers -
clearly, it didn't help that the Shapers thought the Andolians to be
even more ideologically disturbed than the Lightbearers; while the
Shapers thought the Lightbearers perversely misguided in their vision
of embodied human perfection, it was at least a goal that resonated
with their society, whereas the Andolian lust for abstract and
vaporous group progress that would somehow lead to a "bettering" of
mankind and thereby the individuals thereof struck the individualist
Shapers as a demented non-sequitur that was downright
unintelligible. The interjection of an expanding, and, if human,
foreign in an almost alien sense entity into a neighboring polity was
clearly cause for alarm, never mind the Klk'k - they, after all,
weren't human. That the Shapers cut off all support for the
Lightbearers after the revelation of the existence of the Spaceborn
indicates, importantly, that, at the time, they either thought us, or
themselves (depending upon one's perspective) to still be human, and
thus deserving of the fundamental human rights that run through the
core of their belief system. While some Lightbearers did flee to some
degree of protection within Shaper systems toward the end of the war,
it was that same sort of protection offered Nazi scientists captured
by the Russians after World War two - protection of ones life, but not
an opportunity to live, only an opportunity to be worked until
whatever useful knowledge you had could be extracted and you yourself
forgotten. Much has been rumored, but little ever confirmed about what
exactly occurred when the Simons later conducted raids into labs
believed to be housing unaccounted for Lightbearers in Shaper
space. It is well known that the Shapers were absolutely livid over
the Lightbearers' use of human slaves, and it is suspected that a far
darker hell awaited those who fled to Shaper space than they expected.

Notably, the Shapers never gave any indication that they were incensed
by the detailed descriptions of what the Lightbearers had subjected
the Shmrn to, and it is generally conjectured that they were privy to
that information long prior to its distribution by the
Andolians. Apparently, giving the Shmrn a strongly defined sense of
self, dignity, modesty, and a mind muddied by the introduction of
human derived genetic codes for brain development, all in the name of
allowing the Shmrn to truly appreciate and understand the physical and
emotional torture and suffering being inflicted upon them by the
Lightbearers in a way that the Lightbearers themselves could
understand, was not cause for outrage in the Shaper population because
the Shmrn weren't human. This lack of response, though attributed by
some as merely arising from their part in creating the Shmrn, led many
to believe that, if and when the Shapers considered themselves or the
rest of humanity to no longer be human, similar indifference could be
expected to follow were anything to come to affect the rest of
humanity differently than the Shapers. The topic of rights for
associated non-human species has continued to be a topic of contention
between Confederation factions, with the Andolians and the Shapers
tending to lead very different camps in such discussions. Likewise,
even on the topic of human rights, the different principles held by
the Andolians and Shapers has led to rather different stances being
taken on a number of policy decisions. On the topic of human slavery,
however, the two groups stand united, a unity that was formative in
the early Confederation.

Human slavery, as a viable economic pillar, died with the industrial
revolution. Human slavery, for personal pride, amusement, or
perversion, was reborn with mastery over the genome. No longer was it
necessary to steal away random human flotsam and unprotected child
assets, instilling fear and bringing hatred upon the slave trader -
instead, human flesh could be molded into whatever form the customer
desired, and raised from birth into servitude, birthed from a machine
without commentary on the subject. It became difficult in some places
to distinguish between slavery and child abuse. Such undertakings
remained quiet sorts of enterprises until after humanity had finally
ridden nano-technology and genetics to a post-sufficiency economy. In
the introverted, contented meme-group world of that past, the only
objection to someone outside one's meme-group raising slaves was one
of purely moral basis - it was simply unlikely to affect those outside
the participating meme-group. As SuSims, human-derived PAIs, and other
genetically engineered servant species became more increasingly
prevalent, it was increasingly difficult to gain a motivating
consensus on the issue when, despite many believing that human slavery
was over some line, agreement on where in the gray murk such a line
existed was not forthcoming.

It was not until the formation of the Confederation that such things
truly changed. All were free to not do business with polities allowing
behavior deemed "objectionable", but with the Confederation, a
sufficiently forceful hand could be put upon the hilt of a very sharp
dagger that could be aimed, if in unwieldy and cacophonous fashion, at
those sufficiently outside the mainstream of the relevant political
bodies. Human slavery was outlawed. Certain forms of indentured
servitude, especially as concerned AIs, were not outlawed, given
certain contract conditions, much to the dismay of the Andolians and
the relief of other parties unwilling to emancipate their few AIs that
survived the nano-plague. Vagaries in defining both "consenting" and
"adult" as they were framed in the stipulations concerning indentured
servitude - the Shapers being appeased by such, as they believed it
the right of any adult to enter into any contract that did not
permanently disenfranchise them, if the contract was entered into in
an informed and free fashion - have continued to allow fringe
operators to continue practices against the spirit if not the letter
of the laws. However, those who are foolish enough to flagrantly
flaunt their slim loophole existence generally find themselves
ruthlessly and utterly squashed by Andolian and Shaper forces
comprising the military wing of the Confederation taskforce on human
slavery - which often delights in opportunities to expand the
operations slightly to associated enterprises, which often are
subsidiaries of otherwise difficult to legally attack organized
criminal undertakings (especially those with good standing with
CMT). Even those guileful and careful enough to stay within the
loophole lines are sometimes paid visits by the Simons. Strangely,
there aren't as many complaints as one might expect - some things just
become a cost of doing business. Indeed, in polities where the Simons
can move freely, human merchandise tends not to move much at all - but
then again, in places where the Simons can move freely, human
merchandise tends to not be much desired.

\subsubsection{A note on the Dgn within Shaper society}
The Dgn were never really emancipated, because the Shapers don't
consider them to have ever been slaves as such - you have to be human
first to be a slave in the Shaper mind. Dgn are just useful art that
talks, and have been constructed to play particular roles in Shaper
society - the Shapers view them in a way not dissimilar to how AIs are
viewed by many LIHW, Forsaken, and some Purists: thinking machines,
that do work on our behalf when properly treated and guided. The
Shapers have no particular respect for the status of uplifts -
somewhat more for those naturally arriving at sapience, and far less
for AIs - but their quest for superhuman perfection has left them
interested in themselves and what other species can offer them in
terms of genes and understanding thereof, but not in other species, or
even so much in other humans except as if furthers their goals or
offends their tenets. The Dgn are treated with indifference to what
others may believe their "rights" rather than cruelty in oppression.

\subsubsection{Dgn}

Faction data 
Dgn 
Species 	Dgn 
Homeworld (Origin) 	fixme 
Capital 	fixme 

Need a desc for controlled space and space faction strength status/goals FIXME 

\subsection{Unadorned}

Faction data 
Unadorned 
Species 	Human 
Homeworld (Origin) 	Earth 
Capital 	fixme 

The Unadorned worship the cold logic of machines, and seek to abandon
the limitations of human emotion and irrationality. Often referred to
as the "mad monks of Myztheria", the fact that the other factions
believe them to be somewhat insane has not stopped them from making
use of their advances in AI and other computational fields. Indeed,
the Unadorned find religious veneration of computational logic
entirely... logical. Motto: Freed of noise, the mind will make true
music.
 
\subsubsection{Mishtali}

Faction data 
Mishtali 
Species 	Mishtali 
Homeworld (Origin) 	fixme 
Capital 	fixme 

Politically subserviant to the Unadorned, the Mishtali have
significant control over affairs in their home system, but have little
impact on external affairs.

\section{Rlaan Assembly}

Faction data 
Rlaan 
Species 	Rlaan 
Homeworld (Origin) 	4th planet of 
Capital 	
FIXME 
 
\subsection{Rlaan Enforcers}

Faction data 
Enforcers 
Species 	Rlaan 
Homeworld (Origin) 	SCx9362 
Capital 	probably not same 

Vaguely analogous to a policing force, the Rlaan Enforcers keep are
the armed group responsible for keeping order within the confines of
the Rlaan Assembly's domain. Their most active utilizations are as
customs/border patrol agents, anti-piracy patrols, and garrison
duties.  FIXME (more would be nice here)
 
\subsection{Lmpl}

Faction data 
Lmpl 
Species 	Lmpl 
Homeworld (Origin) 	fixme 
Capital 	4th planet of SCx9362 

Politically impotent members of the Rlaan Assembly, they are not
considered citizens. Indeed, the Rlaan view them more as intelligent
tools. They are, however, steadfastly loyal to their Rlaan masters,
and not prone to complaining about their servitude, nor, for that
matter, anything else unless it is disruptive to whatever task at hand
they have fixated upon. They are heavily used throughout the Rlaan
civilian sector in operations on Oxygen-Nitrogen worlds, crewing their
own Oxygen variants of Rlaan civilian craft.
 
\subsection{Nuhln}

Faction data 
Nuhln 
Species 	Nuhln 
Homeworld (Origin) 	fixme 
Capital 	4th planet of SCx9362 

Unlike the Lmpl who have some degree of independent existence, working
on worlds inhospitable to the Rlaan, the Nuhln are completely subsumed
by the Rlaan culture they are surrounded by. Their existence,
somewhere between pets and paper-pushers, is not one that can be
readily disentangled from that of the Rlaan they serve. What
self-governance they exert is more akin to a union than a government,
existing primarily to ensure continued reasonable working conditions,
and created by Rlaan edict rather than arising internally.
 
\subsection{Rlaan Briin}

Faction data 
Rlaan Briin 
Species 	Rlaan
Homeworld (Origin) 	fixme 
Capital 	fixme 

The Rlaan Briin are an anthrophilic Rlaan subculture. Some are so
interested in studying humanity "in the wild" that they perform
full-body transplants into anthropomorphic self-contained encounter
suits so as to live in human habitations.
 
\subsection{Rlaan Hunters}

Faction data 
Rlaan 
Species 	Rlaan 
Homeworld (Origin) 	4th planet of 
Capital 	

Formed post contact with humanity and modeled on the human Hunter
faction, the Rlaan Hunters are almost entirely composed of sterile
Worker/Defender hybrids who failed in the political arena. Although
Rlaan culture was too coherent to have had to deal the issues that
plagued Human space, these Rlaan were quite happy to offer to try out
their solutions - for a fee.  Rlaan Merchants

\subsection{Rlaan Merchant Steering Committee}
Faction data 
Rlaan Merchants 
Species 	Rlaan 
Homeworld (Origin) 	SCx9362 
Capital 	fixme 

The Rlaan Merchant Steering Committee acts to coordinate various Rlaan
merchant groups to supply the Rlaan worlds and military with needed
and desired goods both local and foreign, although it only exercises
notably active control over foreign trade. Like all Rlaan civilian
craft, Rlaan merchant ships are entirely unarmed, though heavily
armored and shielded, are crewed entirely by workers, and are
protected via a convoy system.
 
\subsection{Saahasayaay}

Faction data 
Saahasayaay 
Species 	Saahasayaay 
Homeworld (Origin) 	fixme 
Capital 	fixme 
FIXME 
 
\section{Forsaken}

Faction data 
Forsaken 
Species 	Humans 
Homeworld (Origin) 	Earth 
Capital 	Ajani 


Overview 

The "Forsaken", as they are collectively known, are the
descendents of various victims of one of the more tragic eras of
Humanity's expansion into space. Six centuries ago, slowboats were
made obsolete by the development of FTL travel. However, this didn't
help those colonists already en route, who, upon reaching their
destinations, found the worlds slated for their colonization already
inhabited by settlers that had leapfrogged them in humanities
continued outward expansion. Arriving hopelessly out of synch with the
rest of human society and finding themselves deprived of both their
worlds and nanite-based technology, these groups found they had more
in common with each other than anyone else, and set out to colonize
the worlds no one else had wanted to.

A minor faction due both to the initial power gap and the consequent
constant state of "playing catch-up", they were not asked to
participate in the conferences that begat the Confederation, nor would
they have accepted, having no love for the powers that had done
nothing to stop the leapfrogging of colony worlds that had been slated
them. Likewise, as the dominant strains of thought at the time within
Forsaken space considered the Confederations's protection would be at
best hollow, and at worst a pretense for tyranny, the Forsaken
declined an offer to join the LIHW, and instead focused on settling
the Diaspora sector, keeping their distance, to the best of their
ability, from the developing Confederation.
 
See also : ''On the origin of The Forsaken'' 

\section{Shmrn}

Faction data 
Shmrn 
Species 	Shmrn 
Homeworld (Origin) 	Beckett's Murky Venture 
Capital 	fixme 
FIXME 


\section{Uln (Sul-Gatwa Dynasty)}
Faction data 
Uln 
Species 	Uln 
Homeworld (Origin) 	fixme 
Capital 	fixme 
FIXME 

\section{Interstellar Church of True Form's Return}

Faction data 
Luddite 
Species 	Human 
Homeworld (Origin) 	Earth 
Capital 	Unknown 

Originating as an extremist faction of what was then the growing
Purist movement spawned by the nano-plague, the blatant intolerance of
their doctrines and violent terrorist tactics have put them in poor
standing with every governmental entity worthy of being called
such. Often referred to in derogatory fashion by the misnomer
'Luddites' (it isn't directly applicable because they aren't directly
anti-technologists - their beliefs revolve around returning from the
distinctly altered forms of human existence that, enabled by
technology, humans have explored) the members of this organization are
hunted throughout civilized space, except where given shelter by
extremist interests within the Purist faction.

 
\section{Interstellar Socialist Organization}

Faction data 
Interstellar Socialist Organization (ISO) 
Species 	Humans 
Homeworld (Origin) 	Earth 
Capital 	Somewhere in Forsaken space. 

Undoubtedly one of the oddest memegroups, the Interstellar Socialist
Organization (ISO) seeks to return back to the workers of the
inhabited galaxy control over the means of production. Notwithstanding
the fact that this has more or less happened in some other memegroups,
nor that some others no longer have an economy where this arrangement
is coherrent, the ISO still seeks to have everyone join their utopian
dream.

What the ISO lack in numbers they attempt to make up for in
determination. Unhappy with the dominant free-market capitalist
process employed in the majority of human space, the ISO have turned
to less savory methods to achieve their goals. Being primarily a
paramilitary group, they now focus on industrial sabotage and
political destabilization efforts. Although they heavily shy away from
excessive civilian body-counts, never resorting to anti-civilian
terrorism, they do try to suborn the citizens of the Confederation at
every turn.

With no active control over any planets or industrial sectors, their
resources have been limited. And although they do share a few
co-dominant meme planets, mainly within Forsaken space, their
infrastructure remains minimal and most of their equipment and ships
must be acquired, purchased or stolen from outside suppliers (For
example, the Sickle and Hammer ship designations are post-facto
nicknames applied to ships formerly known as the Goose and the Toad
respectively, before significant ISO acquisition and usage).

Economic backing, what of it there is, is believed to come primarily
out of various ISO communities in Forsaken space and locally targeted
resources funneled from other groups who benefit from destabilizations
in various regions of Confed space. The ISO survives because they have
been long perceived as more a nuisance than a dire threat, and because
the Confed has been unwilling to expend the political capital
necessary to undertake such large scale military operations in
Forsaken space, especially considering the longstanding bitterness of
the Forsaken has held for the Confed.

Due to prompting from major factions intensely opposed to their
economic vision, the confederacy has banned membership in this
memegroup.
 
\section{Lightbearer}

Faction data 
Lightbearer 
Species 	Lightbearer 
Homeworld (Origin) 	Earth 
Capital 	Cradle (defunct)
(Defunct Meme-group comprised of members of the "suprahuman" strain of humanity) 

From a planet discovered on a joint mission with the Shapers the
Lightbearers cultivated some of the early Dgn, further twisting their
development along their own designs until they settled upon what are
now the Shmrn.

When later a Lightbearer exploratory group discovered the Klk'k
homeworld they prepared to subjugate, and "correct" this species,
whose somewhat anthropomorphic construction they viewed as a mockery
of the divine human form. Fortunately for the Klk'k, the Andolians
discovered them only some weeks after the first Lightbearer
pacification vessel had arrived at the planet. The Andolians were
outraged at the Lightbearers' actions against civilized sapients, and
threw themselves against the Lightbearers in the first full-fledged
human interstellar war.

The Andolian's greater manufacturing base outperformed the
Lightbearer's economy. As the Lightbearer military-industrial complex
collapsed, a secret was revealed that not even the Lightbearer allies
had known about, namely, the existence of the Spaceborn a genetically
engineered slave race of humans, designed to live their lives in zero
gravity so as to prevent Lightbearer from having to deal with such a
menial task as laboring in vacuum.

It is this revelation that is believed to bear primary responsibility
for the lack of action taken by any other faction when the Andolians
proceeded to eliminate not only the industrial capacity of the
Lightbearer, but also the Lightbearers themselves.

Those that did not manage to escape to Shaper or Highborn space, or
were not fortunate enough to be killed in the assaults on their
worlds, had the dubious honor of being turned over to the Klk'k, the
Spaceborn, and the Shmrn. The ensuing combination of sterilization and
incarceration served to eliminate the Lightbearer meme from the realm
of dominant thought.
