\section{Gameplay}
\subsection{Objectives - single player}
  What are the objectives of the game?
\subsection{Objectives - multiplayer}
  What are the objectives of the game?
\subsection{Game Progression - single}
\subsection{Game Progression - multi}
\subsection{Mission Structure}
\subsection{Career Structure}
\subsection{Play Flow}
  How does the game flow for the game player
\section{Mechanics}
  What are the rules to the game, both implicit and explicit.  This is the model of the universe that the game works under.  Think of it as a simulation of a world, how do all the pieces interact?  This actually can be a very large section.
\subsection{Physics}
  How does the physical universe work?
\subsection{Movement}
\subsubsection{Normal Space}
\subsubsection{In-System FTL}
\subsubsection{Inter-System FTL}
\subsection{Objects}
\subsubsection{Cargo}
\subsubsection{Debris}
\subsubsection{Resources}
\subsubsection{Passengers}
\subsection{Actions}
\subsubsection{Docking}
\subsubsection{Picking up objects}
\subsubsection{Talking to NPCs}
\paragraph{Ship-to-ship}
In-flight comms.
\paragraph{On base conversations}
Primary mission-initiation dialog, etc.
\subsubsection{Talking to Players}
\subsubsection{Libraries}
\subsection{Combat}
How is this specifically modeled?
\subsection{Economy}  
What is the economy of the game? How does it work?
\section{User Interfaces}
\subsection{Screen Flow Chart}
A graphical description of how each screen is related to every other
\subsection{Screen Descriptions}
What is the purpose of each screen?
\subsubsection{Main Menu Screen}
\subsubsection{Options Screen}
\subsubsection{In-Flight HUD}
\subsubsection{Base interface}
\subsubsection{Trade screen}
\subsubsection{Upgrade screen}
\section{Game Options}
  What are the options and how do they affect game play and mechanics?
\section{Replaying, Saving, and Death - single player}
\section{Replaying, Saving, and Death - multiplayer}
\section{Cheats and Easter Eggs}
