\nomenclature{FTL}{Faster Than Light}

\nomenclature{VSU}{Vega Strike Universe}

\nomenclature{UtCS}{{\it Upon the Coldest Sea}. A time period in the VSU starting around 3276 CE. Also, a game using the Vega Strike engine set in said time period.}

\nomenclature{Rimward}{Generally, in the direction of the galactic rim. Alternately, away from the galactic core. The latter definition is more sensible when considering directions relative to extra-galactic locations.}

\nomenclature{TWHON}{Acronym for ``Those Who Have Only Names''}

\nomenclature{PAI}{Pseudo-AI. A near-AI construct that performs complicated tasks and interactions while lacking actual sentience.}

\nomenclature{Anthrophilic}{Pleasantly disposed towards humans or human culture.}

\nomenclature{Xenolinguistics}{The study of alien languages.}

\nomenclature{APSWAK}{The Andolian Protectorate Space Warfare Academy, Ktah}

\nomenclature{Ktah}{Klk'k homeworld.}

\nomenclature{Kubernan}{The Andolian homeworld}

\nomenclature{Aera}{One of the primary VSU sentient species}

\nomenclature{Deucalion}{The player character in UtCS}

\nomenclature{Uln}{One of the VSU sentient species}

\nomenclature{Andolian}{Human meme-group, primarily composed of Pluralis variants}

\nomenclature{Mechanist}{Human meme-group, primarily composed of Cyberis variants}

\nomenclature{Klk'k}{Vaguely anthropoid sentient species from Ktah. Autonomous clients of the Andolian Protectorate}

\nomenclature{Purth}{Massive, cybernetically enhanced uplifts developed by the Andolian Protectorate. Subservient clients of the Andolian Protectorate}

\nomenclature{LIHW}{League of Independent Human Worlds. A political umbrella group for assorted minor human meme-groups}

\nomenclature{Rlaan}{One of the primary sentient species of the VSU. Ammonia-blooded methane breathers}

\nomenclature{Shmrn}{Sentient species in the VSU. Descendants of the Lightbearer derived variants of the ancestral Dgn.}

\nomenclature{Dgn}{Sentient species in the VSU. Shaper uplift of pre-sapients from Beckett's Murky Venture}

\nomenclature{Beckett's Murky Venture}{The homeworld of the ancestral Dgn. Reference to a joint Lightbearer-Shaper exploratory mission}

\nomenclature{Saahasayaay}{Client species of the Rlaan.}

\nomenclature{Alphan}{Of or concerning the first known species to achieve interstellar spaceflight in the post-Ancient era}

\nomenclature{Betan}{Of or concerning the primary rivals to the Alphans}

\nomenclature{UniServe}{Andolian Universal Service requirement}

\nomenclature{Amakakt}{Klk'k martial art featuring the Tklatl}

\nomenclature{Coreward}{In the direction of the galactic core}

\nomenclature{Lightbearers}{Human meme group composed of Suprahomo variants. Cultural memes extinguished and gene-pool nearly exterminated as a result of Fraternal War}

\nomenclature{Oolak'kl}{A Klk'k beverage analogous to beer.}

\nomenclature{Kt'totha}{A large aquatic predator native to Ktah, once prized for it's hide}

\nomenclature{Rlaan-Briin}{An anthrophilic movement of the Rlaan}

\nomenclature{Retcon | Retconning}{Retroactive Continuity. The retroactive application of changes to canon to enforce a coherent/consistent world view}

\nomenclature{Lmpl}{Client species of the Rlaan. Oxygen breathers, generally deployed as planetside laborers}

\nomenclature{Nuhln}{Client species of the Rlaan. Environmentally compatible with the Rlaan}

\nomenclature{Nanite}{An independent machine constructed and manipulating matter at nanometer scales. Not to be substituted for the broader field of nanotechnology, which deals with all operations and construction occurring at similar scales, not just nanites.}

\nomenclature{Nanotechnology}{The study, practice, and implementation of technology dealing with with construction and manipulation of structures at nanometer scales.}

\nomenclature{Nano-plague}{The somewhat inaccurately named phenomenon
of the activated immuno-defense systems left over from the fall of the
TWHON disassembling nanites. Alternately, the immuno-defense systems themselves}

\nomenclature{Mishtali}{A client species of the Unadorned. The first sentient species to be encountered by humanity.}

\nomenclature{Bzbr}{A client species of the Aera}

\nomenclature{Memetic}{Of or relating to memes}

\nomenclature{Xenocide}{The deliberate and systematic destruction, in whole or in part, of an alien group}

\nomenclature{Cryoship}{A sub-light colony vessel relying on cryogenic storage of its passengers or their genetic material as the primary means of carrying colonists.}

\nomenclature{Terraforming}{The process of transforming a planet in a manner that renders it more Earth-like}

\nomenclature{Kleptocracy}{Literally, rule by thieves. Generally, an ostensible government that exists primarily to extend the personal wealth and power of the ruling class at the, often direct, expense of the population.}

\nomenclature{NeoBarbarians | NeoBarb}{A group that has, usually in isolation and under prolonged economic duress, diverged sufficiently in its cultural norms and expected standard of living, that they are considered barbarians relative to the politically dominant cultures}

\nomenclature{Sundered | the Sundered}{Of or relating to systems colonized but not actually on the jump network, or the populations living therein}

\nomenclature{Sheltersky}{Homeworld of House Blythe}

\nomenclature{Sofono}{Emilio Sofono, inventor of the human variant of the SPEC drive. Alternately, any human produced SPEC drive (as 'Sofono drive'), or a class of Andolian warships named in his honor (as 'Sofono class').}

\nomenclature{Sul-Gatwa}{The current and long ruling Uln dynasty}

\nomenclature{Ingatwa}{Name associated with the Ulnish aristocracy}

\nomenclature{Bifrost}{Shaper homeworld}

\nomenclature{Tklatl}{A body-length stick with bladed ends used in the traditional Klk'k martial art of Amakakt}

\nomenclature{CMT}{Cherryh Mercantile Trust. Precursor to the ISMG}

\nomenclature{ISMG}{Interstellar Shipping and Mercantile Guild}

\nomenclature{Simons (the)}{A paramilitary group operating with the tacit approval of the Andolian Protectorate, self-tasked with oversight of the remaining Lightbearer populations and the tracking and pursuit of war criminals and known terrorists.}

\nomenclature{Fraternal War}{War primarily between the Lightbearers and the Andolians, circa 2988-2994 CE. First large-scale interstellar conflict between human factions post invention of SPEC drive}

\nomenclature{Meme}{}

\nomenclature{Meme-group}{}
%\nomenclature{}{}


% LocalWords:  FTL VSU UtCS Rimward TWHON PAI Anthrophilic Xenolinguistics Ktah
% LocalWords:  APSWAK Andolian Klk'k homeworld Kubernan Aera Deucalion Uln LIHW
% LocalWords:  Purth Rlaan Shmrn Dgn Saahasayaay Alphan Betan UniServe Amakakt
% LocalWords:  Tklatl Coreward Lightbearers Oolak'kl Kt'totha Briin Retcon Lmpl
% LocalWords:  anthrophilic Nuhln Nanite nanites Mishtali Bzbr Memetic Xenocide
% LocalWords:  Cryoship Terraforming Kleptocracy NeoBarbarians Sofono Pluralis
% LocalWords:  Sheltersky Sul Gatwa Ingatwa Ulnish Bifrost CMT Cherryh ISMG
% LocalWords:  Simons Lightbearer Cyberis cybernetically sapients Alphans
% LocalWords:  Suprahomo Retconning planetside NeoBarb Andolians
