\section{Dramatis Personae}
\begin{itemize}
\item Deucalion (Player Character)

Deucalion is a human, albeit of heavily modified genetic stock and
augmented with numerous implants (the latter being the norm among the
citizens of the Protectorate, and the former being a somewhat
distinguishing, albeit hardly unique, feature). He was raised from
near-infancy by Klk'k on Ktah after the untimely demise of the craft
carrying him and, so far as could be discovered, anyone with any claim
to him.  He remained on Ktah throughout his youth, though visits
alongside his adoptive family to other Protectorate worlds were not
unheard of, if infrequent.  His first long term separation from Ktah
and family came when he attended First University on Kubernan. Having
completed his studies in Computer Science, Historical Analysis, and
Xenolinguistics, he returned to Ktah for his Universal Service
requirement, his aptitude exam placing him into Officer's training at
APSWAK (Andolian Protectorate Space Warfare Academy at Ktah), where he
was trained as a pilot.

At APSWAK, Deucalion met Lauktk, a Klk'k flight mechanic who remains
assigned to Deucalion's wing when they both go into active duty.  They
become very close friends, and Lauktk later becomes a bond-mate of
Deucalion's adoptive sister.  Upon discharge, Deucalion is recruited
to return to APSWAK as a flight instructor, while Lauktk takes on a
position as a starship mechanic working for the Protectorate Fleet
shipyard orbiting Ktah.  After a few years of saving up, Lauktk has
accumulated enough resources to purchase his own vessel, and contacts
Deucalion, who agrees to pilot the vessel.

At the start of the game Deucalion is ~28 years of age.  He has left
his position as a flight instructor at APSWAK to pilot his best
friend's ship, with the first launch of the vessel occurring only a
deci-year ago. However, only a week into their travels, a Luddite
attack crippled the ship and forced an emergency landing on an
underdeveloped colony world in Cephid 17 that neither Lauktk nor much
of the cargo survived. A few weeks later, ownership and insurance
issues resolved, the ship has been moved and repaired, and what
systems that time and money could afford have been replaced. Likewise,
in large part due to his genetic modifications, Deucalion has mostly
healed physically, if not mentally.

Finally ready to return to space, he carries with him the cremated
remains of Lauktk and the far heavier burdens of guilt and the
untimely broken dreams of the being that was his closest friend.

\item Lauktk (deceased) (Klk'k) 

Deucalion's best friend, flight mechanic and
companion throughout their stint in the Protectorate military.

\item Mai (Klk'k)

Deucalion's adoptive sister, a Klk'k.

\item Mirabel (Human, designer genome)

One of Deucalion's former squadmates. 

Hers is a striking figure, a dark-elf visage conjuring the essence of
a Boris Vallejo piece writ in flesh. From indigo skin to hair
equivocating with metallic sheen between purple and white, she is
clearly a work of gene-smith's art. There is, however, neither love
nor thanks to be had for her crafters. She was commissioned by sexual
sadists with pockets deep enough for expensive toys and fashioned by a
business group willing to sell scruples and humans alike. Forunately
for her, she was liberated from the facility where she was being
raised before her erstwhile owners could collect her. She, along with
a number of other constructed beings gained their freedom when the
complex was raided during the Blooding of the Purth some thirty years
ago. Mirabel has spent her entire adult life in service to the
Andolian Protectorate. She first met Deucalion during his UniServe,
when they were both posted to the same vessel.
\end{itemize}

{\bf Minor Characters}

\begin{itemize}
\item Jenek (Human) 

\item Simon XII
\end{itemize}

\section{Scenes out of time and sequence}

\begin{itemize}
\item Deucalion and Mirabel after Lauktk's death

"Mommie Dearest": AP slang/jargon for the Andolian Protectorate Ministry of Defense.

Deucalion is sitting alone at a table in a spaceport bar/restaurant,
silently watching the flow of patrons, passively absorbing nearby
conversations, and working with only mild interest through his
meal. Mirabel walks in and sits down at his table.

Deucalion (without particularly looking up): "I see you were on
Kubernan when the Aera decided to wander coreward - or was it
Hephaestus? Either way, not like you to be so far from the front."
(Turning his head up to look directly at her) "Less like you to be
here now."

Mirabel: "Is it that obvious?"

Deucalion (with the first signs of friendliness to his response):
"No." (Smiling) "Only someone already familiar with how quickly your
hair reaches shedding length could tell how long it's been since your
sacrifice to the guardians of sterility."

M: (sincerely) "How - how are you... ?"

D (shaking his head): "I've been better, but I'm sure Mommie Dearest
didn't send you here for a hand-holding session."

M: (More formally, with slight avoidance) "It was on the way. And I do need an answer - how are you?"

D: "As a friend, or as an officer?"

M: "My time is too short right now to be anything but both."

D: "Then I'm not ready to say yet. I need to work through this before
I can tell you which of the other sides I'm going to come out on. The
timing is lousy, but, some things set straighter and more quickly than
others."

M: "I understand."

D: "Do you?"

M: "We miss him too."

D: (with a lopsided grin) "Sometimes I think it's only when you say 'we' that I can really trust you mean 'I'."

M: (With wry humor) "{\em WE} understand. But the timing is all the more unfortunate..." (getting up from the table) "When you return to Ktah, I'll be waiting for you. We'll talk then."

D: "No, you won't. You'll be there, but you won't be waiting, at least, not waiting for me."

M: "You know I - you know, Deucalion, you KNOW." (brief pause) "I have to go. "

D (as she turns to leave): "Mirabel - it was good to see you."

M: (turning back) "And you. But, be well - I'd prefer to remember you when you smiled more easily." (Mirabel smiles slightly, and leaves)

\item Deucalion and K'hoeama, when Deucalion first approaches his ship in the hangar

K'hoeama: Hey, you, HU-NAM! Nice tats. (Points at the Llama) That our ride?

Deucalion: Doctor K'hoeama, I presume.

K: Please, call me Larry. No one else does, but I'm a Klk'k of leisure and I've been told it suits me.

D: ...

D: You do realize I'm not even going to try to figure out what century's archives I'd have to search to comprehend that joke

K: Bah! Lauktk always said your sense of humor was underdeveloped. Still -- (K'hoeama flashes a toothy smile and embraces Deucalion, arm to arm) 

K: It's good to see you up and hopping. From the crash report I expected something worse, but you look at least half as hale as when I saw you last at the bond-set festivities. 

\item Interlude 0: Dreams and Nightmares of Our Fathers Passed

{\center {\it Turn and face the strange

Ch-ch-changes

Just gonna have to be a different man.}}

\centerline{David Bowie - Changes}

[Deucalion] [overlap with credits?? - an interlude?]

Memories. Old. Older. Ancestral. External.

They taste different.

Dreams, like memories, passed down from parent to child. External, yet encompassing - when co-opted, personal.

Dreams. False memories, near memories, fantasies pleasant, surreal - tasting of incompleteness.

The Klk'k do not dream, their unconscious landscapes barren, the troubles of their worlds processed only in lucid detail.

I must have confused them - scared them perhaps.

Nightmares. 

External, induced

Remembered, relived

Archetypal, developmental

Abstract.

Remembered, relived - recurring

A personal taste

My nightmares have always tasted of memory.

My memories

EVENTS IN INTERLUDE 0

Deucalion recovers from the events that killed Lauktk

Aera invade Forsaken space
\end{itemize}
\subsection{Intro Monologue}

[A Dead Man's Ship (end of interlude 0)] [A quiet shape is sitting
towards the end of a not particularly crowded bar.  There is a quiet
hum of lazy activity, and the bar is swaddled in the awkward grays of
artificial twilight. Despite an odd hue to its skin, perhaps the
legacy of a Shaper ancestor or the enduring design of some
gene-smith's art, it is clearly human, and clearly lost in
thought. Two glasses sit in front of him, one an offworld ale, eerily
beautiful in the subtlety of its contrasts with those of Terra, the
other a glass of ice water thinly disguised with a mint leaf. Both
glasses are untouched, both literally and figuratively, while the man
stares somehow both at and through them all at once, as if trying to
will them to complete the act of drinking on their own, or demanding
that they return his gaze with submission. The vaguely confused
motions of the barkeep, unsure now of how to serve him, break his
brief trance. The man's shape, the complete discipline of his
stillness only apparent with the liquid nature of his motion, twists
to direct his dark eyes up and across the marble slab, fixing restless
on the earnest host. A slight curl appears at the edge of the man's
lips. The glimmer of a compassionate smile, a subtle smirk born of
hubris, or the joy of a man rescued from his own mind, the expression
is enigmatic, unrevealing, and ephemeral. The mouth opens again and he
begins to speak, directing his monologue towards the barkeep, although
it is not clear that the audience he craves includes anyone beyond
himself.]

"It was a deci-year ago today - well, day/night cycles being skewed -"

A momentary pause as he breathes while consulting his data-link.

"No, still today."

He smiles genuinely at the barkeep now, but it quickly melts as his
face returns to a more melancholy expression. In a tone laced with
moments of nostalgia and of emptiness, he continues.

"It was a deci-year ago today. We were in a bar, a lot like this one,
in the spaceport district outside the capital on Ktah - maybe a little
busier." He looks down at the two glasses. "Same drinks. Always the
same drinks. He - Lauktk - he wouldn't ever order anything else. Said
that he and his ancestors had been drinking this brew and its
ancestors since back when the monkey-boys had been kind enough to
relegate themselves to one planet. Said that if the flavor has never
been quite the same since the Lightbearers nuked some of the best
crop- land into oblivion, then at least it makes clear that, in
comparison to meeting humanity, drinking Oolak'kl is not too bad for
one's health. I... well, with my metabolism - there's never been much
point in alcohol consumption, at least not in such small quantities."
He glances briefly at the contrasting colors of his hand and the
bar. "I can't speak for what the flavor may have once been, but I must
admit I find the current one equally unappealing everywhere I've tried
it.  Admittedly, and this is no offense to your fine establishment,
shipping costs being what they are, I can't say the price is the same
off Ktah as on." He looks down towards the one empty barstool to his
left and slows slightly. "But that doesn't really matter now does it?
It doesn't really matter at all.

"He'd taken me out to celebrate. In a few hours, he'd be living his
dream. The deed transfer had finalized, launch inspections had passed
- he had a ship.  After years of mucking around with ships that only
came to him in sickness and left his hands the moment he had restored
them to vibrant health, he had a ship all his own." His demeanor
intensifies. "You can't know what it meant to him - his family had
been sailors, captains, explorers, and merchants since the Klk'k age
of sail. His own ship - it wasn't just a dream, it was a birthright
delayed only by economics and circumstance. It wasn't about money - he
was a starship mechanic working for the Protectorate at the Ktah
shipyards, I'd gone back to the academy and was working as a flight
instructor - it wasn't about acquiring some status symbol - he put
every credit he'd saved into that ship. It was about freedom, his
freedom to sail a new sort of sea, and I was going to help him. I was
going to helm that ship wherever his freedom took him. Even I can't
claim to know what it really meant to him, and I knew him as well as
any man could. He was my brother in arms. He was a bond-mate to my
sister - I remember the first time I introduced..." His words trail
off beneath a frigid gust of mental anguish. "I haven't been able to
see her in person yet - yes of course I've messaged, but you see - you
have to understand I couldn't leave, I couldn't...  the dream is still
here... you have to understand, if I could have..."

The loss of composure ends even more abruptly than it began, emotions
submitting again to a mind well practiced in the arts of control. "We
met during our UniServe. He'd come to see the human who was making a
run at top rank in Amakakt at APSWAK. Those were the salad days of
blissful denial, when, somehow, the Rlaan-Aera conflict that was
boiling over next to us kept us calm and cool behind the razor-drawn
wall of political detachment from the years of slaughter that were to
come. Diversions like Amakakt were even more important then; we all
had a fair idea of what it would mean when war galloped across our
borders, but we knew the odds were against it happening before our
tours were up. It wasn't a denial born of callousness to our younger
brethren who would find themselves in the positions we'd vacated in
relative safety - we just couldn't spend a decade or so brooding,
waiting for the nigh-inevitable. The Andolian spirit of counter-
empirical-tainted optimism has always saturated the Protectorate, even
if I could never embrace it the way he did, but much as I'd like to, I
can't blame that or them, or anything so simple for his fate, even if
blame had ever been something I desired. No, for Lauktk, a society
smothered in optimism was a boon; it fueled the infectious energy
intrinsic to his demeanor, intrinsic to his existence. He was
irrepressible and genuinely funny - even the Purth thought so - I
liked him right from the start. We were friends right up until we
completed our stint at the academy and shipped out - good friends,
even if our different specializations meant we didn't see each other
all that often.

"When we found we'd be serving together on the same ship, we both were
surprised. When we found out that he'd actually be my flight mechanic,
I was...  I felt comforted somehow, to know that the person upon whom
my safety in some large part depended on was someone I knew beyond the
casual connections provided by our links. It was a fairly small crew,
and most of us were there for several years, so we all got to know
each other more than most Andolian flat mates, but the social
structure is never flat - there are always some people you click with
more than others.  The details are both too many to recount and too
meaningless without context, but by the end of that tour, we were
brothers in all but birth.  We'd been through... we'd been through a
lot together - you'll just have to trust me that it's less cliche than
classified." He pauses, asking some question of himself that remains
unspoken, and, the question seemingly answered, he continues.

"Mai, my sister - that was always a disappointment for her.  It wasn't
that she was jealous that Lauktk and I were close, or had gone on
'adventures' together, it was that we couldn't share them with her. To
not be able to share some of the most formative moments of the
friendship between two people, both of whom she held dear - as a
link-phile, not being able to share like that, it never sat right with
her. She could understand it, rationalize it, accept it in the higher
regions of her mind, but she could never be comfortable with it. She
longed for the day when enough would be declassified that the three of
us could go out to the cabin and just spend a few days reveling in
freedom from the secrets we'd been carrying.  You can't wear a link
and like secrets, it doesn't work that way.  Even carrying the small
ones entrusted to one such as myself, I gained great respect for the
burdens that high command places upon itself. But she never had that
load to bear. She only got to see it second hand, feeling the locks
and filters in our minds that precluded us from divulging, whose
presence trespassed upon her familiarity with each of us.  The cabin,
the three of us - that'll never happen now - and even if the locks all
cleared today, whatever tales I spin can never truly be his story." He
stares directly at the barkeep.  "But as it is, with secrets
unreleased, I am crippled in my ability to even relate the experiences
comprising our time together. At least Mai has her other two
bond-mates to help her through this. I... I have only the ship, as
fresh recovered from her time as an invalid as I mine."

His head lowers and an aura of tired resignation encroaches upon his
face.  "There's a certain chill sometimes, when I can't escape
pondering past, present, and the difference between them.  When my
hands brush across the weld lines that make the paint cringe in the
subtle dismay of a suture, when I see adverts for a new jump drive,
when some phantom process in my brain convinces me I can still smell
the kt'tothan leather that no longer covers the passenger seats, as I
sit here staring at two untouched drinks - I cannot outrun reality,
and I am left with the knowledge that I pilot a dead man's ship.  This
cold clings to one's skin, like some sort of shrink-wrapped leprosy,
and I can't help but wonder if it's going to follow me to every ship
that I'll ever fly." His tone now shifts, with nostalgia being
overshadowed by pained cynicism. "For I can't stop being a pilot - she
beckons, you see; the cold vacuum of space moves her lifeless arm, and
I am compelled to join the legions of ships that, mastless, sail upon
her.  Days at a time, scant body-lengths away from her condemning kiss
- and there he'll be - a ghost from selves by then long past, waiting
for me to join him.  How many times can I refuse before I am forced to
fold?  How many times can I wake from nightmares to find myself still
sane?" He extends his arm, grasping the water in front of him, and
drains half the vessel before gingerly relinquishing his possession of
the glass, his thirst and anger temporarily quenched, one more
thoroughly than the other. He continues, more calmly than before, "Far
more, I hope, than my melodramatic flares would lead you to
believe. I've no urge to dive into those sunless depths.  Happy or no,
I have it better than many, no worse than most. After all, I was
raised by Klk'k; I've had more than enough time to outgrow pitying
myself merely for being human."

Glimmers of genuine physical exhaustion begin to work their way upon
his face, upon his previously impeccable posture. He leans forward,
right arm upon the bar, supporting his furrowed brow with his right
hand, his splayed fingers covering half his face. "However, even
beyond the shallow realm of self-pity, once you've felt the inexorable
grind of the universe's apathy ... it's hard to see much meaning.  You
can't find a way to convince me that there's any particular reason
that I'm sitting here talking instead of him. Fate has no weavers,
only the mad spider of chance and the ever-spinning spindle of time.
Underneath it all we're not even pawns - pawns can become knights or
queens - we're nameless particles in some sick, twisted Brownian
motion colliding every now and again with each other and changing.  We
dream up gods to play with us, if only so we can pretend to be
pawns. We sup on hubris so that we can aspire to have names.  It is
only a question of which dish we choose to partake of. Do we follow
the Shapers and seek to assault the glass ceiling of perfection
without even the knowledge as to what that would mean?  Do we cloak
ourselves with the counter-empirical idealism of the Andolians,
believing that all problems can be solved, and that our ability to
solve will progress indefinitely?  One could retreat to the scared
futility of the Purist's status quo, or, joining the Unadorned or the
Mechanists, give up the pretense of desiring to be human.  Is there
any solace in the Merchants' proud valuing of wealth or the
High-Born's pride in their idiosyncratic conception of nobility? Hell,
one could give up on humanity and go live with the Shmrn, the Uln, or
the Rlaan-Briin. Despite every oath I've sworn, for all that I do care
about the Protectorate, I've never been able to believe - not the way
he could. Not in a way that mattered.

"When I was a child, the dreams of my parents and their bond-set
sustained me."  He smirks. "Those who didn't know better would think
it bizarre to be sustained by Klk'k dreams, given that they don't have
the whole unconscious labyrinth experience during sleep cycles that
humans do.  There isn't a word for such things in any of the native
languages - no, the Klk'k dream is the daydream, the fantasy, the
meanderings of desires projected onto unfolding futures. Strong dreams
that one can feast on.  As I grew older, the dreams of the societies I
lived in sustained me, and when I met Lauktk, his dream consumed
me. Now I guess I have to make my own... but, as crazy as it sounds, I
feel the need to finish his somehow, and without him... I don't know
what that even means, and even less how I'm going to do it. Indeed, as
clouded as any future seems right now, maybe I shouldn't be surprised
the only dreams I find myself capable of having are nightmares."

He is silent for a few moments, perhaps reliving some fragment of a
nightmare, perhaps preparing his next utterance, perhaps both.  His
face, as still as carved granite and shadowed behind his hand, is
impossible for the barkeep to read. Returning once again to the realm
of motion, he chuckles briefly, in a despondent fashion, and
sighs. For the first time, the motions of his breathing edge into the
realm of mundane perception. He rallies his voice one last time, too
tired for any emotion born of anger to dominate his speech.

"Now, I bet you think this is all just a facade. I bet you think it's
revenge, or anger, or some such that motivates me: that I rage inside
with a desire to kill the Luddites who assaulted us, who drove us
down, who forced me to crash- land, whose actions resulted in
Lauktk's" he falters ever so slightly, "death. I felt that briefly
then, but I feel almost nothing about them now - they aren't important
enough to warrant personal hatred. Pirates, the ISO, the Luddites,
even the Aera - they're all just dancing to the blood rhythms that
cause every cell in their bodies to join in a choral chant of 'Stay
alive! Stay alive!' I can't really blame them for it, even if they'll
probably blame me if I take the lead in the dance and reciprocate
their violence. Pity really. Things would be so much more pleasant if
we could learn to not step on each others' toes, or claws as the case
may be." He turns his gaze up again, letting his fingers fall away
from his eyes, to rest once more on the bar. "Don't think I don't
mourn my friend. It's just that the blackness of an executioner's mask
doesn't make it fit for mourning clothes. If I see the Luddites that
killed him ... I'll probably try to kill them, but for the reason that
they'll be trying to kill me.  Here's my advice, Mr. Robo-barkeep:
don't hold grudges, don't look for comforting answers and don't wait
for magic wands. The first can only hold you back, the second are
never what you want them to be, and the third are always being held by
something that's going to turn you into a toad if you aren't
careful. Feel free to take it with a few grains of salt though. I'm
not anyone qualified to pontificate - me, I'm just someone" he glances
at the still untouched ale "not having a drink at your bar and
... flying a dead man's ship."

[The man finishes his water and leaves, each step clearly a source of
pain from injuries not yet entirely healed.]

\section{Little-Plot segments}
\subsection{The intro plot}

Lauktk's siblings (technically, they aren't all his full brothers and
sisters, as Klk'k families are a bit more complicated than the Western
notion of the nuclear family, but from the Klk'k perspective, they
might as well be) show up in Cephid 17.  Conversing with them about
Lauktk's never to be realized dreams of finding his own personal
truths of freedom in traversing space as his ancestors traversed the
seas of Ktah, they decide the only fit course of action is to have a
prolonged, multi-system funeral commemoration (think "wake") that
takes the remains of Lauktk and his ship to some of the great
landmarks and capitals of friendly space before returning his ashes to
Deucalion's sister and scattering them over the oceans of Ktah.

Implementation issues:

To do this will require the purchasing of a jump drive and some
(perhaps a placeholder for the moment, or even just remove some cargo
space) passenger holding space for the siblings.  The idea will be to
push the player to get enough money to leave the system, and as soon
as they can leave the system, to not just wander aimlessly, but to
force them to see some of the unique places we're going to make. Also,
while on the trip, one should be pushed to take on some sets of fixer
missions that lie on the same path where there are moral/ethical
issues involved as to which of the fixers' missions one takes (thereby
defining somewhat the path the PC will later take, be it more
mercenary, more mercantile, a return to military service, a life of
piracy, or whatever), such that, by the time one returns to Ktah, one
must have enough money that one can buy a new ship (namely, if
possible to script, the original ship should be put aside in deference
to Lauktk), thus completing a rebirth both of the player's psyche,
and, through the change of ship, the rebirth of the player's avatar.

{\bf EVENTS IN ACT 1: Hiroshima (mon ami)}
\begin{itemize}
\item Deucalion engages in side quests for cheap jump drive (optional)
\item Deucalion and Lauktk's bond-siblings take part in wake that canvases friendly space
\item Character development (profession) of Deucalion via encountering certain fixers along the way and interacting with them

 Profession paths:
\begin{itemize}
\item		Officer (Different paths depending on which navy)
\item		Explorer
\item		Pirate
\item		Assassin
\item		Bounty Hunter
\item		Merchant
\item		Tycoon
\item		Fixer (multiple ethical subpaths)
\item		Spook (multiple ethical/political subpaths)
\item		{NEED MORE - ADDME}
\end{itemize}
\item Deucalion meets up with another former servicemate, Mirabel, at
the ceremony on Ktah.  Depending upon actions taken during this
chapter, he will either be told that he can't go back to the military,
entreated to come back to the (Andolian Protectorate military, or
questioned as to his interest in coming back.
\item Likewise dependent upon actions, Deucalion may receive a
communique from interested parties of questionable intent/methodology.
\end{itemize}

