\label{chapt:portfolios}
The following sections give a holistic overview of various key groups
as the exist or existed circa 3276 CE (i.e. the UtCS period, see
~\ref{timeline:UtCS} and~\ref{exptimeline:UtCS}), combining factional, species, and aesthetic
information in a single place, from an omniscient viewpoint. These
portfolios are primarily designed to give insight and direction to
artists and other content creators. Much of the information here may
be duplicated elsewhere. There may be cases where the omniscient
viewpoint and in-game viewpoint disagree. Such disagreements are
likely intentional, reflecting misconceptions or ignorance on the
part of the extant groups. In cases where the disagreements seem more
ambiguous or unlikely in intent, please feel free to direct questions
as necessary.

\section{Cross-Portfolio Information}
\subsection{Vessel Overview}
\begin{itemize}
\item Engines

Retro thrusters, retro thrusters, retro thrusters, retro thrusters!
(to the tune of ``Developers, developers, developers!'')

If your craft is: A) not Rlaan manufactured and B) more than 5 meters
long it probably needs retro thrusters.

Now, the retro thrusters and maneuvering thrusters don't need to be
necessarily as pronounced as the rear engines (giving you some
artistic license in defining a visual front/back) but they NEED TO
HAVE VISUAL INDICATORS OF THEIR EXISTENCE. For some ships this is
mainly a texturing question. For larger vessels, this will necessarily
influence your model design.  

\item Radiators

One of the biggest problems in space is heat dissipation. VS
spacecraft have a lot of heat to get rid of, so they're going to need
radiators. While, as with retro and maneuvering engines, artistic
license is granted in how much of the ship's surface area is going to
be radiator dominated, the same need for visual indicators of
radiators exists. If you have fins, you should probably put a radiator
texture on them, because they're almost certainly not for aerodynamic
flight.  

\item Internals

VS ships aren't, in general, intended to be very dense. This is more
true for larger craft than strike craft. In general, the idea for VS
ships is that the surface-area to internals ratio should be pretty
high. The bigger ships in VS can be seen as big shells of armor and
radiators around many, smaller, internal components. Moreover, a lot
of the internal components would actually be plumbing, of the coolant
(I always had a soft spot for gallium alloys, but no decisions are
finalized on what exactly the coolant is) circulating variety, running
between the reactors/engines (the engines being fusion reactors
themselves) and other heat-generating components (e.g. weapons). The
rest of the space being taken up by insulation (vacuum is pretty good)
and physical shielding to keep the crewed and other sensitive parts of
vessel safe from the engines/reactors, and other hazardous portions of
the ship.  

\item Atmospheric flight

Most spaceships in the VS universe, even if capable of accelerating to
escape velocity, are not designed for atmospheric flight. Some are. If
its role description mentions ``Aerospace'' or ``lander'', ``puddle
jumper'', ``dropship'', ``ground support'', or ``orbital'' then it's a good
bet it has an atmosphere-friendly design. If not - build a spaceship,
not an airplane.

That said, some smaller craft that are not atmosphere-friendly will
still be atmosphere capable. The more developed the infrastructure of
the group using/designing the ship, the more likely they don't intend
for it to ever see atmospheric use on an inhabited planet - this is
what docking stations, space-elevators, orbital shuttles, and other
infrastructure exists for. In fact, (once we get planetary structures
squared away a bit better), the really industrialized planets will
probably start firing on you if you attempt to land in anything larger
than a pinnace/dispatch craft, and even then only with appropriate
permission. On the flip side, if it's a Forsaken ship, or a Luddite
craft, then it probably spends a lot more time visiting places without
well-developed orbital infrastructure and may need to land on
occasion. Also take into account the age of the design you're
considering. Much of the VS universe as of 3276 is post-frontier, but
a lot of places don't have to think back too far to remember when they
weren't. Use your better judgment.

If the ship is very large, even if it is atmosphere-capable, it is
almost certainly {\em not atmosphere friendly}, unless perhaps it's
something very esoteric (atmosphere skimmer for terraforming project
or some such). In general, if the ship is very large, it's best to
assume it's not even atmosphere capable unless there's a very good
reason for it. Large vessels will make use of shuttles and dispatch
craft to take care of business on the ground.

\end{itemize}


\input{portfolio_Aeran}
\input{portfolio_Andolian}
\input{portfolio_Andolian}
\input{portfolio_AndolianProtectorate}
\input{portfolio_Bzbr}
\input{portfolio_Confederation}     
\input{portfolio_CriminalElements}
\input{portfolio_Dgn}             
\input{portfolio_Forsaken}             
\input{portfolio_Highborn}     
\input{portfolio_HomelandSecurity}
\input{portfolio_HuntersGuild}
\input{portfolio_ISMG}                  
\input{portfolio_ISO}               
\input{portfolio_Klk'k}                 
\input{portfolio_LIHW}                 
\input{portfolio_Lmpl}
\input{portfolio_Luddites}
\input{portfolio_Mechanist} 
\input{portfolio_Mishtali}
\input{portfolio_Nuhln}
\input{portfolio_Purist} 
\input{portfolio_Purth}
\input{portfolio_Rlaan}
\input{portfolio_RlaanBriin}
\input{portfolio_Saahasayaay}
\input{portfolio_Shaper}
\input{portfolio_Shmrn}
\input{portfolio_Uln}
\input{portfolio_Unadorned}
%input{profile_}

% LocalWords:  UtCS
