\label{chapt:timelines}
\section{Timeline outlines}

This section contains two timeline outlines, the first, in
Section~\ref{Abbrevtimeline} is abbreviated, and only focuses on the
``playable'' portion of the VS timeline. The second, in
Section~\ref{TimelineOutline} frames all the events in the Vega Strike
universe from the origins of the TWHON in the long distant past
through the decimation of the jump networks that marks the future
``Rimward of Eden'' setting.


\subsection{Abbreviated game-oriented timeline}
\label{Abbrevtimeline}
The following abbreviated timeline shows points in time in the VS
universe that seem like good matches for interactive content, and
highlights some of the key occurances that would occur/would be
occurring during potential gameplay. The Vega Strike engine would be
appropriate for some, but not all, of the points listed. Major point
titles are taken from fragments of the following line:

{\Large \centerline{"Icarus descended upon the coldest sea, foundering inheritance rimward of Eden."}}
list of titles and highlights

(titles in chronological order)

\begin{itemize}
\item	Icarus Descended: ( circa 2684 CE )
\begin{itemize}
\item	[-] Pre-SPEC Expansion in the Diamond Dust
\item	[-] Founding of the major human powers
\item	[-] Origins of the Merchant's Guild
\item	[-] Devolvement of fallen colonies into piracy, etc.
\end{itemize}
\item	The Stars Are Also Pyres:
\begin{itemize}
\item	[-] Fraternal War
\item	[-] Origins of the Confederation of Inhabited Worlds
\end{itemize}
\item	Upon the Coldest Sea: ( 3276 CE)
\begin{itemize}
\item	[-] Aeran War
\item	[-] Maturation of "Grandchildren"
\item	[-] Andolian powershifts
\end{itemize}
\item	Black Paralysis:
\begin{itemize}
\item	[-] Exploration/exploitation of Jump Network
\item	[-] FTL maturation, discovery of "locked" gates
\item	[-] Changes in Human-Rlaan relations
\item	[-] Final death throws of Aeran Ascendency
\item	[-] Origins of "marked" Aeran co-operators
\end{itemize}
\item	Foundering Inheritence: ( circa 3663 CE )
\begin{itemize}
\item	[-] Further exploration of locked gates
\item	[-] Revelations on the Ancients' libraries
\item	[-] Opening of gates
\item	[-] Launching of the Fleet of 10,000
\item	[-] Invasion of Rlaan space from beyond locked gate	
\item	[-] Decline of Rlaan assembly
\item	[-] Near extinction of Saahasayaay
\item	[-] Devestation of Rlaan population
\item	[-] Evidence of TWHON stirring
\item	[-] Revelations on the intended role of the Nano-Plague
\end{itemize}
\item	The 10,000:
\begin{itemize}
\item	[-] Travels and struggles of the stranded fleet
\item	[-] Origins of the Pen-Ul and Pen-Pen
\end{itemize}
\item	Rimward of Eden: ( circa 11007 CE )
\begin{itemize}
\item [-] A young, marked Aeran struggles to claim his people's
birthright and escape from human imposed bondage on a world never
truly recovered from an interstellar cataclysm, the cause of which has
been lost in time and strife to the inhabitants of his world. Old
relics lead him to seek the help of the "demon" Kilroy, who is
actually one of the original batch of Grandchildren.
\end{itemize}
\end{itemize}

\subsection{Vega Strike Universe Timeline}
\label{TimelineOutline}
The following is a timeline, in outline form, of the currently
relevant timespan of the Vega Strike universe, spanning from tens of
millions of years BCE to tens of thousands of years CE. The events
mentioned in the timeline will be examined in greater detail in the
sections following it.

\begin{itemize}
\item TWHON Era: --irrelevant-- to 80 Million BCE

pre-TWHON develop technology allowing creation of jump network and
begin work thereon

pre-TWHON subjugate/destroy/incorporate all other competitors in
galaxy

pre-TWHON become increasingly post-technological (no longer building
tools)

proto-TWHON abandon bodies, transcend tool/body distinction entirely

proto-TWHON lead move toward unification of entire TWHON species

proto-TWHON emerge victorious after defeating other pre-TWHON factions

singular TWHON mind coalesces

TWHON spreads back out take advantage of galactic scale material and
energy resources

thought-localizations within the TWHON bodies lead to heterogenous
faceting of the singular TWHON mind

various aspects of the TWHON mind begin designing sapients for a wide
and varied array of reasons

\item Ancient Era 80 Million BCE to 2 Million BCE
\begin{itemize}
\item	Several Million Years BCE 

Ancients A, B expand throughout this region of the galaxy, interact
with at least 4-6 other groups of similar developmental level

\item 2 Million Years BCE 

Interspecies warfare among Ancients at the behest of TWHON

Advancement of subset of Ancient species to early post-technological
development stages

Indecision by TWHON as to whether to nurture, co-opt, or terminate the
more advanced Ancient cultures leads to internal unrest in the TWHON
mind

Internal discord in the TWHON mind manifests as conflicts between
physical TWHON bodies

TWHON aspects sympathetic to Ancients seek to use several remaining
advanced Ancient species to assist in forcibly convincing the TWHON
mind of the correctness of the pro-Ancient views

TWHON internal conflinct and TWHON-Ancient conflicts come to a head in
battle near research planet on current Uln homeworld. Attempts at
using Ancient-implemented devices for forcible convincing fail
catastrophically, causing insanity and rapid devolvement of the TWHON
mind.

Increasingly mindless TWHON bodies continue to act out echoes of the
last few semi-coherent thoughts of the TWHON mind, seeking out and
destroying all Ancient species in increasingly self-destructive acts
of violence.

Slag worlds and planetary debris fields formed as TWHON bodies
alternate between vaporising various Ancient worlds from orbit,
hurling themselves into said worlds, and engaging in other equally
inefficient acts of mayhem and self-destruction.

After some thousands of years of turmoil, and the complete destruction
of all Ancient species, the remaining TWHON bodies begin to migrate
coreward and go dormant.

\end{itemize}
\item Inheritor Era  2 Million BCE to present (3276 CE at beginning of "Upon the Coldest Sea")
\begin{itemize}
\item	 960,000 to  830,000 BCE 

Alphan/Betan sublight expansion 

Conflict between Alphan and Betan species leads to development of FTL technology 

Both Alphan and Betan species rendered extinct due to conflict with Nano-Plague 
\item	 100,000 BCE 

Rlaan Workers and Defenders speciate 
\item	 13,000-11,000 BCE 

Humans develop agriculture 
\item	9783 BCE 

Rlaan writing system formalized across entire planet 

Rlaan Assembly formed 

\item "Icarus Age" begins (11th Century BCE)
\begin{itemize}
\item	11th Century BCE 

Rlaan begin sublight expansion 
\item	1st Century BCE 

Aera begin clearing forests to create agricultural land 
\item	3rd Century CE 

Rlaan research into FTL re-activates Nano-Plague 

Rlaan expansion halted as Rlaan recover from Nano-Plague 

\item	5th Century CE 

Uln reach Sul-Gatwa High Castle 

Gatwa dynasty cements Ulnish supremacy 

\item	15th Century CE 

Rlaan finish consolidating post Nano-Plague recovery, reintegrating
surviving colony worlds back into the Rlaan Assembly

Rlaan encounter the Saahasayaay homeworld 

New Rlaan Assembly expansion policies instituted, including stringent requirements for core-system development before expansion 

\item	18th Century CE 

Uln begin sublight expansion 

Negative Rlaan experiences with Saahasayaay lead them to concentrate
on the construction of designer species from pre-sapients, including
the Lmpl and Nuhln, rather than relying on encountering useful
sapients.

\item	23rd Century CE 

Death-throws of the nation-states on Earth 

Increasingly cemented meme-groups 

Humans begin sublight expansion 

Origins of the Forsaken 

\item	25th Century CE 

Second generation colonies launched (Human space) 

Massive construction projects common (Human space) 

Cherryh station 

Hephaestus 

FIXME MORE EXAMPLES TO COME LATER 
\item	26th Century CE 

Humans begin forays into jump technology, in particular, the Unadorned
and Andolians

\item	27th-28th Century CE 

Nano-Plague activates in Human space 

Collapse of nanite-based technologies in Human space 

Human societies stabilize, having adapted or failed. 

"Interstellar Church of True Form's Return" founded 

\item "Icarus Age" ENDS 
\end{itemize}
\item "Diamond Dust Age" begins (28th Century CE) 
\begin{itemize}
\item	"Reconstruction period" begins 

Beginnings of human interstellar economy. 

Initial formation of Merchant Cartels 

Displacement of Forsaken by FTL claim-jumpers begins. 

\item	29th Century CE 

SPEC drive invented by Emilio Sofono 

Displacement of Forsaken accelerates. 

"Reconstruction period" ends 

\item	"Exploration period" begins 
\item	30th Century CE 

First Contact (for humans): Unadorned discover Mishtali 

Joint Shaper/Light-Bearer venture discovers ancestors of Dgn and Shmrn
in system dubbed "Beckett's Murky Venture".

Light-Bearers discover Ktah 

Light-Bearers occupy Ktah 

Andolians discover Ktah 

First human interstellar conflict (later dubbed "The Fraternal War")
begins

Hoshino uprising, Light-Bearer presence on Ktah neutralized, nuclear
retaliation kills over 1 billion Klk'k.

Andolians liberate Ktah system

"Ktah restoration project" begins 

Andolians reveal Light-Bearers' Space-Born project, muting
Light-Bearer support from other factions

Widespread ramp up of space-based military assets begins (Human space)

Andolians systematically debilitate Light-Bearer faction, liberating
Space-Born and Shmrn. Andolians now responsible for more human deaths
than any other group in human history.

\item	30th-31st Century CE 

Surviving Light-Bearers turned over to Klk'k, Space-Born, and Shmrn
custody.

Light-Bearer meme-group rendered defunct 

Andolians cede control of several former Light-Bearer colonies to
other factions, primarily Shapers as compensation for lost
investements.

Independent Shmrn colonization effort begins. 

"Andolian Protectorate" established as entity in control of Andolian
and Klk'k affairs

Diplomatic talks concerning the formation of "Confederation of
Inhabited Worlds" begin

League of Independent Human Worlds (LIHW) formed

\item "Exploration period" ends 
\item "Diamond Dust Age" ENDS 
\end{itemize}
\item "Confederation Age" begins (31st Century CE)
\begin{itemize}
\item	"Paranoid Avarice period" begins (Human space) 

Independent human meme-groups aggressively absorbed by larger
meme-groups.

Andolians discover Purth, begin redesigning them

Aera encounter Bzbr, become increasingly worried about potentially
abundant sentient species

\item	"Confederation of Inhabited Worlds" formed 

Internal conflicts abate. Military buildups continue due to increased
assessments of risk from potential alien sapients.

\item	(MORE PRECISE TIME-GOES-HERE) 

Rlaan encounter Uln 
\item	(MORE PRECISE TIME-GOES-HERE) 

Humans (specifically Shapers) encounter Rlaan 
\item	(MORE PRECISE TIME-GOES-HERE) 

Aera encounter Uln 
\item	(MORE PRECISE TIME-GOES-HERE) 

Aera/Humans/Rlaan/Uln expand to each others' borders 
\item	(MORE PRECISE TIME-GOES-HERE) 

Aera colony passage negotiations fail 
\item	(MORE PRECISE TIME-GOES-HERE) 

Aera attempt to sneak convoy through Rlaan space 

Aeran convoy destroys Rlaan civilian vessels 
\item	(MORE PRECISE TIME-GOES-HERE) 

Aera-Rlaan war 
\item	(MORE PRECISE TIME-GOES-HERE) 

Aera-Rlaan cease-fire 
\item	3276 CE 

Aera invade Forsaken space 

Confederation grudgingly mobilizes to support Forsaken 

Aera/Confederation enter into a state of war 

\item Upon the Coldest Sea events go here
\phantomsection
\label{timeline:UtCS}
\item Black Paralysis events go here
\item Events between Black Paralysis and Foundering Inheritance go here
\end{itemize}
\item Fleet of 10,000 launch goes here
\item Foundering inheritance goes here
\item Fleet of 10,000 events go here
\end{itemize}
\end{itemize}


The following sections provide details for key historical events. The
sections have been placed in roughly chronological order, but may
overlap somewhat in time.

\section{TWHON Era}
The species that would become the TWHON (for simplicity, let's call 
them pre-TWHON) were not by any stretch of the imagination the first 
lifeforms, nor were they the first sentients, nor even the first 
species to leave their own worlds. They were not even, by many 
Nietzsche-esque metrics, the best of their contemporaries among the first 
crop of spacefarers, being not particularly remarkable beings of not 
particularly remarkable chemistry and rocky-body planetary 
origins. What the pre-TWHON were, however, is driven, and they 
survived, and they prospered, and they expanded - sometimes with their 
neighbors, and sometimes at their expense. They boiled out from their 
homeworld, rolling out in great waves of ships limited to no more than 
a few hundred C, filling the stars with trillions upon trillions of 
pre-TWHON. Advancing in technology and power, they eventually did 
achieve a 'first' status, becoming the first group in the galaxy to 
discern how to re-carve the topology of space to connect pairs of star 
systems together. They constructed the beginnings of the jump network, 
and leveraged this to further accelerate their expansion. Eventually, 
over hundreds of thousands of years of exploration, conquest, 
subjugation, assimilation and extermination, they achieved their most 
important 'first': first species to bring the entire galaxy under 
their rule. 

However, even with the speed of communication and travel offered by 
the increasingly inter-connected jump network, time and distance took 
their toll on the pre-TWHON cultures - for there were many cultures 
now, not only one. With their common goal of dominance achieved, there 
was progressively less to hold the pre-TWHON together. With few 
constraints remaining, and no clear goals, the pre-TWHON erupted into 
a decadent exploration of the possible. Cultures fragmented, groups 
willingly speciated, and all manner of experiments were performed on 
often grandiose scales. Many groups gave up on embodiment, preferring 
to live within the data streams, and others transitioned to amorphous 
colonies of pico- and femto-tech manipulators housed within 
space-going shells containing the more complicated mechanisms of power 
generation, propulsion, and thought. The latter groups frequently 
joined together to form ever larger ecosystems of body and thought. 

One such group of the re-embodied (let us call them the proto-TWHON) 
yearned for a return to the unity of purpose that once had propelled 
the pre-TWHON across the entire expanse of the milky way. They had 
come to the conclusion that no species, nor even culture, could stand 
the test of time without coming undone. They believed fervently that 
such unfragmented immortality could only be achieved by an individual, 
and set out to become only one in number. A return to unity was a 
common yearning among the pre-TWHON, and they found many willing to 
join them in the promulgation of their cause. Gradually, the 
undirected fountain of exploration following the great achievement 
died, calcifying as the strongest few sentiments coalesced and 
overshadowed all other contemplation. The splits among the pre-TWHON 
had become too deep, and they were so divided that war once again 
played out across the galaxy. 

In the many millennia that followed, the proto-TWHON would emerge 
victorious, assimilating the willing and exterminating those who 
defied unity. When all that had been pre-TWHON had been at last 
reconciled, the entirety of the proto-TWHON massed in one place to 
begin a period of nomadic existence, a single great ecosystem composed 
of an intricate and complex menagerie of spacecraft bodies in changing 
hierarchies and purposes, moving from star to star, grazing on the 
resources of system after system while the minds that shared the 
network of bodies coalesced and reduced in number. Emerging from this 
long meditative slumber would be a being of only one mind, and one 
body, albeit a body in innumerable parts. 

This being was not yet, however, the TWHON as the Ancients would come
to know them, for it had, for a time, truly achieved unity. By the
time of the creation of the Ancients, the being had become internally
heterogeneous. The singular being had spread itself across the galaxy,
repairing, replacing, upgrading and expanding the jump network and
greatly enhancing the communication networks tying stars
together. This time apart, however, introduced a slight drift. And
from the drift, came a fluidity previously held at bay by rigid
homogeneity. Localizations in the thoughts of the singular being began
to form. Particular clusters of bodies became increasingly responsible
for particular modes of thought. As this trend progressed, and then,
eventually slowed again in a new equilibrium, the singular being had
become the TWHON: a collection of personified aspects of a shared
mind, distributed across body elements spread over the entire jump
network, with all the communication hurdles such a galactic scope
entailed. It was in this environment that the different TWHON aspects
began their own independent projects, some of which would eventually
give rise to the various species of the Ancient era.

\section{Ancient Era}
There were, as you assumed, more than two species of Ancients - in
fact, I'm fairly sure it is mentioned somewhere (although I can't
recall where, and if I can't, then... ) that there is evidence of
interactions with at least six other groups of Ancients. The two in
question were just the locals, and their domain was demarcated by the
locked gates, an imposition of the TWHON. The TWHON, whether the
Ancient group in question arose more or less naturally to a meaningful
existence, or was created entirely from scratch by TWHON whimsy, were
the creators of all of the Ancient groups in that the molded them and
shaped them to their own designs and for their own reasons. Had their
era not passed, there would have come a time when, whether directly or
by Ancient proxy, the TWHON would almost certainly have manipulated
Humans, Rlaan, and Aera alike for whatever motivation struck them at
the time - and, if by odd chance we had remained sufficiently
insignificant to escape the notice of the TWHON, the local Ancients
would almost certainly have had their own games to play, more likely
than not at our eventual expense. There were, to distill the essence
of relevant groups, only the Ancients and the TWHON, as no other
groups were given license to develop.

The downfall of both groups was to come from a great indecision within
the mind of the TWHON. As the Ancients progressed into an early
post-technological stage of development, certain aspects of the TWHON
mind no longer saw it proper to use them as playthings and
experimental subjects (both of the former as groups, not individuals -
the TWHON didn't do "alien abductions" they just started massive
interstellar wars, conducted xenocides, created plagues, monsters, new
Ancients, challenges, rewards both equitable and not...  standard
god-like-being stuff). To keep things brief, and use metaphor and
analogy (though it leads to an incomplete description), one can
picture this all culminating in an eventual Ancient revolt, assisted
by elements of the TWHON mind that, rather than being squashed,
instead set off a "civil war" within the TWHON mind, spilling over
into the TWHON bodies. The defensive assistance rendered by the
pro-Ancient TWHON lines of thought against their urges toward
continued usage of the Ancients as inconsequential entities, would
eventually become very well known to the extant species (and cause
many later species to cease to be extant) as the "nano-plague." A
majority of the Ancient groups joined the revolt, though some
continued to fight by proxy for their masters.

One decisive battle would prove to be the beginning of the end for the
era of TWHON and Ancient alike. On one of the two research planets in
what is now the Uln home system, indeed, on the planet that is now a
noticeable debris field in the Uln home system, research was
frantically being conducted on how to reinforce the allied lines of
TWHON thought at the expense of antagonistic patterns withing the
TWHON mind. The future Uln homeworld was evacuated and the proto-Uln
abandoned. Fleets of moon-sized Ancient craft, a fragment of one such
eventually becoming the Sul-Gatwa high-castle, and numerous TWHON
bodies gathered in the system to defend and participate in the ongoing
research. An armada of TWHON bodies attacked the system, and prototype
device were activated. The battle unleashed enormous devastation upon
both fleets, but the TWHON-mind-altering devices were to prove far
more lasting in the devastation they were to unleash. In addition to
cognitive dissonance increasing within the TWHON mind from the
operations of the mind-altering devices, the battle was sufficiently
fierce that TWHON ideas, and thus Names, were dying along with TWHON
bodies. These deaths echoed, and then reverberated, and then fed back
and amplified on the devices, burning themselves into, and by turn
burning out the TWHON mind. As the battle proper neared physical
resolution, the forces allied on the side of the Ancients' liberation
looking to have gotten the better of the engagement, though the
victory would likely have already been Pyrrhic in the longer term, the
TWHON mind sank below the threshold of sanity, all Names becoming
muddled and losing sufficiency of form (the TWHON mind already having
been many times weakened by prior TWHON-TWHON engagements). What
moments ago were their supporting deities now turned on the Ancients
in the TWHON slide from supremacy and, along with the remains of the
invading TWHON bodies, themselves equally deranged, slaughtered every
Ancient in the system and destroyed, too late, the planet on which
their downfall was birthed. The Uln homeworld was spared a more
sterile fate only for the hurry with which what remained of the TWHON
mind wished to distance itself from a place of suffering.

What had before been a war, if a galaxy-devastating one, became a
cataclysmic spectacle of heretofore unrivaled savagery, devoid of
rhyme or reason. As the TWHON mind degenerated, the Names of Rage,
Confusion, and Suffering fared better than many, and were allowed to
greatly influence the actions now being undertaken by the TWHON bodies
throughout the galaxy as they began to purge every system of Ancients
of every kind, revolutionaries and loyalists both. It was in this
period, not during the war proper, that the slag worlds were created;
the great overkill measures of boiled atmospheres, muted suns, melted
crusts, and shattered worlds left to re-coalesce into the not
infrequent debris clouds seen in UTCS era systems were not the actions
taken by victorious armies upon their foes, but by a great singular
mad mind, fragmented and devolving, lashing out in sometimes
self-annihilating fashion (see: worlds rendered uninhabitable by what
would appear to be prolonged bombardment by large impactors,
i.e. TWHON bodies) at anything that could be detected by its
increasingly twisted perception. Every living Ancient in every system
on the jump network was exterminated. Every gathering of Ancients in
systems off of the jump network was obliterated. What was once TWHON
hunted the Ancients, and in finding and annihilating them, found the
occasional glimmering shred of purpose to its new existence. All that
remained of the Ancients were arks and tombs, drifting alone in the
relative safety of the void between stars. When it could no longer
find anything to hunt, the TWHON bodies wandered, some disassociating,
some taking to slumber, some few wandering beyond the edge of our
galaxy, to slowly plough the deepest of voids toward ambiguous end,
and a great many retreated to the warm embrace of the old and dying
stars far coreward of our portion of this galaxy.

\section{Alphan-Betan period}
As for the Alphans and Betans, as to whether they were, like the Uln,
beings existing primarily because of Ancient intervention, or
creatures independently evolving into sentience now that the Ancients
and TWHON no longer policed species developments -- I didn't really
have any particular feelings on that, so that could go either way for
whoever wants to flesh their origins out further.

As for their demise, however, the "conflict with the nano-plague" is
exactly that. They attempted to combat it with military force, and it
wiped them out wholesale in ever-escalating fashion proportionate to
their increasingly desperate attempts to eradicate it.

The Alphans (so named as the first race in the post-Ancient era to
leave any evidence of interstellar travel) were quite technologically
advanced, but appear to have simply missed or ignored avenues of
research involving FTL, shields, and related technologies. They
expanded outward in relativistic vessels over tens of thousands of
years, eliminating or curtailing any potential competition without
incident until encountering a colony world of the Betans. Unlike the
stately march outward of the Alphans, the Betans had expanded in more
chaotic fashion, constantly fighting amongst themselves, even in the
interminable wars that sublight travel forces. The initial encounter
with the Alphans was nearly instantly hostile, and the Betans (as
evidenced by a sharp temporal dividing line in the nature of the
artifacts left behind) rapidly reverse-engineered much of the
encountered Alphan technology and began to counter-expand into Alphan
space. The light-speed trickle of information and sublight flow of
craft and personnel meant that the wars (for there were no true
star-empires at sublight - just collections of vaguely associated
colonies) raged on for thousands of years before FTL was discovered by
the Betans.

With the vast tech and industrial bases under their command, FTL tech
spread quickly enough that numerous vessels were completed before the
nano-plague had sufficiently reawakened to begin to attack the Betan
infrastructure. The extermination fleets that were launched struck
Alphan world after Alphan world without mercy or warning, first those
on the jump network, and then, methodically, bearing down on worlds
outside it. The Alphans were rapidly destroyed as a power. However,
they would not have died out completely save for the nano-plague
rearing itself in the wake of the Betan attacks. Believing the
nano-plague to be but yet another Betan weapon, most Alphan worlds
threw what remained of their military at the problem and were utterly
snuffed out in quick succession. The Betans similarly believed the
nano-plague to be an insidious Alphan response to their FTL
attacks. Their military being more intact than that of the Alphans
(though crippled by the loss of nanites), their response was greater,
and thus was their demise more profoundly enacted at the hands of the
nano-plague (see page 57 - the system (incorrectly) assumed to be the
origin of the Saahasayaay creeper organism was in fact a Betan base so
heavily damaged that it was not recognizable as Betan in origin (the
chlorine atmosphere also helped obscure things) ).

While the last remnants of the Alphans withered away in scattered
pockets of decay, their nanite-reliant technologies worthless, and
their worlds and colonies rendered inhospitable by Betan weapons of
mass destruction, the Betans, as was their nature, fought to the very
end, and thus became utterly extinguished, to the last Betan, far
sooner than the Alphans. Those Betan sublight craft still in transit
would all fall prey to their own agressions against the nano-plague
within weeks of reaching their destination systems. Their Alphan
counterparts, on the other hand, would, technologically crippled,
having relied on nanites to reanimate the cryo-preserved colonists and
soldiers, in many cases limp on for years after their arrival. In the
more successful cases, the handful of Alphan crew awake during the
sublight voyages founded downteched outposts and set up beacons,
calling out for rescue that would never come from brethren they did
not know were already gone. These last Alphan outposts are the richest
source of artifacts from that era. Without sufficient tech or
industrial base, and without a viable breeding population (usually 5
to 8 waking crew), even those lucky enough to find worlds capable to
some degree of hosting Alphan life pre-Alphaforming died off or
degenerated within a small number of generations.

Unlike Ancient technology, Alphan and Betan technology, while often
extremely advanced (in certain areas and deficient in others) is not
so advanced as to be completely indecipherable in function and/or
mechanism. Many Alphan and Betan languages have been studied and
heavily translated, in stark contrast to the Ancient languages, which
have only yielded to limited understanding, and even that only through
the use of other Ancient technologies on the Uln
homeworld. Unfortunately, as both groups' name for themselves
translates to "people" the names Alphan and Betan remain the common
descriptive terms. Alphan and Betan artifacts are heavily sought
after, especially those that are sufficiently intact that they can be
reactivated (as, in contrast to Ancient artifacts, the Alphan and
Betan ones are found on worlds that were merely devastated rather than
obliterated or assaulted until their surface was molten slag, Alphan
and Betan artifacts are sufficiently common as to be merely
outrageously expensive rather than nearly priceless).

\section{End of Alphan-Betan to beginning of `Icarus Descended'}
\subsection{On the origin of The Forsaken}

The Forsaken are, at top level, sub-light settlers abandoned and
displaced due to the FTL expansion of other entities. As a
non-coherent group, the Forsaken are much more diverse than any of the
meme-groups, with members derived from an assortment of late embarking
colonies, far ranging colony missions, and, this being a distinct
minority, arriving quite late with respect to other groups, refugees
from Sundered worlds. However, this assortment of influences actually
makes them even more distinct, because of the already diverse nature
of each of the subgroups that make up the Forsaken.

Some of the Forsaken actually set out long before most of the dominant
Human meme-groups, but in slower, more primitive vessels, or toward
significantly more distant stars. Earth, at the decline of the Nation
States and the beginnings of the nanite economic revolution, was a
place of significant turmoil. China, grown increasingly insular and
denied of information and innovation from outside as it rejected the
meme-trends, shielding its population from what it saw as memetic
viral infections through draconian information control, was the last
of the truly great Nation States. It sought to capitalize upon the
destablization of the other world powers and eventually plunged the
world into war by proxy support of the now shell-like governments of
numerous other former great powers. Much of Asia, in particular, was
ravaged, but the war ended far more swiftly than any had anticipated -
the first real signs of the economic change that the nanite revolution
that had been simmering only outside the Chinese borders had
wrought. Defeated planetside, the mankind's first manned interstellar
venture (following probes already launched decades prior) was
undertaken in haste, rather than jubilation. In a grand twist of
irony, the greatly distilled population of Chinese Nationalists and
assorted Nationalist allies that were to crew the two craft that left
the solar system embodied, more than their particular nations, the
meme of nationdom. Their haste, however, was to prove costly in the
long run. Their cryoships were efficient and simple, but extremely
slow. Neither of the vessels would actually be the first human craft
to reach its intended destination, being in fact usurped by later
sub-light settlers.

Soon after the war ended, there was one more last gasp of escaping
fragmentary Nation States, launching a final four more vessels, which,
though less hastily constructed, were all aimed toward more distant
stars rimward of Sol. Only one of the four craft would be the first
human ship to reach its destination, a destination which would turn
out to be a Sundered star system, unconnected to the jump
network. With the passing of this last great undertaking, the
nation-states were no more, as nanite economics led to profound
individual freedom of association for populations throughout the
entire solar system. However, even with the nanite economy, the
devastation wrought on its own and surrounding territories, combined
with the resource allocation that had been necessary to produce
interstellar craft in haste, served to cripple the further aspirations
of those in former China's territories both on Earth and in space.

Even as humanity lurched toward a post-sufficiency economy, there were
still those far better off than others. In an effort to leverage their
superiority in a world of increasingly homogenous meaning to one's
wealth, the precursors to the Highborn set about getting a jump on all
other meme-group's colonial aspirations - and the colonial aspirations
of meme-groups were many, seeking, much as those who came to be known
in America as the Pilgrims, a place away from everyone else as much as
anything glorious. Moreover, at sublight speeds, the nearby pickings
were few, and the choice planets were likely to fall to the
swift. Thus groups that had resources to spare rapidly saw themselves
embroiled in a spacerace. While the nanite economy was allowing
post-sufficiency living for Sol's booming population and the ever
advancing communication networks were allowing collaboration on
remarkable scales, interstellar undertakings were still monumental
undertakings, and groups without initial reserves, or with significant
catching up to do with respect to those immediately benefiting from
the new economy were in no shape to immediately mount interstellar
expeditions. Likewise, many groups were not interested in placing
leaving Sol among their goals.

Still, interplanetary developments in Sol were legion, and many
impressive interstellar craft began their construction in what were
distinctly two phases of expansion. There were the groups, such as the
precursors of the HighBorn and the Great Mormon Mission, which
expended their existing advantage to immediately begin their
interstellar expansion, and then there were groups, such as the then
newly formed Andolians, who would catapult themselves, through their
efforts in building the craft for the first wave of launches into a
position amongst the second wave of launches. (The Andolians, by
virtue of their extensive engineering undertakings for other groups,
actually made a number of unmanned launches during the first wave,
effectively securing themselves their future home system long before
they themselves would leave.) By the beginning of the third wave of
launches, however, the real-estate suspected of being most prime had
already been spoken for, and while some were willing to race the
long-already-launched Chinese, the general pace of launches slowed
down as immediacy lessened and those groups in position to expend
capital in exchange for travel had mostly already done so.

The third and longest phase of sub-light expansion from Sol was
therefore a sporadic process. Minor groups often had to, much to their
chagrin, band together in order to construct their interstellar
steeds, especially knowing that the worlds they would be headed to may
well be more marginal, leading to such mixed-meme enterprises as the
settling of Bantam. Eventually, the more prosperous or lucky colonies
would even begin sending out second-generation colonies, but the
number of those was sufficiently small as to not warrant separate
discussion.

It is from the later portion of the third phase of sub-light expansion
from Sol that the Forsaken would draw the bulk of their ranks. Those
who could not, or would not, leave earlier found themselves in flight
when FTL was discovered. Those headed to worlds that would turn out to
be on the jump network tended to find their planets already occupied
and their presence unwelcome at the end of their flight - this if
enough of them survived re-animation in the absence of their expected
nanites (should they have waited until contact to begin
re-animation). Others would arrive at their worlds, jump network or
otherwises, to find it not in the state their run-ahead terraformers
should have left it in, their terraforming nanites consumed by the
nanoplague. Still others in mixed fortune would find their destination
not on the jump network at all, and remain wholly isolated, having to
survive the unexpected nano-plague (though an admittedly slower onset
plague experience, according to all reports) without human aid or
intervention (except in rare cases, such as exemplified by the origins
of House Blythe, even if they came too late) until the invention of
the SPEC drive by Emilio Sofono.

Those that survived to see their erstwhile new worlds generally found
themselves turned away or worse. It is widely believed that there were
several massacres of late arriving sub-light craft that managed to
escape sufficiently detailed reporting to be able to assign
responsibility to any extant parties. The increasing frequency of such
sub-light craft turning up led to the formation of a dispossessed
settler's union, which, though ostensibly run by the displaced
themselves, was funded by the other meme-groups, and thus
strong-armed, until the later development of the Union's own
government (Forsaken being a colloquial rather than official term),
into adhering to a policy of expedient removal. Those lucky enough to
survive seeing their erstwhile destinations from orbit were forced to
stay on board their ships while crews outfitted them with jump drives,
and re-fuelled them (if possible), or crammed the colonists onto
converted freight ships when such was infeasible, sending the
dispossessed off toward what had begun, even then, to be called the
Rimward Badlands - a zone already looking from early exploration and
astrometrics to be an oddly less bountiful region of both space and
jump network that the major meme-groups were willing to lose
colonization of.

The final members of the Forsaken's population base would be the
smallest portion, those repatriated from Sundered worlds. Sharing only
poor luck and temporal dialation from the rest of humanity with the
Forsaken, most Sundered populations worth calling such were too large
to repatriate en mass and may not have been particularly compatible
matches with the Forsaken in the first place, but a number of them
supplied either a one time disbursement or a steady drip of persons
fleeing their black paralysis off of the jump network for the only
group willing to take them in.
 
\section{Icarus Descended}
\subsection{On various Pirate Groups and their origins }

The history of many of the pirates seen in the UTCS era is closely
tied to the history of the Sundered and the fates of numerous fallen
and less prosperous civilizations. Roughly classed, there are three
tiers of groups generally deemed pirates and one that, though rarely
called as such, are: `organized' `criminal enterprises' (as is common
knowledge, organized crime is generally only somewhat organized and
somewhat about crime) such as House Blythe and the Tribe of Eliana,
the economically downtrodden seeking to leech from the more wealthy or
willing to be used as pawns or intermediaries by groups looking for
some level of deniability such as may be found on various LIHW and
Forsaken worlds, and finally, the odd Neo-Barbarians who've been
gifted or sold spacecraft in generally less than above-board
transactions.  Finally, there are the kleptocracies, degenerate
governments where corruption, bribery and extortion are so the norm as
to find the committing of what amounts to highway robbery by the local
police or military forces to be commonplace - while the actions of the
latter are sometimes acts of piracy, they would assuredly bristle at
being called pirates.

The first grouping are both the least and most dangerous, as they are
uninterested in random acts of violence - just economic gain - and
their visible position requires some modicum of discretion. They are
arms dealers, loan sharks, extorionists, traffickers in forbidden
goods, and not above killing the odd fellow, but only if he's proven
bad for business.  They are, most importantly, reasonable in that they
can be dealt with via credits more often than guns.

The second class number the most numerous of pirates, those, as
persons or as groups, seeking to profit from theft, destruction, or
illegal activities, the lattermost often as subsidiaries to a member
of the first tier. They are local phenomena, cropping up in places of
poverty or insufficient oversight. They are, in effect, the most
well-to-do of some system or systems' gangs, rich enough to outfit
themselves with spacecraft in the first place, insufficiently wealthy
or powerful to become organized competitors to the first tier, and
existing in systems welcoming, tolerant, or impotent to act concerning
their presence. Especially in Forsaken space, such groups rarely
attack local vessels, and are sometimes more akin to local militias
"enforcing tolls" than the true pirate groups operating out of LIHW
worlds (the major powers being rather better at scrubbing bases of
operation out of their own systems). These should not be confused,
however, with the paramilitary groups operating in Forsaken space, for
though the line is blurry, there is a distinct non-locality and size
to the paramilitary groups that is lacking in amongst the local armed
rabble. Unfortunately, pirate groups rarely confine themselves to
their systems of origin. Thus, problems with local sources become
pandemic scourges. Fortunately, those groups with an eye toward
continued existence tend to extract cargo rather than lives or vessels
- attracting too much of the wrong attention can lead to Confederation
crackdowns, or worse.

Finally, there are the (so-called) Neo-Barbarian groups, much more
rare than the second tier, and actually sharing a more similar initial
origin with the first tier, though their origin as Sundered colonies
tends to limit their range to areas much closer to Sol sector. To
understand the existence of the NeoBarbs, one must recall that only a
small fraction of all systems are connected via the jump network. A
fair number of colonies settled and colony missions in-flight at the
time of FTL's discovery and the ensuing nano-plague were of systems
not on the jump network. This did not spare them from the nano-plague,
but did greatly hinder their recovery, as SPEC travel was not invented
for several hundred years after the development of the jump drive and
has remained inconvenient (albeit decreasingly so) for interstellar
travel through into the UTCS era.  Many of these colonies never
recovered. A fair fraction of them were completely lost, including
such formidable notables as the Great Mormon Mission. Some, in a dark
sense lucky, produced so few survivors by the time exploratory SPEC
craft arrived that rescue operations were affordable, and resettlement
often occurred in Forsaken space. Those living on more innately human
habitable worlds met a variety of fates, each living through the
nano-plague in their own ways. Of tangential, but important, note is
the fate of some worlds that bear evidence the colonists attempted to
fight the nano-plague directly: Their civilizations were razed to the
ground when it fought back, although exactly how remains
unclear. Returning the point at hand, however, of those that survived
the nano-plague, many did not survive as anything resembling their
initial civilizations.

It is, of course, a bit of a misnomer to call them Neo-Barbarians, as
they are not so much barabaric or savage as merely culturally and
economically divergent from the rest of humanity (far more so than
even the Forsaken). Of course, taking the original meaning of
barbarian, it is perhaps an accurate moniker. Those who continued on
in their existence outside of the jump network were not privy to
centuries of advancements or cross-cultural pollination. This made
many of these groups of particular interest for Luddite recruiters,
seeking both isolated bases of operation and untainted samples of
humanity. Thus, through Luddite and the odd profiteer, robber baron,
or fool, did the isolated and technologically backwards Sundered (as
those off the jump network had come to be called) gain access to more
modern craft and resources. This gave rise, among those civilizations
that had become more aggressive during their isolated survival
struggle, to the use of their newfound toys to prey upon their
neighbors for anything they could not produce or otherwise
acquire. Neo-Barbs are the most unpleasant of pirate encounters
because they are generally more interested in acquiring one's ship
than one's cargo.

It is interesting to note that, alone among the human groups, those
more prosperous among the Sundered continue to colonize systems off
the jump network. However, due to the impracticality of scaling SPEC
travel to trading routes beyond astronomically adjacent systems, the
overhead of even that, and the impossibility of swift communication,
there are no empires as such, with each system or at most cluster of
systems being given to its own governance, and development is greatly
slowed due to the trickle of interaction with other groups. The
particular interest the Luddites have placed in a subset of the
Sundered has led to significant tensions between those Sundered groups
receptive to the Luddites and all other Sundered polities.


Specific "Pirate" groups:
\begin{itemize}
\item House Blythe:

House Blythe's origins lie with the Sundered colony of Sheltersky,
which happened to be very close, only a couple of light-years, from a
far less successful colony (Gorky) actually on the jump network. As
the latter colony failed during the nano-plague, they sent out
distress messages, which, arriving some few years later at Sheltersky,
spurred an expeditionary force to be sent. The commander of this
expedition, one Nidhi Blythe, upon finding the residents of the nearby
system already nearly expired (in large part due to overadaptation to
their current tech base rather than anything truly insurmountable)
upon her arrival, with only a few dozen survivors, set about turning
the remains of the former colony into her personal fiefdom. The system
had abundant natural resources and viable orbital infrastructure, but
no planets with admirable living conditions. Acquiring, through murky,
but generally believed thieving means, jump drive technology, Blythe
maneuvered her way into leveraging the support of Sheltersky for her
own benefit, being their only gateway to the outside universe. With
this support (a decade long pipeline exchanging personell and luxury
resources from Sheltersky for information and external access), remote
as it was, Blythe, her lackeys and their lineage were able to control
an entire star system against the minor threats of predation present
in the pre-SPEC era. Gorky came to prosperity during this era as a
smuggler's capital, a port so free as to be thought lawless, except
that it was always under martial law - it just so happened that as
long as the people being affected weren't of House Blythe, it likely
didn't happen to be illegal within Gorky's borders - although there
were some standards of behavior that fell beneath what House Blythe
considered civilized (such as slavery). Following in their mother's
aggressive tradition, the scions of House Blythe set up operations, in
a mix of above and below board trades, on an expanding number of
worlds. As they had no interest in bulk goods and that manner of
market control, they managed to co-exist with the CMT, rising to its
own power in the same time period. With the advent of SPEC, Sheltersky
became once more somewhat accessible, but by this time, it had been
sufficiently infiltrated by House Blythe that it was Sheltersky that
was consumed by Gorky and not the other way around.

As the SPEC era continued, and power consolidation ramped up, the sort
of activities for which Gorky was famous became less acceptable. By
the time of their joining the LIHW, most of the less reputable
operations had been shifted from Gorky to the less accessible
Sheltersky. Along with the arms dealers of Tribe of Eliana, House
Blythe is considered to be one of the most respectable of the criminal
groups operating within human space, with a public face, a reputation
for honesty in deals that it makes (albeit a keen, hungry, and
ruthless eye for advantage in anything not covered by agreement), and
an avoidance of more objectionable forms of illicit activities which
stems in part from their belief that they are far superior to any
common pirates - that they are a civilized organization that happens
to operate under their own code of laws, and not those of the
Confederation.

\item The Tribe of Eliana: 

An odd tale to be sure - the tribe of Eliana is so called because
every member of the group is a clone of the sole survivor, the
eponymous Eliana, of an otherwise failed colony. Their homeworld
rendered uninhabitable by the internal conflicts that broke out during
the nano-plague, the Elianas took up sparse residence on the other
worlds, moons, and worldlets of their system. While there had been
significant orbital infrastructure surrounding their colony, produced
on grand scale before the nano-plague, much of it had been damaged or
destroyed. Significant portions of the Elianas' efforts for several
generations focused entirely upon salvage operations performed on the
remnants of the orbital infrastructure so that items of value could be
retrieved before the chunks in question decayed in orbit and burned up
in the atmosphere. Their salvage expertise would eventually become
their trademark, as their post FTL acquisition undertakings were
primarily of the salvage variety. They became involved in many Ancient
artifact hunts, but became truly notorious for their "valkyrie" role
in stripping wrecks of all valuable systems and subunits. Their
criminal tinge comes from their willingness to sell what was formerly
anyone's to whomsoever is willing to pay for it (species
notwithstanding), whether it be insured cargo or military grade
weaponry. It is the weaponry trade that has proved to be the most
profitable for the impoverished Elianas, with many disreputable groups
finding them the only, if very expensive, potential suppliers of arms
generally not available to civilians. Policing of the Elianas have
proved of limited utility, as, while deals may be brokered in their
home system, such deals are made by individuals, not the government of
the Elianas, so responsibility is more difficult to sanction in
accordance with the gravity of offense. Moreover, while the deals may
or may not take place in their home system, the actual transfer of
goods rarely does, so concentrating efforts at policing the home
system of the Elianas (though not originally known as such, the
Elianas came to call it Gehenna, but, as there already existed another
system of that name in Confederation, the official name became
Yesteryear) has not proved remarkably fruitful.

Elianas outside of Yesteryear often have a quasi-nomadic existence,
moving from salvage operation to artifact dig to mining operation,
etc. There are no permenant Elianas settlements larger than outposts
outside of Yesteryear. However, there are a number of Elianas outposts
in otherwise lawless or unpatrolled systems, as local aggressive
groups tend to leave them alone, as they are often a source of
business partners for either the acquisition of, or disposal of, goods
of questionably transferred ownership. The Elianas themselves,
however, are not generally considered dangerous unless provoked or
interrupted, and the Elianas' government receives such a substancial
portion of its revenue from kickbacks and taxation of questionable
earnings that it has no practical choice but to decline to enforce all
aspects of LIHW law on its citizens. Such enforcement would be all the
more difficult due to the culture of sisterhood which defines their
unique existence.

\item The Order of the Dynast Shrub: 

The origins of the name of this group are lost to history but
generally believed to stem from a mistranslation of some older parable
or idiom. Starting out as a local family of robber-barons expanding
from a business selling fusionable fuels and antimatter to passing
starships, upon expanding outward from their home system, one branch
of the family took to collaborating with elements of various criminal
organizations with roots centuries old. The influx of new blood and
even murkier ethics moved the investment strategies out of fuels and
into the sorts of operations that wouldn't put them into direct
competition with House Blythe or the CMT - namely, operations
considered too disreputable for the merely greedy, such as human
slavery, vendetta by proxy, kidnapping, etc. The Order thus overlaps
the first and second tiers of pirates, being at heart a group of
violent thugs, but being in practice a group of very wealthy thugs
hiding behind their shell corporations. The Order is known to have
engaged in gang-wars with other smaller criminal enterprises,
absorbing them, or making them its vassals. Thus, the Order, along
with subordinate groups, though it is internally fragmented, through
luck and sheer callous brutality, has become one of the larger and
more powerful criminal enterprises in humans space, large enough to
negotiate with the Ulnish pirate cartels for "gentlemen's agreements"
as to what constitutes "invasions of territory."

\item Uln Pirate Cartels:

Though there are several of them, they are loosly organized and fight
internally more as violent siblings than as bitter rivals, coming to
actual exchanges of fire only when major assumptions need to be
revisited or on the demise of a powerful leader. They can thus be
treated as one entity, as they will consider the actions of any
outsider in similar fashion even if the actions did not take place in
their particular territory.

The Uln pirate cartels are part and parcel of Uln culture. Uln culture
expects corruption. It is deemed natural and appropriate that laws
will be circumvented, and a sign of personal power that one is in a
position to do so without being punished. The Uln pirate cartels
intermix their shipping with that of the Ulnish merchants trading with
all of their allies as well as within the Uln borders
themselves. Membership in merchant or pirate groups is fluid, and the
same Uln may pass back and forth easily between the low ranks of both
groups, albeit the loyalty required for the nepotistic cronyism
inherent in Uln leadership succession requires few such flip-flops if
one wishes to advance. Be that as it may, the fluid low-level
membership makes interdicting "known" members of the pirate cartels at
the border nearly impossible without squeezing trade to a standstill -
which none of the major powers is willing to do, as it would prelude
their access to the Ancient artifacts on the Uln homeworld.

Oddly enough, those at the most risk of Uln pirate attack are those
seeking to trade with the Uln, as the Cartels believe they have been
disrespected and slighted of their traditional cut of commerce by the
arrangements made between non-Uln traders and the Ingatwa and ranks of
royals. Fortunately, if they are particularly well armed for a pirate
group, they are not, in the grand scheme of things, a decided menace
outside the Uln borders, as they are generally outmatched by
Confederation, Aeran, or Rlaan responses. The Ulnish Cartels have
proved more problematic for the less well-off Shmrn, but remain an
aggravation and annoyance rather than a threat.
\end{itemize}
\subsection{The Stars Are Also Pyres}
\subsection{The Fraternal War}
\subsubsection{Fraternal War Historical Analysis}

\begin{itemize}
\item 2927 CE: First Contact \\ 

First contact with an alien sentient species by human explorers would
occur as the Unadorned discovered the bronze-age civilization of the
Mishtali.  These nomadic creatures, while actually quite civilized,
were possessed of the disturbing belief that eating the remains of a
vanquished foe would grant power to the victor, thus improving one's
life.  Still, the Mishtali adapted quite well to post-contact life
eventually being employed by their benefactors as baggage handlers.
The Mishtali are tripodal/polypodal beings with one strong upper limb
and two smaller manipulator limbs.  Each appendage is made up of three
individual limbs which can be used together as one limb with three
claw-like phalanges for grasping and handling objects, or separated
into three individual limbs terminated in a hooked claw.  Their main
body is supported by three separate spines connecting the jaws to the
stub of their three leg trunks.  The jaw/head portion is comprised of
three sections which hinge radially to reveal rows of long, sharp
teeth.  An optical organ is housed in each jaw section giving the
creatures approximately three hundred degrees of peripheral vision.

\item 2956 CE: Beckett's Murky Venture \\

The Lightbearers would be the next meme-group to encounter sentient,
or at least pre-sentient life in the form of the Dgn.  These bipedal
creatures were discovered dwelling in coastal marsh areas and
surrounding shallow seas on a world closer to Shaper space than
Lightbearer.  While the ancestral Dgn were not civilized, they were
exhibiting early signs of sentience in their use of tools and were in
possession of the required dexterity.  Oxygen breathers, they were
capable of living indefinitely on land as well as in oxygenated seas,
though without a humid environment, their skin and breathing orifices
would become too dry resulting in the need for re-wetting.  Moderate
resources were invested by the Shapers to ensure the success of the
Lightbearer mission until their role was expanded.  This was primarily
due to the need for their advanced experience with non-human DNA.
Though their goals were different, the potential benefits outweighed
the differences between the two meme-groups and a joint venture was
fully entered.  Genetic modification of the ancestral Dgn by the
Lightbearers gave rise to the Shmrn.  Eventually it would become clear
that key human genes were spliced into the Shmrn for the sole purpose
of equipping them to attain a full understanding of the superiority of
their creators.  This in contrast with the Dgn, who were uplifted by
the Shapers.  Their respect of individualist ideals, while not truly
extensible beyond humanity, prompted a more benign treatment.  Still
today the Dgn are treated justly, if not equitably in Shaper society.
While the Shapers may have been aware of it during their joint
operation with the Lightbearers, no action was taken to prevent any
mistreatment.  Ideological differences between the two factions
eventually led to the dissolution of the partnership, but not before
the two divergent species were crafted.

\item 2988 CE: Discovery of Ktah \\

Some time after the Lightbearers success with the Shmrn, they
discovered another sentient race dubbed the Klk'k (yes, it's a
transliteration).  The Klk'k, in contrast with previously encountered
races, were quite advanced, though not sufficiently to have begun
interstellar colonization.  Still, the race had all but conquered
their home world of Ktah and were well on the way to developing
sub-light capability. (The Klk'k had a limited orbital infrastructure
before the Lightbearers arrived. The debris from the destroyed
stations was one of the things noted by the Andolian scouting party
prior to deciding to engage the Lightbearer forces.)  These sentients
were bipedal with legs hinged similar to those of a bird.  Unlike an
avian, their legs are very muscular and quite long, accounting for
roughly two thirds of their total height.  The Klk'k have torsos with
a spine similar to a human's and a head which houses the most complex
portion of their central nervous system as well as two ocular organs,
two aural organs, and a single hinged jaw.  In fact they are
sufficiently anthropomorphic that their very existence was an affront
to the Lightbearers who believe that humanity is the embodiment of
evolutionary perfection.  They considered the Klk'k to be a mockery of
that perfection.

\item Occupation of Ktah \\

It was for this reason that the Lightbearers began a cruel campaign
designed to subjugate and 'correct' the Klk'k.  While they attempted
to mount a defense with their full military force against the
invaders, the Klk'k were simply no match for the technologically
advanced Lightbearers.  After five days (It wouldn't have taken
weeks. The Lightbearers didn't just happen upon the Klk'k and
immediately invade. Their initial scouting of the system detected
something worth investigating, and they returned after allocating
sufficient resources for an effort in force. While the sensors and
weapons available were few and paltry by the standards of 3276, they
were more than sufficient, in unscrupulous hands, of crushing the Klk'
k military. Many major military resources were assaulted before the
Klk'k were even aware of the Light-Bearer presence in their
system. The largest delay would have simply come from the small number
of actual primary vessels involved and fuel-conserving orbits of the
planet to reach all of the desired targets.) of planetary bombardment
resulting in heavy civilian casualties and an awed state of shock of
the general populace, the first wave of pacification vessels landed on
Ktah's surface.  The shock troopers were merciless in their handling
of the citizens of Ktah, engaging in horrible atrocities and causing
considerable suffering for the Klk'k. (Note that there were numerous
Klk'k guerilla groups retreating into the hinterlands of Ktah after
the bombing, but with the Lightbearers in complete control of the
orbital high-ground, air superiority, and with vastly superior
sensors, major actions were precluded from success, and not
attempted. Mostly, however, the Lightbearers just didn't care that
much about forces they were convinced couldn't hurt them.)

\item Andolians Discover Ktah \\

As the second wave of pacification vessels were being prepared aboard
the Lightbearers orbiting frigates (I wouldn't classify them as
frigates. The Lightbearers took the mentality that they were on
expeditions to pacify and rework planets, even those without ornery
native populations (intelligent or otherwise). Thus, their primary
craft of choice were essentially troopships armed with extremely clean 
"beach-heading" fusion warheads to clear out jungles, forests, and
any other potential inhospitable terrains so as to begin any outpost
on their terms, and not those of the planet in question. Each
troopship housed significant populations of soldiers practiced at
subduing local wildlife, numerous dropships and shuttles for moving
personnel and supplies, and sufficient supplies and materials to found
an outpost on an inhospitable world. The troopships would be
accompanied by attendant smaller craft designed to deter any pirates
or NeoBarbs encountered and perform recon and other secondary tasks.),
a small flight (as for details, I generally envisioned there being 3
Andolian ships and 2 LightBearer Pacification vessels, but any small
number on both sides would be fine with me, with the caveat that there
must be more Andolian scout ships than there were Pacification
vessels) of Andolian expeditionary vessels stumbled upon the system.
When the Klk'k's level of sentience and advancement became clear to
the crew, and thus the information was available to the entire
Andolian network, (Lacking communication infrastructure on both sides
of the jump point, and relays between there and the Andolian worlds,
the rest of the Andolian people didn't find out about Ktah until one
of the Andolian vessels went back to an Andolian world.) the Andolians
people were outraged at the callous disregard the Lightbearers had
shown for the sanctity of sentient life, and detached one of their
group to begin a return to Andolian space to alert the rest of the
Andolian population.  A decision was hastily reached to take immediate
action to prevent further launches and liberate the besieged world.
Captain (If commanding a squadron of ships, the rank is more likely
Fleet Captain (if temporary) or Commodore, if he is normally in charge
of a group of ships) Gil Neragel, Senior Officer of the Andolian
scientific vessel ASV Poole(?), the lead vessel in the exploratory
flight, desperately launched a series of unmanned scientific probes on
a collision course toward the Lightbearer frigates and landers.  The
surprise maneuver caught the inexperienced pilots unawares and he was
able to destroy a handful of the would be invaders.  Once the initial
surprise diminished, however, the frigates closed on the lightly
shielded vessels and The Poole and two other vessels were incinerated
before they could flee the system, the first Andolian casualties of
the war.  Captain Neragel is still lionized today by the Klk'k as the
embodiment of Andolian integrity and fortitude.  (So, here's how I
always had this going down in my head: As per "Though the Andolians
had only a small squad of exploration craft, they defeated the
Lightbearer subjugation force by using their own unmanned vessels as
weapons" (wherein by unmanned I really meant ``unmanned because they
had been evacuated'' namely, they rammed the
troopships. More detail follows:

	With one of their initial grouping departed back through the
	jump point, the remaining Andolian scout vessels, still
	unnoticed by the Light-Bearer forces, and having observed that
	the troopships were maintaining a highly regular, minimal
	power orbit, set about plotting an intercept course. The event
	was timed to occur at night over a populated continent (not
	that difficult with a low, 100 minute orbit). The scout ships
	began gathering momentum before going into SPEC, briefly
	dropping out well short of the target area to disembark
	lifeboats on a lazy ballistic arc that would intercept the
	planet a day later. They then reoriented and continued
	accelerating full throttle before returning to SPEC. While the
	Lightbearers could now begin to detect something incoming,
	they were not prepared for a well-planned kamikaze attack, and
	were not able to intercept the incoming scout ships far enough
	out to avoid damage. One of the two pacification vessels
	exploded from the impact, a bright death-flare visible in Ktah'
	s night sky. The other was sufficiently damaged to be no
	longer truly spaceworthy, and the surviving crew and supplies
	had to be removed by support vessels, and transferred to the
	planetary outpost. The Lightbearers sent back one of their
	precious few remaining jump capable craft, a refueling tanker,
	to gather more reinforcements, escorted to the jump point by a
	sizeable fraction of their parasite craft. The bright flash,
	disappearance of one of the two vessels, and damage on the
	other visible to any Klk'k with a really good telescope did
	much to leave the natives especially restless. The two dozen
	crew in the lifeboats from the two scout ships would land in
	one of the many bayou areas of Ktah on the larger of the two
	super-continents. Allocation of Lightbearer forces to deal
	with Klk'k unrest and guerilla attacks in the wake of the
	visible orbital occurrences would serve to allow the Andolians
	enough time to flee their landing sites before they were
	pinpointed and bombed.

(Note: the particulars of names, exact numbers, and ranks don't
usually concern me too much (and this applies to things I've written
as well), so feel free to name most ships/people as you see fit. If
anything seems particularly inconsistently named or disproportionate,
or needs to be a specific something-or-other I'll make a specific
comment at that time. Otherwise, assume the names are fine) )

\item Andolians Declare War \\
	
At this time interstellar navies were barely a shadow of what they are
today.  But, The Andolians were able to assemble a small fleet of
their most modern armed vessels, dispatching them to the Ktah system.
While the Andolian fleet was small and led by officers with little or
no combat experience, the Lightbearers' frigates were driven from
orbit in a grueling contest.  Not to be denied, the Lightbearers
returned with a larger force prompting the first major full scale
spacial naval engagement in human history.  Both factions were able to
land several waves of drop ships and the battlefield became a
threefold undertaking, with forces engaging in orbit, in Ktah's
atmosphere, and on the ground. (While I didn't envision air and close
orbital combat taking place this early, nor even the order/magnitude
of some the events above (see below), what most doesn't fit for me is
the ground combat part -- the Lightbearers specialized
in making pacification vessels, troopships for conquering (admittedly
uninhabited) worlds. Even repurposed from civilian
colonization/infrastructure tasks, the Andolians wouldn't have had
anything that could both be sufficiently militarized on short notice
and could land enough troops to counter the sort of ground presence
the Lightbearers would have by this point --
particular, the lack of truly viable large-scale ground assault craft
in the existing Andolian ship repertoire is a key reason for why the
Hoshino uprising was instigated. That said, other than the ground bit,
as long as it doesn't end up affecting the Hoshino Uprising too much,
the early fleet actions can be pretty flexible as to what happened, so
we'll see how everything looks when we read it all over again)

(Again, this part played out a little differently in my head: The
Lightbearer forces were nearer to their homeworlds than the Andolians
and were the first to return in force, albeit not with one
specifically designed for the tasks at hand. However, the
Light-Bearers did make good use of the ships they did have,
dispatching several partially and some fully loaded troopships to
reinforce the garrisons on Ktah and begin work on entrenching their
defenses. Accompanying them was a mixed group of assorted smaller
vessels and parasite craft. While the Lightbearers sought out the jump
point from whence the Andolians had arrived, they did not find it
before a detachment of the Andolian's most modern interdictor vessels
appeared to defend it. Attempts to drive the Andolian craft from the
jump point were rebuffed, and additional Andolian craft poured into
the system. As a new wave of Lightbearer reinforcements gathered to
transit from the jump point leading to Lightbearer space to the forces
encircling Ktah, the first full scale spacial naval engagement in
human history began. The more nimble Andolian craft exacted a heavy
toll on the sluggish troopships and supply vessels before finally
succumbing to the superior numbers of the escorts and parasite
craft. Still, the battle was an economic victory for the Andolians,
and hindered efforts to further fortify the Lightbearer positions on
and around Ktah.  ) At 16:40 hours on the 245th day of the year 2988
CE, Aldra Heldous, Commander of the Provisional Andolian Expeditionary
Navy(?)(They certainly wouldn't worry about it being too long, so it
sounds fine to me. Also feel free to change your mind if anything you
think is better comes to you.) announced at a press conference that
all diplomatic overtures had been exhausted and were insufficient to
bring an end to the already broiling conflict over Ktah.  One hour
later, the an Andolian Prime Minister(?)  Speaker (In the
hierarchically elected Andolian government, many elected officials
are, informally, Speakers, and more formally, Speakers for XYZ (the
previous level in the hierarchy). The Andolian government is also
intentionally redundant (and overlapping, but that's beside the
point) to some degree (to both avoid problems manifest with the
beheading of a hierarchy and to limit the depth of the hierarchy while
still managing individual workloads) thus, even at top level, there
are multiple Prime Speakers, hence, "an Andolian Prime Speaker")
made an official declaration of war before the Joint Colonial
Council(?). ( ? Not sure what you wanted the "Joint Colonial Council" 
to be. There wasn't any central organization between sovereignties
before the confederation, and the Andolian government doesn't
physically meet anywhere (beyond all being on or near Kubernan for
reasons of communication latency) in a way that would really be 
"before XYZ council". Any public announcement for the Andolian
government is an announcement "before all Andolians".  If the JCC is
a Lightbearer organization, then a Prime Speaker wouldn't be there,
so I'm assuming that's not the intent. The key action, as unromantic
as it sounds, was most likely an Andolian foreign minister contacting
a set of Andolian ambassadors, and having them deliver appropriate
notification to appropriate parties. If I've completely missed what
you were going for, do let me know.)  The political backlash was
immediate and severe, originating primarily from Bifrost, the Shaper
capital.  Rhetoric from most of the major meme-groups included
statements condemning the use of force against the sovereignty of the
long established Lightbearers.  The Shapers secretly pledged to
support Lightbearer efforts to repel Andolian forces, favoring their
arrogance to Andolian imperialism, while pointedly avoiding overt
conflict. (Yup).

\item 2989 CE: Hoshino Uprising \\

(Some parts of this are particularly well set in my mind, and differ
notably.)  Word of the Andolian's arrival spread across Ktah's ravaged
surface like a deluge soaking a parched field.  The welcome news of
any power that might stem the tide of the unstoppable invaders became
a powerful catalyst for the populace.  Components of the nearly
decimated Klk'k military began to come out of hiding.  Civilians took
up arms against their would be conquerors using whatever means
necessary to defend themselves, home by home, street by street, and
city by city.  The Andolians gained a foothold on one of the larger
continents, shipping in thousands of crates of small arms and military
supplies.  They instituted hastily conceived training programs to
prepare the Klk'k freedom fighters to face the formidable Lightbearer
shock troops.  As the liberators regained orbital and aerial
superiority, the Lightbearers were being pushed back on all fronts.
Fierce combat raged in the orbital arena with the Lightbearers
attempting to shut down the Andolian supply lines.  And indeed, it
required virtually all available Andolian warships to maintain those
critical lines through the jump network along with the continued
suppression of orbital resistance over Ktah.  But the Andolians were
as determined to break the siege as Ktah's people were to reclaim
their homes.  It became obvious to the Lightbearer command element
that the world would be lost.  And worse, Andolian forces were now
targeting their assets in systems considerably removed from Ktah and
the supply convoys.  A snap decision was made to pull back whatever
forces could be quickly rescued from the planet's surface.  The
admiral(?) in command of orbital operations ordered the launch of
several high yield laser induced nuclear fusion devices.

(Here's how it has played out when I've mulled over it before)
Firstly, to give more substance to the name Hoshino (for whom the
uprising is named): Anders Hoshino was the ranking survivor (at the
time of the uprising) of the initial two-dozen Andolians to make
planetfall.

Some of the scout-craft crew both survive to, and have survived the
experience of, making contact with a Klk'k guerilla group. Key
assistance rendered by the planetside Andolians is knowledge of the
larger picture. Assistance from the Andolian forces in the rest of the
system to the handful of planetside crew is slow in coming, and
consists of small teams of specialists with communications gear. Even
with covert insertion tactics only a portion of the teams dispatched
make it through the orbital blockade and increasingly entrenched
planetside Lightbearer anti-ship defenses, and not all of those manage
to make contact, or are accepted by Klk'k guerillas. However, a
couple of insertion squads do reach the scout crew, and establish
contact between the ground forces and the Andolian fleet.

Hoshino's crew and the Klk'k guerilla fighters bide their time until
informed that the orbital supremacy of the Lightbearers is nearly
broken. Unfortunately, many of the Lightbearers' ground defenses have
been built in the middle of heavily populated Klk'k areas so as to
use the civilians as a shield against bombardment from beyond their
engagement range. In order to mitigate losses of the (converted
civilian) Andolian landing craft, some significant portion of these
emplacements would have to be neutralized.

The Leader of the Klk'k guerilla group proposes inciting a mass
uprising using symbolism from the overthrow of the Theocracy that once
ruled much of Ktah. A video message is recorded featuring Hoshino and
the guerilla leader each blowing into one side of a traditional Klk'k
nasal horn (While not how it's usually played, i.e. it's usually
played by a single Klk'k blowing through both nasal passages, the two
person playing of a horn is a long-standing Klk'k symbol of unity
(one of these days I'll have to remember what the heck, if anything,
I called the nasal horn)) while the guerilla leader holds an Amakakt
Tklatl (dual-bladed martial arts staff of the plebian classes and
traditional weapon of the River Kingdoms of the lesser continent,
traditionally carved from the "wood" of an Obelisk "tree". For
Tklatl and Nasal horn together, see: Klk'k Flag/Emblem) and Hoshino
holds a family heirloom mono-blade that he has tended to use in lieu
of the standard issue Andolian exploratory service machetes (Hoshino
was authorized to use the non-standard equipment because it had proved
more effective in testing than the standard issue machetes. A brief
investigation at the time of authorization into using similar
equipment to replace the standard issue machetes terminated upon
initial cost evaluation.) . Textual and auditory messages concerning
the precarious Lightbearer position are overlaid. The guerilla forces
lead an assault to take over a Lightbearer communications node
slightly before the time as the Andolian fleet is scheduled to attack
again, and upload the recorded message into the Light-bearer installed
propaganda distribution network (which they'd have, because what's
the use of being superior if you can't lord it over your lessers
(people are that petty, and the Lightbearers were everything we are
and more, etc.). Surprisingly, the Andolian technicians were able to
keep the transmission link open for longer than expected, perhaps due
to a combination of shocked dumbfoundedness and a lack of
understanding of the importance of local symbolism on the part of the
Lightbearers. The recorded message completed, and they cut to a live
feed. Hoshino, who had been badly wounded in the taking of the
communications node, uttered a variation on a Klk'k phrase amounting
to something like "Should my bones here rest, so shall I be honored" 
upon which the guerilla commander (who I never bothered to give a Klk'
k name to because I always thought of him by the human name he would
later take as the first Simon.) gingerly took up Hoshino's mono-blade
that was lying on the floor, turned and spoke directly to the camera
what translates to "And the stranger came to his brother's house",
the beginning of a famous passage of a recounting of the aid of the
River Kingdoms in overthrowing the Theocracy (For notoriety level
think "Friends, Romans, countrymen, lend me your ears" ). Just after
this, the scene was rocked by explosions as the Lightbearers sought to
retake the communications node, and the signal was shortly thereafter
returned to Lightbearer control. Almost immediately, anywhere that the
message had propagated to erupted in conflagration. Even those places
where it had only come as audio, heralded by the unmistakable sound of
a Klk'k nasal horn, were roused. Within minutes, Klk'k were being
shot down across the planet by the thousands, but continued to spring
to the battle call with projectile weapons, vehicles, chemicals,
knives, sticks, rocks, and their naked rage and indignation. They had
freed themselves from tyranny before, and all they had needed to spark
a roiling fury was a hope that blood shed would not be in vain. Wave
attacks overwhelmed more isolated Lightbearer positions, driving them
back into their central compounds.

Guerilla stockpiles were employed in more organized fashion against
key targets, and holes began to open up in the Lightbearer ground
defense network. The Lightbearer forces in orbit were crushed. With no
means for extraction, and an Andolian invasion now imminent, set to
pour through the holes opened up by the disabled ground emplacements,
the Lightbearer commander opted for a scorched-earth policy,
retargeting the anti-orbital and anti-lander missile batteries to
launch ordinance on paths that would fall back to Ktah.

 ( and here we reconverge )Most of the heavily populated metropolitan
 centers were targeted.  Had the Lightbearers been equipped at that
 time with dirtier fission warheads, they most certainly could have
 salted the planet, effectively rendering the Klk'k and their world
 extinct.  Fortunately for the native species, and for the Andolian
 liberators on the surface, they did not have that option.  Still, the
 massive blasts alone caused over four hundred million deaths
 instantly, and another one billion died in the following days.
 Whatever Klk'k infrastructure that had survived until this point was
 now defunct, blanketing the beleaguered planet with utter chaos.

\item 2990 CE: Andolians Liberate Ktah\\

The furious response by the Andolians was swift and brutal.  Most of
the ships in the Lightbearers' occupation fleet were destroyed before
they could make the jump out of the system.  Those that survived were
hunted mercilessly and either destroyed or chased deep into
Lightbearer space.  Some pacification troops still remained on the
surface, abandoned by hard pressed naval vessels fleeing orbit.  These
forces were quickly neutralized in a joint effort with Andolian troops
and Klk'k freedom fighters.  Many of them were taken into custody as
Andolian military units established POW camps in former Klk'k
facilities.  These prisoners would be the first of many Lightbearers
to be captured during the war.

\item Ktah Restoration Project \\

Andolian naval forces would soon discover that the Lightbearers were
loathe to accept defeat.  Incessant attacks on shipping lanes would
stall humanitarian relief efforts for Ktah's citizens.  Some of the
other meme-groups, and particularly some of the more benevolent
corporate entities, would pledge additional resources to aid in these
efforts.  Still, sharp criticism was targeted at the Andolians from
the Lightbearers' historical allies, primarily the Shapers.  The media
would dub the conflict The Fraternal War citing the Andolians'
disregard for the mutual good of humanity.  Hundreds of thousands of
Andolian citizens, both private and military (in time of war, this
distinction is even more difficult for Andolians than it is usually( I'
d use the term "civilian" rather than private, as they would
certainly be effectively an arm of the Andolian group will
(i.e. government) in such an effort rather than ``private citizens''
which is what I take to be the current meaning.), began arriving on
Ktah to aid in rebuilding critical infrastructure components and
restoring basic necessary services to the Klk'k populace.  This effort
would become the sole focus, at least publicly, for the Andolian
military-industrial complex.  Minor skirmishes were to occur, both
over Ktah and beyond the system, between the Andolians and the
Lightbearers.  However, most of the major battles during this time
took place along the shipping lanes as the Lightbearers stepped up
their campaign to sabotage the restoration efforts.  Over the course
of the following months many overtures would be made by the Shapers
and several other meme-groups calling for a peaceful resolution.
Offers were levied to broker a peace between the warring polities.
But, Lightbearer arrogance and Andolian indignation would continually
block these efforts at every turn.

\item 2991 CE: Andolians Reveal Spaceborn \\

As Lightbearer raids began to come with less frequency, the Andolians
started to make forrays into Lightbearer space, often staging intense
raids on orbital facilities.  On occasion the stations would be
captured to serve as a forward operational facility for the Andolian
navy.  It was during these raids that the Andolians discovered the
existence of the Spaceborn.  The Lightbearers had secretly plied their
considerable skill in manipulating human genes to create a subservient
race of humans known as homo sapiens cosmonatalis.  Their basic
genetic code had been altered in such a way that they were unable to
survive within a gravity field, but provided considerable advantages
for operating in micro-gravity.  The Spaceborn suffered none of the
ill effects of prolonged exposure to zero-gravity environments, and
were being used as slaves by the Lightbearers to perform construction
functions considered too menial for themselves.

In an effort to mute Lightbearer support, this information was leaked
to GNN correspondents.  The Andolians, being well aware of the
Shapers' strict policy against any form of human slavery, knew that
this information could serve to demonize the Lightbearers in the
public view.  The story became headline news across the human
colonies, especially on Shaper controlled worlds.  Shaper leadership
was so shocked by the claim, that they dispatched an independent
investigative team to confirm the validity of the Andolians' report.
The committee was not only able to verify the claim, but made further
discoveries of the depth of the Lightbearers' use of the Spaceborn.
Some of the interviews with freed Spaceborn men and women were leaked,
most likely by the Andolians again, which set off a devastating public
backlash against the Lightbearers.

The financial and intelligence support which the Shapers had been
secretly providing to the Lightbearers dried up almost immediately,
though the Shapers and other polities were still calling for a
diplomatic resolution and releasing statements condemning the
Andolians indifferent use of force.  The Lightbearers themselves,
having little defense against the accusations, were eerily silent in
the face of bitter objections over their Spaceborn program. (I'm not
sure that "leaked" is the best word to use in the above two
paragraphs -- maybe the latter of the two, but not the first. The
secret being kept by the Lightbearers, it would only be a "leak" if
the information came out of their end. The Andolians, certainly in the
middle paragraph, would have been quite loud in announcing the
existence of the Spaceborn. As to whether the Andolians would leak
interviews it's not entirely in character, insofar as they'd
probably have done them in public if they wanted them broadcast as
such, or debriefed in tight secrecy if they were concerned about
secrecy. Now, working with third parties of known unreliable
informational security to handle external validation of claims -- that'
s not beyond Andolian character, but they wouldn't generally bother
with leaking something themselves. )

\item Arms Proliferation Begins \\

The Andolians, deeply embroiled in an ever widening conflict with the
Lightbearers, had already increased production of military vessels and
equipment, as well as investing a large share of their capital
expenditures into projects with weapons production potential.  This
shift in resource allocation was not isolated to the Andolians,
however.  As the stories continued to pour in from the front lines of
the war, the leadership of every major human political group soon came
to recognize the need for increasing the strength of their military
arms.  A new arms race began, one motivated primarily by fear of
Andolian dominance, though in some small part the discovery of now
three non-human sentient species gave rise to a fear of potential
threats from sources beyond mankind.

The Shaper and Purist factions were perhaps in the next best position,
along with the Lightbearers who had already begun to suffer some
rather humiliating defeats to the ever advancing Andolian navy.  The
general consensus among the human polities not directly involved in
The Fraternal War centered around the belief that the Andolian and
Lightbearer military influence would diminish with time, creating a
power vacuum to be filled by the fastest sprinter.  Other factions,
such as the Highborn and Mechanists, would place a greater focus on
research, the latter making great strides during this period in
cybernetic advancements.

Within the span of barely one year, governments which once commanded
small policing navies designed to foil piratical raids would be
transformed into formidable offensive forces capable of projecting
power far from their operating bases.

\item 2992 CE: Andolians Expand War \\

No group seemed better able to refocus their production capacity than
the Andolians.  As the evidence of their success continued to grow.
The Andolian leadership seized the opportunity for conquest.  Research
was underway that would eventually lead to the development of early
capital ships, including the eventual production of the Thales class.
Predecessors of this ship, still in use today as a corvette by several
minor groups, began pouring out of the Andolian shipyards.  In
addition, many advances were made in engine and reactor design,
enabling the production of more advanced, more maneuverable fighters
and assault craft.

The Andolian navy would soon become a proving ground for these new
designs.  Prototypes would be tested in the heat of battle.  New
secondary systems were being retrofitted to existing vessels.  And the
Andolian Joint Colonial Council (Ah, here it is. That's what you were
talking about earlier. Hmm. I don't like the name. but I don't
have a better one yet. It certainly wouldn't have the word "Colonial" 
in it. "Speaker's Forum" or something along those lines might be
better.) authorized a broadening of the conflict with the
Lightbearers.  Studies published by the Andolian intelligence
community indicated a high degree of likelihood that there could never
be successful negotiations with the overly proud Lightbearers.  In
fact, these were accurate, if exaggerated, claims.  News media in
Lightbearer controlled space often featured biting commentary and
political cartoons which depicted not only defeat, but total
elimination of the Andolian people, coining phrases such as 'Link
Junkie'.  It is a widely held belief that the inner circle of Andolian
political influence had decided early on to bring the entire
Lightbearer faction to its knees. (Note that there isn't any "inner
circle" in the traditional sense in Andolian politics. Note also that
this wouldn't dissuade other groups from believing that there was/is
because they don't really comprehend the Andolian mindset and
existence. (But, it does mean that word choice and phrasing can mean
something when discussing it.)

That is exactly what the Andolian navy was doing.  Orbital facilities
were falling under Andolian control on a regular basis.  Several
successful raids had been staged in many of the Lightbearers' eleven
core systems, with focus on infrastructure targets and production
facilities.  The Lightbearers had made several efforts to mount
counteroffensives into Andolian space only to be repelled nine times
in ten.  And each effort only served to strengthen the resolve of both
the Andolian military and her civilian populace.  The war had taken
its toll, however, on both sides.  The Lightbearers had surrendered
several key systems housing minor colonies in favor of withdrawing
their fleets to more populated and valued systems.  And the Andolian
war machine began to grow thin as the battle line grew longer.  The
war had become a war of attrition, each side's navy diminishing over
time.

This fact did not escape the notice of the Shapers' intelligence
community.  Calls for a cease-fire increased in regularity and
intensity, even to the extent of veiled threats to enforce these calls
with military backing.  Word from Cradle, the Lightbearer capital,
indicated that the normally arrogant meme-group would be receptive to
talks, drawing more pressure on the Andolians to scale back their
offensive efforts.  Many military analysts agreed that the Andolians
had played out their capacity, claiming that further advances would
carry too high a cost for the Andolians to continue.
	
\item 2994 CE: Andolians Defeat Lightbearers \\

Then, in a move which stunned the entire human community, the
Andolians launched a fresh campaign employing the largest single
military armada ever seen in human history prior to its time.  The
Andolians had reserved nearly two thirds of their production capacity,
effectively concealing the vessels from the public.  Further,
investment in weapons research had led to the production of second
generation vessels far more capable than anything previously produced.
This fleet is often referred to as the Ozawa Fleet in reference to the
ancient sea battle in which a japanese naval officer used a large
force as bait to conceal two aircraft carrier task groups.  However,
unlike the catastrophic result for the japanese carrier force in that
ill-fated battle, the Andolians had played their trump card to
perfection.  The fleet stormed past several prime targets in minor
systems, going straight for the Lightbearers' jugular vein.  It was
divided into five task groups, each striking separate leading
production worlds simultaneously in a fabulous display of coordination
and planning.  Orbital superiority was established surprisingly
quickly, then light bombardment operations were undertaken as heavier
kinetic bombardment tactics were prepared.

The Lightbearers, expecting the Andolians to become bogged down with
restoration efforts on the battered worlds as had been the case with
Ktah, desperately withdrew forces to neighboring systems in
preparation for guerilla campaigns mirroring those enacted against the
Ktah Restoration shipping lanes.  This was a serious miscalculation on
their part, for the Andolian strategy had shifted from capturing
assets to annihilation.  The Ozawa fleet moved from system to system,
destroying all off-world defense before refocusing on planetary
bombardment.  These kinetic bombardment operations were devastating to
the planetary ecology.  The goal of the attacks was to produce as much
ejecta as possible, filling the atmosphere with dust and ash and
effectively blotting out any solar energy from the systems sun.  The
practice came to be known as 'dusting a world,' and was immediately
condemned by the human community as a whole.  Public outcry against
the war increased to a disturbing level, several polities entertaining
plans to intervene.

But there would be no time, neither for diplomacy nor for
peacekeeping.  Within just over two weeks, the core worlds had been
rendered dead.  Andolian troopers soon gained control of the lesser
colonies, those which were deemed small enough to effectively control,
and the Lightbearer military was rendered ineffective.  The human
colonies stood aghast at the utter devastation, and at the utter
success of the Andolian strategy.  During the weeks that followed,
Lightbearer citizens were herded into internment camps, many of those
in leadership positions being charged with war crimes and crimes
against humanity.  Millions of Spaceborn and Shmrn slaves were freed
and provided with relocation plans.  The Shapers, while still
condemning the Andolians for their methods, were livid about the
Lightbearers' creation and use of human slaves.  They began to
construct cases against several of the accused, insisting that their
trials should not be an Andolian matter, but an issue for all of
humanity to weigh.

Over the course of several weeks, in the absence of any supporting
infrastructure, the last of the Lightbearer military was summarily
destroyed or captured.  Final casualty estimates indicate twelve
billion Lightbearer dead (I'd up the death toll a bit from ~60\% to
75\%-80\%), including heavy civilian casualties (*cough*
understatement *cough*), perhaps as many as one billion Shmrn slaves
who died beside their masters in the bombardment, and around three
hundred million Andolians (I'd imagined the Andolian casualties to be
a bit lower, given that the bulk of all casualties were from planetary
bombardment and most of the war was fought in Lightbearer or
under-occupied space. However, it would only take a handful of attacks
on metropolitan areas to drive up an Andolian deathtoll quite a bit,
so that's not an implausible sum. [aside -- just had the odd thought
that those who read "billion" as a million million will find this
much more disturbing]) , this in addition to the one billion four
hundred million lost on Ktah, brings the staggering total casualties
to almost fifteen billion living souls.  The Andolians are now
credited with more human loss than any other polity in history, both
before and since, The Fraternal War.

\item 2996 CE: Andolians Turn Over Captives \\

The Andolians now begin the daunting task of handling the remaining
eight billion Lightbearers.  Through intense diplomatic debates it is
decided by the major human powers that the prisoners be placed in the
custody of the three groups which had suffered the most under their
policies, the Klk'k, the Shmrn, and the Spaceborn.  A human oversight
and steering committee is appointed the task of advising those
custodians.  Surprisingly, the Shapers were leading proponents of a
forced sterilization program for all surviving homo sapiens suprahomo.
During the final days of the war, Lightbearers fled the core worlds en
mass, many seeking refuge in Shaper and Highborn systems.  It is said
that many of those refugees that fled to Shaper space met with a far
darker reception than they might have expected, though any individuals
who possessed useful knowledge, most especially in the field of
genetics, were afforded some safety, if not freedom, to impart that
knowledge to their new hosts.

Many of the high level military commanders and political leaders would
eventually be tried and sentenced, usually to death.  However, most of
the Lightbearers were subjected to forced sterilization and allowed to
live out their days with minimal intrusion and monitoring as long as
no tendencies toward their previously held beliefs were detected.  And
wherever those traits were detected, The Simons were quick to respond.
The Lightbearer meme-group was essentially rendered defunct, its few
remaining survivors being integrated into other societies.

\item 2997 CE: Andolian Protectorate Established \\

Having maintained a protective watch over the Klk'k since their
liberation, the Andolians make an official declaration continuing this
arrangement indefinitely.  The Treaty of Ktah is signed by an Andolian
delegation and leaders of the Klk'k in the chief city of the named
planet.  The Spaceborn, most of whom lived under the protection of the
Andolians at that time, are later included in the Protectorate.

The treaty establishes military and financial backing of the two
lesser factions and promises of cooperation among the three.  The
Klk'k government is afforded some autonomy, though technically the
Andolian Protectorate, as a governing body, maintains jurisdiction.

\item 3002 CE: Shmrn Begin Colonization \\

The Shmrn, now free of Lightbearer oppression, begin independent
colonization efforts.  They are forced to accept many less than
desirable worlds, as the richer opportunities have already been
capitalized upon.  Still, they begin to establish themselves as a
minor power in the political landscape.


\item 3003 CE: Andolians Cede Control of Gained Colonies \\

After seven years of occupation, control of several former Lightbearer
worlds is ceded to other factions, primarily the Shapers, as
recompense for lost investments.  The deal is brokered by the Shapers
and agreed upon by the Andolians in an effort to mute continued public
condemnation for their brutality (ruthlessness is generally a term
more applied to the Andolians than brutality) during the final days of
the war.

\item 3006 CE: Confederation Talks Commence \\

In the aftermath of The Fraternal War, all of the major human factions
begin to suggest the formation of a cohesive confederation of worlds
to cement mankind's security and prevent a second fraternal war.  This
comes on the heels of several skirmishes between minor human groups
which threaten to destabilize certain fringe regions.

Most major meme-groups, in spite of their vast differences in
ideology, increase diplomatic pressure on the Andolians to capitulate.

\item 3006 CE: League of Independent Human Worlds Formed \\

As rumors and reports of the confederation talks persist, several
independent human worlds schedule a convention to discuss how it will
affect them.  Many of the leaders in attendance author a document
establishing the League of Independent Human Worlds.  The main thrust
of the league is to provide some representation of these worlds when
the confederation is formed, and to avoid eventually being co-opted by
the major meme-groups.

\end{itemize}

Important Dates: (dates are proposed) (Mine aren't perfect either, but they're my first pass at things.)
Exploration Period Begins
\begin{itemize}
\item 2927 Unadorned discover Mishtali (First Contact: Humans)
\item 2956 Beckett's Murky Venture
\item 2988 Lightbearers discover Ktah
\item 2988 Lightbearers occupy Ktah
\item 2988 Andolians discover Ktah
\item 2988 Andolians declare war on The Lightbearers (Fraternal War)
\item 2989 Hoshino uprising, nulcear retaliation by The Lightbearers
\item 2990 Andolians liberate Ktah
\item 2990 Ktah Restoration Project
\item 2991 Andolians reveal Spaceborn
\item 2991 Arms proliferation increases across majority of human space
\item 2992 Andolians begin to systematically debilitate Lightbearer faction
\item 2994 Andolians defeat LightBearer faction, liberating Spaceborn and Shmrn
\item 2996 Surviving Lightbearers turned over to Klk'k, Spaceborn and Shmrn custody.
\item 2996 Lightbearer meme-group rendered defunct
\item 2997 Andolian Protectorate established
\item 3002 Shmrn colonization efforts begin
\item 3003 Andolians cede control of several former Lightbearer colonies to other factions
\item 3006 Diplomatic talks commence concerning Confed
\item 3006 LIHW formed
\end{itemize}

Historical Information:

Klk'k

The Klk'k had a level of technology roughly equivalent to 21st century
Earth when they were discovered by the Lightbearer faction. The
Lightbearers, believing themselves evolutionarily superior to all
other species and factions, made little note of this fact and
proceeded to enact harsh subjugation on the Klk'k. Not long after this
subjugation began, the Andolians stumbled across the
situation. Perceiving the Klk'k's level of advancement in both culture
and technology, the Andolians decided they could not permit their
enslavement by the Lightbearers. Though the Andolians had only a small
squad of exploration craft, they defeated the Lightbearer subjugation
force by using their own unmanned vessels as weapons (What I really
should have said here was ``evacuated vessels'' instead of
``unmanned'' much more clear that way) This event led to The Fraternal
War between the Andolians and Lightbearers, and to the Klk'k joining
the Andolian Protectorate.

Lightbearers

From a planet discovered on a joint mission with the Shapers the
Lightbearers cultivated some of the early Dgn, further twisting their
development along their own designs until they settled upon what are
now the Shmrn.

When later a Lightbearer exploratory group discovered the Klk'k
homeworld they prepared to subjugate, and "correct" this species,
whose somewhat anthropomorphic construction they viewed as a mockery
of the divine human form. Fortunately for the Klk'k, the Andolians
discovered them only some weeks after the first Lightbearer
pacification vessel had arrived at the planet. The Andolians were
outraged at the Lightbearers' actions against civilized sapients, and
threw themselves against the Lightbearers in the first full-fledged
human interstellar war.

The Andolian's greater manufacturing base outperformed the
Lightbearer's economy. As the Lightbearer military-industrial complex
collapsed, a secret was revealed that not even the Lightbearer allies
had known about, namely, the existence of the Spaceborn a genetically
engineered slave race of humans, designed to live their lives in zero
gravity so as to prevent Lightbearers from having to deal with such a
menial task as laboring in vacuum.

It is this revelation that is believed to bear primary responsibility
for the lack of action taken by any other faction when the Andolians
proceeded to eliminate not only the industrial capacity of the
Lightbearer, but also the Lightbearers themselves.

Those that did not manage to escape to Shaper or Highborn space, or
were not fortunate enough to be killed in the assaults on their
worlds, had the dubious honor of being turned over to the Klk'k, the
Spaceborn, and the Shmrn.  The ensuing combination of sterilization
and incarceration served to eliminate the Lightbearer meme from the
realm of dominant thought.

Spaceborn

The Spaceborn were genetically engineered in secret by the
Lightbearers as a servant race, to perform undesirable zero-g tasks
for their suprahuman "superiors". During The Fraternal War between the
Andolians and Lightbearers, the Spaceborn were discovered and
liberated, and sided with the Protectorate as a result.

Shapers

Initial Shaper political, financial, and resource support for the
Lightbearers during the Fraternal War was linked not to political or
ideological support of the Lightbearers, but a firm opposition to what
they felt to be Andolian Imperialism and the desecration of the
natural sovereignty of the Lightbearers (clearly, it
didn't help that the Shapers thought the Andolians to be even more
ideologically disturbed than the Lightbearers; while the Shapers
thought the Lightbearers perversely misguided in their vision of
embodied human perfection, it was at least a goal that resonated with
their society, whereas the Andolian lust for abstract and vaporous
group progress that would somehow lead to a "bettering" of mankind
and thereby the individuals thereof struck the individualist Shapers
as a demented non-sequitur that was downright unintelligible. The
interjection of an expanding, and, if human, foreign in an almost
alien sense entity into a neighboring polity was clearly cause for
alarm, never mind the Klk'k -- they, after all, weren'
t human. That the Shapers cut off all support for the Lightbearers
after the revelation of the existence of the Spaceborn indicates,
importantly, that, at the time, they either thought us, or themselves
(depending upon one's perspective) to still be human, and thus
deserving of the fundamental human rights that run through the core of
their belief system. While some Lightbearers did flee to some degree
of protection within Shaper systems toward the end of the war, it was
that same sort of protection offered Nazi scientists captured by the
Russians after World War two -- protection of ones
life, but not an opportunity to live, only an opportunity to be worked
until whatever useful knowledge you had could be extracted and you
yourself forgotten. Much has been rumored, but little ever confirmed
about what exactly occurred when the Simons later conducted raids into
labs believed to be housing unaccounted for Lightbearers in Shaper
space. It is well known that the Shapers were absolutely livid over
the Lightbearers' use of human slaves, and it is suspected that a far
darker hell awaited those who fled to Shaper space than they expected.

Notably, the Shapers never gave any indication that they were incensed
by the detailed descriptions of what the Lightbearers had subjected
the Shmrn to, and it is generally conjectured that they were privy to
that information long prior to its distribution by the
Andolians. Apparently, giving the Shmrn a strongly defined sense of
self, dignity, modesty, and a mind muddied by the introduction of
human derived genetic codes for brain development, all in the name of
allowing the Shmrn to truly appreciate and understand the physical and
emotional torture and suffering being inflicted upon them by the
Lightbearers in a way that the Lightbearers themselves could
understand, was not cause for outrage in the Shaper population because
the Shmrn weren't human. This lack of response, though attributed by
some as merely arising from their part in creating the Shmrn, led many
to believe that, if and when the Shapers considered themselves or the
rest of humanity to no longer be human, similar indifference could be
expected to follow were anything to come to affect the rest of
humanity differently than the Shapers. The topic of rights for
associated non-human species has continued to be a topic of contention
between Confederation factions, with the Andolians and the Shapers
tending to lead very different camps in such discussions. Likewise,
even on the topic of human rights, the different principles held by
the Andolians and Shapers has led to rather different stances being
taken on a number of policy decisions. On the topic of human slavery,
however, the two groups stand united, a unity that was formative in
the early Confederation.


\subsubsection{Discussion on The Fraternal War}
Swright

Ostensibly The Andolians' empirical eminence and their considerable
authority in The Confederation was largely established by an effective
campaign against the Light-Bearers and an even more effective
political maneuver with regards to both the Spaceborn and the
Klk'k. Not only did the injustices perpetrated against the Spaceborn
dissuade The Shapers from opposing the war, but the tradegy on Ktah,
namely the Hoshino Uprising and the attempted genocide of the Klk'k
would have brought sharp opposition against the Light-Bearers from
other meme-groups as well. Further, the Andolians' successful efforts
to ressurect Ktah and their continued protection of The Klk'k would
serve to advance the group's political clout if not morally, at least
by proxy from the gratitude and loyalty of those whom they liberated.

JackS
	
Worth noting is that a significant impetus in the formation of the
Confederation was the fear of the Andolians shared by the other
meme-groups (and the belief that including them in the Confederation
was the most effective way to constrain them from future similar
actions). Given the systematic destruction of billions of
Light-Bearers, the restoration efforts on Ktah did little to convince
anyone of the Andolians having a particular advantage in moral
authority. While the attempted xenocide on Ktah and the exposure of
the Space-born did much to remove any support for the Light-Bearers
and their actions, neither served to significantly strengthen the
external perception of the Andolians.

Swright

Following the war, the remaining Light-Bearers were taken into custody
and turned over to elements of The Spaceborn, The Klk'k, and The
Shmrn, the three groups being those who had suffered injustices
enacted by the former, though it is believed that some may have found
sanctuary in systems controlled by The Shapers.

JackS

I believe that forced sterilization programs were also mentioned in
existing documentation.

Swright

Are there any Light-Bearers remaining at the 'current' time? If so,
are they pure-bred suprahomo, or has their genetic line become
diluted, perhaps with superioris genes? (Certainly those Light-Bearers
which may have survived the war in Shaper custody would have found
their hosts to be sympathetic to any desire to preserve a carefully
engineered genetic line.) Further, if they have survived, what level
of freedom do they have in current society? One would not think that
the children of war criminals would be punished or detained due to the
sins of their fathers, though they certainly might find themselves
victims of unending scrutiny by organizations such as Homeland
Security's IntelSec.

JackS
	
Members of the Light-Bearer meme-group: yes, but extremely few within
the Confederation, and none publicly, as any hint of the Light-Bearer
meme tends to cause members of the Simons to conduct a hunt. Those
fleeing into unexplored space rarely stayed ahead of future human
expansion and were hunted diligently by the Simons. Those few who fled
off of the jump network entirely gained a longer respite, but the
practical limitations of such journeys meant that all but the most
well-hidden outposts could only be temporary homes -- anything
reachable by a defeated and fleeing group via SPEC would clearly soon
become reachable by those searching for them, especially with greater
resources behind them.

Descendents of Light-Bearers: yes, but only in small numbers, and
generally integrated into other societies, if not necessarily into
other gene pools. As the Andolian view of the war was not a war
against a polity or a people, but a culture, ideology, and general
worldview ( i.e. a set of ideas and beliefs) that the Andolians
believed intrinsically diseased, given the behaviors undertaken by
their adherents, in the most general case, only those young enough to
be deemed viable candidates for successful re-education were even
given an opportunity for a post-internment existence. After dispersion
among many worlds, the total population of living Suprahomo is quite
small, especially pure-strains. Existing populations are generally
ideologically integrated into their surroundings, and, while subject
to surveillance, are not otherwise infringed upon provided no contact
with Light-Bearers has been observed.

Shaper custody: As mentioned in the description of the Shaper faction,
they find the concept of human slavery fundamentally abhorrent to
their ideals of personal freedoms. Thus, while those Light-Bearers who
fled to Shaper space may have faced a less dire fate than those
remaining to face the Andolians, they were not welcomed with open
arms. Precise details on the number of Light-Bearers who submitted
themselves to Shaper control have never been made available, and it is
generally believed that only those Light-Bearers with useful skills or
information were well treated in Shaper custody. Within a few months
of the end of the war, only limited signs of the Light-Bearer influx
could be seen, and the number of Light-Bearers presented by the
Shapers to the Shmrn, Space-born, and Klk'k delegations for oversight
by said aggrieved parties is widely believed to be much smaller than
the number arriving within Shaper space.

Swright
	
What did/do they look like?  Are they distinquishable from the homo
sapiens sapiens, or remarkably differing in appearance like The
Shapers?  Other than their ideological desire to be seated at the head
of the evolutionary table, what can be said of their culture?

JackS

Being supra- rather than super- human in nature, all of the genes in
the Lightbearer genepool were of human origin, but the mix doesn't
reflect the configuration of any particular historical group. An
individual Lightbearer can, unlike the Shapers, blend in (to a degree)
with any sufficiently diverse population of homo sapiens sapiens,
although groups of them would be conspicuous, and there would likely
be clear suspicions of heritage. Differences would be most acute at
age extrema -- suprahomo tend heavily toward earlier growth spurts,
and age very gracefully compared to sapiens. Suprahomo skin tends to
be devoid of any visible blemishes and only memorable injuries tend to
leave visible scars. Suprahomo do not suffer from hair loss, hair
discoloration, or other similar aging effects stemming directly from
isolated genetic traits.


Lightbearers stood 1.9 - 2.3 meters in height, and had universally
muscular builds and bulk scaled to their size. Skin tones, while given
to some variation, were heavily centered in bronzes, such as can be
found in Mediterranean-Persian hues. Hair color varied widely, but the
hair itself was always straight and frequently worn long. Lightbearer
men traditionally wore full beards, again of arbitrary color,
frequently of no correspondence to hair color, but, unlike their hair,
tightly curled, and kept close and neat.

One of the keys in understanding the Lightbearers was that they didn't
have a {\em desire} to be seated at the head of the evolutionary table,
but a steadfast, axiomatic belief that humans already were, by
definition, at the head of the evolutionary table, and that all that
was necessary was a perfection of humanity. The Lightbearers believed
universally in a creator deity with specific interest in humanity, and
in a literal interpretation of humanity being created in said deity's
image. This literal view was central to the negative reaction of the
Lightbearers to the Klk'k, as the only (vaguely) anthropomorphic
sentients discovered.

At all levels of Lightbearer society, the hallmark of their culture
was confidence. While in retrospect, hubris may seem the more
appropriate term, for the Lightbearers, their place in the cosmos made
them feel entirely secure in their sense of selves. Lightbearers, in
attempting to be the ultimate expression of humanity, developed what
many other contemporary cultures considered an indulgent hedonistic
streak. Excesses in all manner and direction were not uncommon, but
the excesses themselves were not of uncommon origin -- the
Lightbearers did little that others did not, but they did it moreso,
both "good" and "bad".

Swright

What level of preeminence did they have among the other human
meme-groups before The Fraternal War? Obviously their
military-industrial complex was weaker or at least less prepared than
The Andolians', but how did it compare to the other human factions?

JackS

The Lightbearers quickly shot to power in the early human FTL era
through aggressively expansionist growth and exploration policies and
favorable starting conditions, having weathered the nano-plague in
relative stability. While their relative power was waning even before
their destruction, especially as the Shapers became increasingly more
effective in adapting themselves to inhospitable environments in the
post-SPEC age, the Lightbearers were, if not second to the Andolians,
then sufficiently close to other contenders as to be considered tied
for "second" in their military output leading up to the war. However,
as all space navies were quite diminutive prior to the Fraternal War,
it is difficult to draw much from such comparisons. The Andolians were
particularly advantaged, not just by their industrial capacity, but by
the rapidity with which they were able to reorient total capacity
toward military ends.

Swright

Are there any particular artstyle comments about Light-Bearer
technology? Further, do any Light-Bearer equipment, bases, or vessels
still exist? If so, in what state or condition might these be? (Two
hundred years is a long time, but in the vacuum of space it is highly
conceivable that technological elements of the faction, whether fully
or partially operational, might remain.)

JackS

The Light-Bearers were capable, if not always masters of every art
pursued, and their constructions captured both their solid workmanship
and reverence for the aesthetic. Light-Bearer ships were unnecessarily
pretty, but such beauty was, pragmatically, only skin deep, and any
damage would reveal that the artisans working on the exterior had been
working on an exterior separate and almost distinct from the craft or
station underneath.

Precious few Light-Bearer vessels still exist, but some of their
outposts, both those constructed pre-defeat and those constructed
post-flight still exist in under-inhabited and remote systems. Many
have been directly assaulted, but others merely pillaged, relocated,
or otherwise repurposed by pirates or other tertiary entities. Several
stations abandoned or surrendered, rather than defeated, are still in
use by one entity or another, and others have become the nearly
unrecognizable centers of subsequent colonizations.

Swright

How long did the war last? Did it 'start' on Ktah? (The timeline might
lead one to believe The Andolians' impetus for military action against
fellow humans may have stemmed from The Light-Bearers' obdurate
slaughter of a sentient race, though in all likelihood there were
other less noble factors instigating hostilities between the two
groups.) How did the Shmrn respond during the conflict? And After? How
large was the Light-Bearer population, and how many colonies, bases,
etc? What was their capital system and/or world?

JackS

The Shmrn had little response during the conflict, not being well
equipped in either information or supplies for any sort of slave
resistance. Many Shmrn died alongside their Light-Bearer masters in
the planetary bombardments. They were afterwards pleased to be freed
from cruelty at the hands of the Light-Bearers, but not particularly
thankful to the humans in general (although they hold no particular
grudge against the Andolians for those killed in the bombings). Shmrn
general opinion of humans is frequently dim, but not antagonistic --
they are more cynical than angry.

The Light-bearers began the war with a population of ~20 billion
spread across 11 major colony systems and with outposts in another 2
dozen or so systems, the whole of which was ruled from a world called
(thanks here to Oblivion) "Cradle".

The war did indeed start on Ktah, and lasted for a few years, but in 3
distinct phases. Preliminary activity centered on Ktah, with assorted
skirmishes elsewhere in uninhabited or newly inhabited areas. Without
existing infrastructure in the system, and with a lengthy resupply
route, a significant portion of the Andolian fleet had to be kept at
and along the route to Ktah to support the safety of the Ktah
restoration efforts. Once the situation in space and on the ground at
Ktah stabilized, the war moved in a new direction, with Andolian
forces taking much more aggressive stances, and heavily raiding
Light-Bearer colonies and pushing the lines of battle back into
Light-Bearer space. The political climate among the other human
meme-groups became much more volatile at this point, as what had
previously seemed, if unpleasantly violent, something that could be
rationalized as merely a logical extension of the previous era's
border conflicts was turning into something more disturbing. However,
it was at this juncture that the Space-Born were revealed by
successful raiding parties, thus considerably disrupting the political
will of third parties for mediation. Throughout this phase of the war,
the Andolians seemed to be winning, so to speak, but only by
attrition. However, the Andolians were in fact committing only those
forces necessary to continue to make minimal progress while holding
back (and physically concealing) the rest of their production (greatly
misjudged in quantity by Light-Bearer intelligence when on full
war-time resource allocation), especially of next-generation vessels,
for a reserve strike force. The third, and final phase of the war came
swiftly (and not far ahead of renewed stirrings of calls for peace),
with massive Andolian strike fleets taking Light-Bearer defenders by
surprise in simultaneous attacks on five of their core worlds. Further
complicating matters for the defenders was the supposition that, as
had happened with Ktah, the Andolians would invest significant time
and resources in asserting control over a populated planet. Instead,
the Andolians mostly bypassed the planets after securing orbital
superiority, completing initial planetary bombardment, and beginning
kinetic bombardment operations. The infrastructure of the Light-Bearer
core worlds was destroyed in a matter of days, as the Andolian fleets
marched relentlessly on. The smaller Light-Bearer outpost worlds
actually fared better, as they were deemed small enough to manage
through more traditional means.

The planetary meteor bombardments were unexpected in scope, and
utterly devastating, lasting, in the case of Cradle, the most
developed of the colonies, continuously for well over two
weeks. Bombardments did not cease until both sufficient debris was
present in the atmosphere to make unsupported existence on the surface
short-term untenable and all detectable planetary infrastructure had
been damaged beyond likely ability to provide support. Third party
media, when reporting on the aftereffects, would come to call the
practice "Dusting a World" and frequent comparisons were made to the
salting of Carthage, among many other more horrific comparisons,
depending upon the political alignment of the media in question.

Swright

Was this engagement premeditated by The Andolians, or precipitated by
the atrocities of the defunct group? (Certainly it would seem the
former, as the devastating effectiveness of the anti-Light-Bearer
campaign would suggest.) Was there any diplomatic response by
non-human powers? (Specifically, the Rlaan, Uln, or Aera.)

JackS

The Andolians had been preparing for some form of more intense
military action as a contingency ever since the invention of the SPEC
drive, but had no particular plans to engage the Lightbearers until
they were witness to the Lightbearer occupation of Ktah. In
particular, they did not deem conflict, which would no doubt be very
costly in both lives and materials, necessary when expansion was still
A) proceeding in an unhindered fashion, and B) going well for them
relative to other powers.
 
The Rlaan, Uln, and Aera had not yet been encountered at the time of
the Fraternal war.

Excerpt from Music Brainstorm: HUMANITY The Fraternal War would result
in the Lightbearer memes becoming defunct, the liberation of the
Klk'k, Shmrn, and Spaceborn, and, as a reaction to Andolian military
dominance and the totality of Lightbearer destruction in the conflict,
the stirrings of what would become the Confederation of Inhabited
Worlds. It was not lost on the other major human groups that the
Andolians were unapologetically responsible for killing more humans
than any other group in history. A greater chance of survival through
a more structured means of inter-meme interaction than armed conflict
was a sufficiently common goal to bring together the, by this time
extremely ideologically divided, human polities, but not sufficient to
bind them more than loosely. The polity governing the dispossessed
settlers, for reasons of jump network topology and lingering distrust
and animosity did not join the Confederation.

The Fraternal War also marked the beginnings of rampantly increased
militarization of budgets throughout human space. This tangible
symptom of unease and distrust was only accelerated by contact with
the Rlaan, Uln, and later the Aera, in the two centuries following The
Fraternal War. This served to slow the pace of economic and physical
expansion somewhat, but the intense R\&D efforts were not without
benefits, and contact with the Rlaan, Aera, or even Uln, would likely
have been quite a different underatking for a humanity not devoting
itself to unkindly causes.

\section{Upon the Coldest Sea}
\label{exptimeline:UtCS}
Act Titles, Quotes, and Big-Picture Events in

{\bf Upon the Coldest Sea }

\subsection{Act 1: Hiroshima (mon ami)}
\begin{center}
{\it Give me Christ or give me Hiroshima.}

Leonard Cohen - The Future
\end{center}

The Aera have invaded the Union of Dispossessed Settlers seeking to
carve a path through human space to the coreward regions of the jump
network. The Confederation of Inhabited Worlds lends increasingly
direct aid while the Forsaken are driven further and further back in
what are clearly little more than delaying actions against the far
superior Aeran forces.

\subsection{Act 2: As the sea begins to free them}

\begin{center}
{\it Mourn, England, mourn and complain

For the brave Lord Nelson's men

That died upon the main}
\end{center}


An official state of war is reached between the Confederation and the
Aeran Ascendancy. The Aeran-Human border is ablaze with combat, and
many Human border worlds are lost or compromised as the Aera strike
first. Shmrn space is invaded, and the Andolian 9th fleet, sent to aid
the Shmrn, is cut off. Victories by the Andolian 6th fleet, and
Confederation 4th fleet blunt the Aeran advance on the Spinward front,
but at significant cost in men and materiel. Aeran forces penetrate
deeply on the center front, while much of the anti-spinward front is
pushed to near collapse, save for pockets of bypassed Forsaken
worlds. Confederation forces resort to scorched earth policies in many
systems still belonging to the LIHW or Union of Dispossessed Settlers,
completely abandoning the Diaspora sector, and regroup in more
sustainable positions. Confederation fleets engage the Aerans in
skirmishes and strikes throughout the center front, with profound
carnage on both sides. Andolian counter-attacks all but eject the Aera
from Shmrn space, reestablishing direct communications with the
Uln. Andolian forces begin a methodical incursion into Aeran space on
the spinward front, relieving some pressure from the anti-spinward
front as the Aera redistribute their forces. The Rlaan watch intently,
hoping to be able to avoid intervening, while putting significant
effort into expanding their corward possessions while humanity is busy
fighting the Aera.

\subsection{Act 3:  Children of an unconstructed god}

\begin{center}
{\it Gaudete, gaudete! Liber est natus ex machina sapiens, gaudete!

Gaudete, gaudete! Liber est natus ex post-homo paternus, gaudete!}
\end{center}

The Andolians and the A.I. Quorum unveil the Grandchildren, the fruit
of decades of research and material investment. A new class of
thinking machines, designed explicitly with military application in
mind (indeed, it is generally assumed by all other parties that they
were originally designed for use against or for leverage with the
other human polities), vast multitudes of Grandchildren, in the form
of a veritable armada, issue forth from inside the hollows of
Hephaestus, and begin an all out invasion of Aeran space. Taking
advantage of wartime powers granted by the Confederation Senate in
unintended fashion, the Andolians allow their Special Forces
operatives to wage an unrestricted campaign against the Interstellar
Church of True Form's Return without fear of Purist legal
entanglement. The major Confederation powers play kingmaker among the
less than reputable and less than legal organizations, bringing in
from the cold those entities willing to assist the war effort, and
removing from existence any competitors less eager to be of service,
such as the Order of the Dynast Shrub. Some of these actions deeply
anger various Ulnish clans, but the proximity of several fleets to Uln
space deters overt action, and retribution takes the form of Uln
relaying of intelligence to the Aera. The Simons are employed in
overthrowing the corrupt, and externally funded leadership of the ISO
as a precursor to instigating coordinated guerilla actions from the
pockets of Forsaken worlds behind the Aeran lines. Shaper forces make
their first real impact of the war, as troopships full of Shaper Hulks
are dispatched alongside Mechanist and Purth forces to reclaim worlds
fallen to the Aera. The initial engagements by the Grandchildren are
stunning victories, greatly demoralizing the Aeran forces, and
disrupting their efforts to concentrate their fleets.

\subsection{Act 4: Fallout - Ecce Homo Post Diluvium}

\begin{center}
{\it ... [it] will not start a chain-reaction in the water converting it all to gas and letting the ships on all the oceans drop down to the bottom. It will not blow out the bottom of the sea and let all the water run down the hole. It will not destroy gravity ... } 

Admiral William Blandy on Ivy Mike
\end{center}

As the Grandchildren produce an increasingly impressive set of
victories from the Spinward front and ever deeper into Aeran space,
the progress on the other fronts is much slower. Aeran raiding parties
continue to harass and impede. The Forsaken guerilla attacks, while
marvelous as a delaying tactic, can no longer be sustained, as the
Aeran response has destroyed all remaining Forsaken colonies behind
the front. Especially troubling, an unexpected Confederation defeat,
due more to distrust and negligent communication than individual
incompetence, coupled with the deployment of the Ascendancy's Leonidas
class dreadnaught reserves, has opened up Vega sector to Aeran
assault, and many long developed worlds have been raided, threatened,
and attacked. With the devastation being wrought by the Grandchildren,
the Aera are increasingly desparate to either lure the Confederation
into ill-planned action via destruction of ancestral worlds, or to
construct a deep enough corridor that a spread of colony convoys may
be launched in a last-ditch attempt to bypass Human space primarily
via SPEC, retreading the waters that started the Rlaan-Aeran
war. Fighting on the front becomes especially fierce, as both sides
commit themselves deeply to each fray. Domestic Confederation politics
boils and froths as assorted scandals are rooted out when entities are
no longer capable of focusing the attention necessary to hide them,
and the Andolian Protectorate woos the Shapers into their fold.

\subsection{Act 5: Sailing on Embers}

\begin{center}
{\it And all that remains is the faces and the names of the wives and the sons and the daughters}
	
Gordon Lightfoot - The Wreck of the Edmund Fitzgerald
\end{center}

With the Grandchildren winning some key decisive victories in Aeran
space, and the incursion into Vega Sector blunted, the long term
outcome for the war seems increasingly likely to be an Aeran
defeat. However, the cataclysms of combat have quieted somewhat as
both sides have exhausted themselves, and any final closure is surely
years, if not decades away, leaving much time for potential
reversals. The Rlaan are the obvious short-term beneficiaries, but it
is clear that they are somewhat uneasy about the Grandchildren, having
been taken by surprise by the revelation of their existence. On the
domestic front, the Confederation is undergoing profound political
upheavals. The Highborn have been cowed into silent disgrace as a slew
of rumours about their involvement in illegal activities in Forsaken
space issues forth following an orchestrated series of information
leaks. With the Grandchildren firmly in their camp, the Andolians have
finally found the trump card in their long contest with the Shapers as
to who would helm the march to post-humanism. With the Shapers now
aligned with the Andolians and their traditional allies, the
post-human agenda now dominates the policy-making bodies of the
Confederation. The Forsaken, once again dispossessed, are scheduled to
once again be relocated, but this time will remain under the auspices
of the Confederation, with semi-autonomous status, sharing in the
opportunities available to other humans. As humanity adjusts to both
life during wartime, and the possibility that humanity's children may
be much further on their way to surpassing them than expected, the
first trickle of Aera who have come to believe that the future of
their species may rest in siding with humanity lest they be erased by
it journey to Uln space to conduct the first of a series of
clandestine meetings with Andolian and, later, Confederation agents.

\subsection{Black Paralysis}

\section{Foundering Inheritance (Orion region)}
One of the key points about the Foundering Inheritance era mentioned
in the conversation concerning the overall VS universe timeline is
that the Rlaan do not realize they are encountering a new alien
entity. The destructive force that renders the domain of the Rlaan
Assembly broken and the Saahasayaay nearly extinct is first
encountered as the local variant of the nano-plague on the other side
of the locked gate (the expidition of the 10,000 finds no local
nano-plague, as the Ancients in that region had been supporters of the
status quo) in the Saahasayaay home system. As the variation between
observed nano-plague entities in locals space was greater than that
between the two spaces, seeing more of the ubiquitous nano-plague on
the other side of the locked gate was seen as mildly interesting, but
generally unremarkable. It may have even been somewhat unremarkable,
save for the other-side nano-plague recognizing the symbiotes in the
Saahasayaay (see wiki description of Saahasayaay life cycle) and
activating in a decidedly organized fashion. It is no coincidence that
the habitable worlds of the jump network are disproportionately suited
to oxygen breathers, and it is no coincidence that near the
Saahasayaay system there are several other chlorine worlds, and near
the Rlaan homeworld, several methane-ammonia systems, and another
locked gate. These areas were provided for neighboring Ancient groups,
resident across the locked-gate boundaries. The symbiote in the
Saahasayaay was a tool of one such neighbor, and the other-side
nano-plague, another tool, recognized its kin, even in its kin's
damaged state. The other-side nano-plage effected changes to fix the
symbiotes, allowing the other-side nano-plague (hereafter OSNP for
brevity) to use the Saahasayaay as a conduit for information, and then
for dissemination of the symbiote fixes back to the Saahasayaay
homeworld. The eventual goal of both OSNP and symbiote was simple, if
vain and likely futile: to prepare the way for a return of their
long-dead Ancient creators, existing currently only as dormant
information. While the OSNP increased its activity, building
progressively larger colonies, the colonies macrostructures, and the
macrostructures becoming macroscopic machines themselves, it remained
resource limited - the immediate region of several star systems was
quite barren, strip-mined in its hey-day during the Ancient-TWHON war,
and then ravaged in the aftermath, and it would be some time before
the OSNP would be able to produce drone ships either as an end product
or to transport goods from richer systems. Neither could the OSNP act
directly in our region, as our own nano-plague would react negatively
to it if it were to try to expand. Instead, the erstwhile-Saahasayaay
(not truly Saahasayaay anymore, but merely avatars for their finally
flourishing symbiotes) staged a revolt, taking numerous Rlaan
civilians as hostages, knowing this would delay the Rlaan response
long enough for their ships, now crossing freely into the other-side
to be upgraded by the OSNP - if not yet to anything resembling Ancient
standards, then enough to be a great surprise to the Rlaan when the
time for combat was to finally come.

Bargaining, in a very uncharacteristic way for Saahasayaay (this
itself already greatly disturbing the Rlaan), for the safety of the
Saahasayaay outside of the home system with the lives of the Rlaan
Workers, the ex-Saahasayaay gained both more crucial time and powerful
weapon: the Rlaan Workers were themselves modified by the OSNP to have
an innate desire to sabotage, kill, and disrupt - while the Rlaan
would check them for bio-weapons and such, the Rlaan mindset precludes
consideration of Workers as themselves being instruments of violence,
and so clean bills of health and belongings would see them released
back into Rlaan society. In a move reminiscent of the battle of
Puebla, the Rlaan forces charged directly through the jump-gate and
the inevitable fixed defenses, rather than flanking with SPEC. The
Rlaan suppression fleet met with horrendous casualties at the hands of
the upgraded Saahasayaay vessels, and, the first glimmer that more
than just a Saahasayaay ploy to escape Rlaan rule via the lock-gate
was underway, upgraded vessels of remnants of the Rlaan expeditionary
force itself (there having been some resistance in the Saahasayaay
overtaking of the expeditionary force). Though unsuccessful in
retaking the system, the attempt had bought the Rlaan valuable time,
though they did not then know this, for the OSNP ramp-up would be
significantly slowed by the damage to many of the already quite small
number of vessels at the command of the OSNP and
ex-Saahasayaay. Nonetheless, while the Rlaan re-inforced and
re-grouped, planning to send an even larger force through, this time
via SPEC, this remained a "local" issue, with news of the Saahasayaay
rebellion being an event not shared with the human powers. However,
when, shortly after returning to Rlaan space, the modified Rlaan
Workers set about their assigned terrorist tasks, the entirety of
Rlaan space was overtaken with hysteria and paranoia - Workers
committing violent acts of terror, even if the actual damage was so
small as to be trifling, and most of the Workers in question perished
in the execution of their tasks, was so antithetical to the Rlaan
worldview of the possible that it threw their entire society into
gridlocked turmoil. This was to prove fatal for long term Rlaan
aspirations, and to be a saving grace for all other species in the
region, as it was a disaster now too large to keep quiet. Fleet
movements that could have all directed themselves in a bloody, but
likely ultimately successful assault on the OSNP beachhead, were
curtailed, as the Rlaan, in paranoid meltdown, saw threats everywhere
that the released Workers could still be, and held back significant
fleet resources for use with the Rlaan Enforcers to retrace the paths
and hunt down every one of the "abominations."

When the second attempt to retake the Saahasaay homeworld and the
lock-gate failed (even more poorly than the first, thanks to new waves
of ships of unknown design), the fact that the situation was
exceptionally dangerous became increasingly apparent to all
parties. Joint efforts were undertaken to organize an interspecies
response. As these efforts began, the first counter-attacks began to
issue forth from the lock-gate. As time wore on, the enemy ships
improved in design rapidly, if only slowly in number. Analysis of
probable causes leads to the the gate in (formerly) Forsaken space
from which the fleet of 10,000 left being locked from both sides,
isolating both fleet and research colony from the rest of known
space. Strikes are radiating further and further out from the
lock-gate, battles are slowly worsening in per-ship losses, and the
combined forces of the extant species are being repeatedly forced to
withdraw to keep from being overrun. The situation is grim,especially
as it is obvious that time is not on the side of the newer species. It
is only by becoming much grimmer that the situation will improve.

As the first wave of nearly-Ancient quality drone craft is completed
and larger craft construction commences and continues, it is launched
en-masse in a series of attacks against the Rlaan, both fleets and
worlds, carving a vast barren region into the Rlaan Assembly's
territory. However, this coordinated use of Ancient tech stirs TWHON
fragments slumbering in the other-side, and the resolution of this
conflict is set in motion. A single small TWHON fragment plows
straight through into our region, bypassing the unprepared OSNP, and
engages the drone craft. Though the fragment is destroyed, the drone
craft are decimated in number and universally damaged. All OSNP and
ex-Saahasayaay forces retreat to an expanded beachhead area in the
systems near the lock-gate, the drones being recalled for repair and
re-deployment against the stirring TWHON fragments in the OSNP's own
region. Meanwhile, the presence of a TWHON fragment in our own region
fully activates our own nano-plague (it already having a many hundreds
of years head-start on the OSNP due to our own FTL
travelling). Nano-plague factories begin defensive preparations
against subsequent TWHON incursion. Reconnaissance of the beachhead
proves difficult, and recon through the lock-gate, nearly
impossible. However, it becomes clear that the preparations being made
at the beach-head are for an OSNP/ex-Saahasayaay escape from their
region, and no longer directly a pacification of ours - the OSNP are
losing. Though drone-craft from our region have begun to appear in
number, they do not cross over into the next region to assist - they
still obey the long meaningless borders between their long dead
masters.

With the evidence of a distinct drop in craft from the beach-head area
signalling a likely imminent collapse, the joint Human, Grandchild,
Klk'k, Rlaan, Aera forces embark upon a massive suicide mission of
genocide - having realized that the OSNP cannot operate freely in our
region except through the ex-Saahasayaay, and that they could not hope
to defeat the OSNP forces in direct engagement, they instead chose to
exterminate the Saahasayaay hosts to the Ancient-designed symbiotes
and as much of any construction infrastructure as they could
manage. The strike was well-timed, and the tactic not one the
ex-Saahasayaay had weighted with high probability, having been too
colored in their view by lingering effects of their time in Rlaan
service. The Saahasayaay homeworld was carpeted by anti-matter bombs
even as the fleet delivering them was being savaged by OSNP drones. It
was unclear exactly how successful the mission had been - very few
craft returned, and there were too many unknowns as to other potential
bootstrapping sources for the OSNP. However, only days after the raid,
TWHON fragments broke through the locked gate, signalling the totality
of the OSNP demise in their own region, and the entire Saahasayaay
home system, when it was next visited, was found to be
slagged. Whether through previous TWHON attrition, or superior
preparation, our own dones were faring somewhat better against the
TWHON fragments. Also remarkable was that our nano-plague took to
using jump-point deactivation as a time-gaining tactic in fighting the
TWHON fragments, which either could not or would not be bothered to
reactivate the jump points and instead would travel through some more
advanced SPEC-like manner directly between stars in the absence of
jump points. Though the TWHON-nano-plague fight raged, the priority of
the extant species remained the re-locking of the locked gate. The
expedition charged with doing so was assembled as quickly as possible
given the extreme losses all parties had already endured, and expected
to endure again.

Upon reaching the, now disconnected from the jump network, former
Saahasayaay home system, they found no fleet waiting for them,
instead, they found a single, small TWHON fragment body and a series
of slagged worlds. Much to their surprise, the TWHON body completely
ignored them as long as they remained distant from the lock-gate. When
local drones engaged the TWHON fragment and it chased the drones out
of the system, the expedition set up to lock the gate and retreated.

Though the few TWHON fragments that entered our region were eventually
defeated (the long-active nano-plague prevented local procurement of
re-inforcements, that being the true key point of the nano-plague, the
drones being entirely secondary, as Ancient fleets were assumed to
exist at the nano-plague's creation) the TWHON-drone conflict made the
area around the lock-gate unfit for habitation both through further
destruction of worlds and severing of jump-links.

At incident's end, the Rlaan were no longer the power they were, with
many worlds destroyed, many more damaged, their fleet broken, and the
breach of trust that started the incident a sticking point that would
be quite well remembered in the coming period of rebuilding - one they
would pay dearly for in control over their future, and one which would
lead the Rlaan-Briin to societal dominance with the support of the
Post-Human powers. Moreover, there was now a great pallor over the
futures of all of the extant species, as it became clear that this
could not be a lasting defeat of the TWHON - rather, the very stirring
of our nano-plague, it's true purposes at last revealed, would
undoubtedly bring more suffering in the future. The sword of Damocles
would now rest over all of their heads, and it would clock-like sound
each hour to pass of doom creeping forward while salvation remains
nebulous.



d: simply that the 'cleaning up' of the whole mess seemed "easy"
(22:28:20) d: as in, besides the space fleets, the only real losers from the whole lot were the rlaan and the s*
(22:29:01) d: which didn't quite match what I thought led to teh later timeline you talked about
(22:30:07) d: but I get now that it was the long term events set in motion (ie the activation of the local nano plague) are what sets ruin to teh current crop of spacefaring races 
(22:30:42) d: but don't let that comment mislead you, melikes :-)
(23:06:05) t: yeah, the Post-Human powers didn't get their turf destroyed
(23:06:13) t: but their fleets were shredded
(23:06:19) t: I should make that more explicit
(23:06:44) t: most of the ships involved in the late action, even defending the Rlaan, weren't Rlaan anymore 
(23:07:05) t: there just weren't many of them left
(23:08:23) t: The sweep made by the OSNP drone craft was a bit understated
(23:10:00) t: We're talking, Rlaan needing to ask for help to clean up all the dead Rlaan bodies understated
(23:11:02) t: And the suicide run.... would be the largest congregation of extant species ships seen in a long time
(23:14:03) t: especially with a hyper-active nano-plague playing serious hell with production (bad tendency to eat some high-quality materials) although that would die down soon enough 
(23:14:39) d: cooly, that's some better sense of proportion ;-)
(23:17:23) t: that, and they weren't able to confirm that they killed all of the ex-Saahasayaay
(23:17:42) t: minor potential problem, that, over the next few thousand years
(23:18:02) d: as in, point for focus of paranoid delusions?
(23:18:24) t: that, and just because you're paranoid doesn't mean they aren't out to get you
(23:18:27) t: :)
(23:18:37) d: :-)
(23:18:41) t: closing the gate was intentionally anti-climactic
(23:21:23) t: it deprives the survivors of the solace of accomplishment, false or otherwise
(23:21:55) t: forcing them to face the unsettling reality that, at most, they've only delayed the inevitable
(23:22:22) d: point
(23:22:45) t: which, I think, is a nice note to end that chapter on :)
(23:22:59) t: condensed in form as it may be

One of the things that I should have mentioned more explicitly is that
the above was only a description of what was going on in Rlaan space
during this period. The events in Human/Post-Human/former-Aeran space
are fairly important too, and brought all the more to a head because
of the fall of the Rlaan - it is worth remembering that a key reason
that warfare was fought in a "restricted"
i.e. non-genocidal/collateral-damage-minimizing fashion in the UTCS
era was the Rlaan intolerance for civilian body counts (so if you were
going to butcher planetary populations, you had to keep it on the
serious down-low). With the Rlaan gone, there's nothing but human
conscience to prevent atrocity (which is not a historically viable
solution). Add to this the tensions between human (Basically Purists
and allied factions - the Forsaken still remaining distant from the
affairs of the Confederation worlds) and post-human (The Andolian
Protectorate (having expanded (via negotiated annexation) by this
point to include the former Unadorned, Mechanist, and Shaper polities)
and the Grandchildren) political groupings in what has become of the
Confederation and the destruction of the vast majority of the
Confederation and its constituents' military forces (worth noting that
the majority of the ships on the "suicide run" on the Saahasayaay home
system were civilian vessels hastily armed as cannon fodder due to a
lack of remaining military forces), and you have a recipe for violent
unrest, even if the human worlds themselves were not
ravaged. Likewise, even before the actual destruction of the fleets,
there was significant political maneuvering occurring within
Human/Post-Human space concerning both the fleet of 10,000/other
expansionst projects and scheming by various groups to position
themselves to profit from what was initially seen as a minor mishap in
Rlaan space.

To briefly summarize where things go from here: 

Sensing imminent Protectorate weakness, the Shapers plot and execute a
revolt. With internal stress between the post-human and human elements
of the Confederation already near boiling, the Shapers hope to ignite
a civil war with their revolt, believing that they will come out on
top when all the dust has settled and no other group remains to stand
against them.

Preliminary movements of the Shaper revolt involve prodigious use of
bio-weapons, including morphologic virus packages aimed at Purist,
LIHW and other traditional human populations in Confederation
space. The Forsaken, the bulk of their population now in former Aeran
space, are not targeted due to extremely low population density. Chief
among these virus packages are ones that overwrite the germ-line DNA,
causing erstwhile human parents to either run sterile or produce
offspring with Shaper-designed genomes. Viral packages for the Klk'k
populations are more simply plague oriented, and, as with the human
packages, are designed for stealth and long incubation periods. The
Shapers have also been producing modified fungal strains that act as
factories for "Quickships," extremely fragile but extremely cheap
in-system pilotless craft. The Quickships require engines and
weaponry, but are otherwise selfcontained. As they are considered
totally expendable, great liberties are taken with the safety,
reliability, and sustainability of the weapons and engines produced
for the Quickships - Quickships would invariably die from radiation
and heat stresses in prolonged service, even if they were not engaged
in combat.

As their bio-weapons begin to activate, the Shapers actively revolt,
laying a surprise siege to Kubernan. However, their intelligence
gathering proves mortally flawed when their attack on Hephaestus is
insufficient to halt production of Grandchildren. The blockade of
Kubernan greatly reduces the flow of resources into Hephaestus, but
the failure of the Shapers to break the Grandchildren's homeworld
means that they cannot readily press home their intended beheading of
the Andolian Protectorate.

The Andolian-led response of the Protectorate is swift and
brutal. Having actually made some preparations for a Shaper attack
back during the cold-near-war period prior to the Aeran war and the
unfurling of the Grandchildren, the signal is given to activate a
number of deep-space asteroid bases set up a few week's SPEC from each
of the core Shaper worlds. These bases will spend the next several
days accelerating at maximum thrust before activating their SPEC
drives and proceeding on a collision course.

Meanwhile, plague and the creatures dubbed "murder-babies" for their
habit of consuming their mothers just prior to expected birth, the
latter turning out to be an extremely rapid-growth form of Shaper
Hulk, inflict terror upon the human populations of the
Confederation. The Andolians, in their sterile compounds, and many
Mechanists, devoid of the necessary pathways for infection, are not
directly affected, although there are riots and violence in the
off-worlder sections of Andolian holdings, and the Unadorned will find
themselves devastated by the Shaper plagues. The Shapers embark on a
widespread propaganda campaign, and use the promise of cures and
vaccines as leverage to gain support from the human governments.
Several capitulate outright, and many more are overthrown by their own
populations, most, at this point, having long preferred to side with
the Shapers over the Andolian led Protectorate and their
emblematically non-human Grandchildren. The Shaper revolt is now a
full-fledged civil war.

After six centuries of existence, the Confederation is now dead in
name as well as fact. The Rlaan are powerless to do more than watch in
horror, and the Forsaken and Marked-Aera are hard pressed enough as it
is just trying to keep any potential plague-bearers out of their
space. The Uln bide their time, preferring to wait until one side is
closer to victory before considering any benefits possible from
backing one side or the other or pursuing territorial ambitions in the
resultant power vacuum.

With the bulk of human space-naval forces already having been
destroyed in efforts against the ex-Saahasayaay, the engagements seen
are quite different from those during the Aeran war. The Confederation
Joint fleet has completely disintegrated, with individual ships
falling to one side or another of the conflict, or simply returning to
the crew's system of origin. Small numbers of ships, many of them
conscripted civilian craft, or Regional Guard forces are often
deciding factors in individual battles. Capital engagements are only
seen around key worlds, and ground engagements are frequently fought
between small forces among swarms of ambiguously allied
civilians. Neither side engages in significant restraint, and the toll
in collateral damage is staggering.

While the Shaper assault on Andolian space gains little headway, and
the blockade on Kubernan is broken, the retreat of Protectorate forces
to defend their own territory has left the Shapers and their "allies"
a fairly free hand to attack and consolidate power in the rest of the
former Confederation. In a move that signifies to all with sufficient
understanding that the Shapers no longer consider mainline humans to
be members of the same species as themselves, the Shapers deploy
mind-altering parasites to ensure the submission of both willing and
protesting conquest alike. Throwing the resources, both material and
population, of their vassals into the fray as cannon fodder, the
Shapers are able to keep Kubernan and Hephaestus under constant threat
of attack, forcing the Grandchildren to be deployed defensively rather
than be allowed to engage in strikes against their own forces. With
the fall of Sol, Bantam, and New Camelot, the core region of
humanity's oldest colonies is almost entirely in Shaper hands, with
the Purists, ISMG, and High Born polities in collapse.

Attacking through space ostensibly controlled by the crippled Rlaan,
but in truth barely patrolled, the Shapers attempt to plague-bomb the
Shmrn worlds, releasing a virus package designed to turn future Shmrn
generations into Dgn. Given the widespread panic in the wake of the
war, countermeasures are in place, and the effect, as intended, is
more psychological than military. This act, however, pushes the
neighboring Uln to believe that neutrality would, in the end, be no
armor against a victorious Shaper entity, and the Uln throw in their
(small) lot with the Protectorate.

This would be the peak of Shaper aspirations for dominance. Coinciding
with the impact of the asteroid-base kinetic weapons on nearly
four-dozen Shaper homeworlds, the Protectorate launched a
counterattack spraying forth from Kubernan, spearheaded by a fresh
crop of Grandchildren produced at Hephaestus. The Protectorate strike
force handily defeated the vassal-ships surrounding the recent Shaper
conquests. However, they neither charged toward the Shaper worlds nor
attempted to liberate the conquered human populations. Instead, the
Protectorate had come to the conclusion that any world already
conquered by the Shapers was too polluted by Shaper bio-weapons,
traps, and incubators (human hosts being used to produce new Shapers),
and was safer cleansed than liberated. Using what would eventually be
revealed as a set of bio-weapons of Rlaan origin, the Protectorate
fleet embarked on their scorched earth and flesh campaign, destroying
every extraplanetary structure found, raining down orbital
bombardments upon every exposed spaceport and industrial area, and
leaving behind a cloud of extremely voracious organisms capable of
rapidly degrading planetary human populations to "manageable levels" -
especially given the Shaper efforts on every conquered world to make
sure their own bio-agents could have mass effect.

Certain that they could easily best the Shapers in direct engagements,
the Protectorate sought out existing and potential hosts within Shaper
reach and exterminated them with extreme prejudice. Both Shaper and
Protectorate sides had known from the beginning that there would be no
negotiated end to this conflict - the vast populations of the Purist
and LIHW holdings, however, had never anticipated this war, and yet
were the ones doing the bulk of the dying, being killed by the 10s and
eventually 100s of billions. The last contribution of the Unadorned
before succumbing entirely as a political and cultural entity, having
been devastated by the Shaper plagues, would ironically feature
heavily in accelerating the demise of mainstream humanity. Following
on from the Ancient library research that had led to knowledge about
the locking and unlocking of the far-gates, the Unadorned had
discerned how to temporarily disrupt normal jump points. The Culling
fleet, as it would come to be known, used this new ability both to
delay Shaper reinforcements and to protect their own rear. While the
disruptions lasted only a day or two, that was frequently more than
enough time for the fleet to crush the Shaper presence in a
system. The support of their vassals pulled out from under them, and a
third of their own planets now without an inhabitable surface due to
the impacts, the outcome of the war had been determined, and only the
shape of things to come remained to be decided.

The human population of the old Confederation no longer a potential
threat-in-the-making, Protectorate forces throughout the entirety of
their space rushed out from their defensive deployments and began
attacking relentlessly. The rush came from the perceived threat of the
Shapers fleeing out of the jump network, slinking off to where they
may not be found for centuries, an everpresent ghost of a menace,
leaving all the bloodshed for, in the end, naught in terms of
safety. However, owing to some degree of confidence in their eventual
victory, the cultural unity of the Andolians, and, perhaps moreso even
than the other two, to not having the AI Quorum cease its infiltration
of Shaper computing systems when the Shapers joined the Protectorate,
the Protectorate had a much better informed covert operations presence
in Shaper space than the Shapers in the Protectorate.

The Shapers had not begun work on their escape efforts until after the
Culling fleet's offensive began. Though the preparations were still
ongoing, the distributed nature of the Shaper efforts made it unlikely
that sabotage would sufficiently slow the progress to allow
Protectorate forces to reach all systems before an exodus could
occur. Rather, the Protectorate's covert operatives were employed in
gathering information as to the destinations of the Shaper
vessels. The seed fleets were allowed to leave, but would either find
themselves ambused when they came out of SPEC to resupply (the key
point being that, absent fuel, they would no longer be threats to
meaningfully flee the ambush), or, for the sublight "drifter" fleets,
chased down and exterminated. What few Shapers slipped through the
Protectorate's net would no longer constitute a people, a culture, or
anything other than stray scattered individuals, living out the rest
of their lives in exile. Historians would say of the war that, despite
the Shapers' superior intellects, which had led them on what should
have been more successful strategies, it was the superior Andolian
knowledge of the situation at hand that determined the outcome of the
war. A combination of ignorance and mild hubris (belief, not that just
that they could win, but that they deserved to, for being "better"
than their Andolian competitors) proved fatal for the Shapers.

However, before their demise played out, the Shapers did leave a few
more marks. The punitive plague-raid on the Uln (a simple killing
virus) was far superior in efficacy to that on the Shmrn, and the Uln
would prove notably cowed by the experience in their later
negotiations with the First Empire. Realizing that the bio-weapon used
by the Protectorate was of Rlaan origin, the Shapers incorrectly
concluded that the Rlaan had collaborated with the Protectorate in
engineering the Shaper demise, and likewise sent the Rlaan a "farewell
present" targeting their civilian population - less perhaps a parting
shot than a warning that the Rlaan would not be immune from the
Shapers' intended return. However, even in their shattered state, the
Rlaan were adept at producing counter-viral agents, and the plague
bombing was perceived more as an incomprehensible act of deranged
malice than an action of war. This act by the Shapers was to prove
fundamental to supporting the Protectorate arguments to the Rlaan,
utterly aghast, nauseated, and made profoundly uneasy by the
Protectorate tactics of the Culling fleet, that the Shapers were so
deranged as to leave euthanizing them as the only option. Indeed, this
act was so fundamental to the Protectorate arguments that there is
much speculation as to whether Protectorate operatives may have played
a role in convincing the Shapers to attack the Rlaan.

The Shapers sufficiently defeated, the Protectorate turned now to
salvaging, rather than slaughtering the human survivors of the
post-human war. Quarantine centers were organized on marginal,
devastated worlds long abandoned by the Forsaken during the events of
the Aeran war. There were so few survivors, that the operation, even
given the ravaged nature of the Protectorate fleet and economy, was
feasible. Actual Shaper worlds, rather than briefly conquered worlds,
were not landed on for centuries, instead being dusted with
radioactive materials, bombarded with asteroids and anti-matter bombs,
seeded with counter-terraforming machines, and kept under constant
vigilance and quarantine until no signs of life were visible. They
would remain under quarantine even after this point until ground
crews, only now visiting the planets, could verify no residual "gifts"
lingering on. Where intransigent Shaper bio-forms were still found,
they were rooted out until the planets were deemed sanitized,
sterilized, and untainted, only then becoming eligible for
re-terraforming. Rlaan, specifically Rlaan-Briin, were brought in to
help clean up the bio-weapon mess that was the former human inhabited
region of space. Compensation for this assistance, along with the new
political and military realities, was to make the Rlaan-Briin, as an
anthrophilic force within the Rlaan, the dominant group in Rlaan
politics.

On a sad aside, the Dgn died along with their masters. Indeed, the
only Dgn to survive the war would be those Dgn babies inflicted upon
the Shmrn by the Shaper terror attack.

While many former human worlds were deemed too dangerous for human
habitation, they were a fine environment for a new generation of
Grandchildren to settle, for the first time moving out of their
Andolian cradle. Backed by these new colonies of Grandchildren, the
Protectorate would eventually become a different entity - the First
Empire, expanding to include Shmrn, Uln, and eventually Rlaan space
under a single sovereignty. This expansion would be one of implied,
rather than exercised force, and would be performed through a series
of peacefully negotiated annexations.

Under the Empire, there would be a distinct split between those
populations merging into the increasingly collective population of
Andolian origin and those relegating themselves to an increasingly
zoo-like status, preserved as the Ur-Human, Ur-Aeran, Ur-Rlaan,
etc. on worlds both garrisoned under and protected by
Grandchildren. The Grandchildren themselves would change as well, the
garrison forces becoming increasingly detached from the path of the
Empire, while the Grandchildren who became settlers of the former
human worlds would become vital participants in all aspects of the
Empire. As millennia passed, these divisions would come to have names,
the Legion, the Ur-, the Firstborn, and the Inheritors,
respectively. There also arose another class, beings constructed to
bridge generational gaps within the Legion, who, due to their
indefinite lifespans, were known as Methusalans. Though a very small
portion of the total population, their role in monitoring and
mitigating the negative effects of cultural drift would give them
important positions within the Empire.

The Pax Legion would last for many generations before interruption by
ex-Saahasayaay forces that had regrouped outside the jump network for
some millennia. While the war was brutal, it was decisively won with
knowledge gleamed from study of Ancient library information. However,
evidence mounted that the aggression between the two sides both using
vaguely Ancient derived technology had served to further rouse TWHON
enitities. The majority of the Legions would end up heading off in a
mass exodus to dwell around the deep-space devices responsible for
maintaining the jump-network, as these would provide cover for further
investigations into Ancient derived technology and limit the danger to
the Ur. The Inheritors would remain behind to pursue their own ends
absent the Legions' supervision, with maintenance of contact and
mitigation of drift falling to the Methusalas as the go-betweens.

This arrangement would last only for a few centuries before, in a
series of tragic confusions, the return of the descendants of the
10,000 would once more break the peace.

\section{Foundering Inheritance (Fleet of 10,000)}
\section{Rimward of Eden}
\section{Post Rimward of Eden}
To Be Discussed later.

\section{A conversation (edited)}

rimward of where the THWON have returned to

in the madness that remains of them

back to where old memories took them

but the stories are all quite far apart in time

the one I want to attach to "Rimward of Eden" takes place several thousand years after "Upon the coldest sea"

indeed, several thousand years after "Foundering Inheritance"

which takes place hundreds of years after Upon the Coldest Sea

there was a struggle between the Ancients and their creators, TWHON

think Gods and Titans, only with a different outcome

The Ancients lost their lives, the TWHON lost its/their (it's hard to choose the better term) sanity, among other things

there were, of course, many different types of Ancient

diverse playthings crafted while the mind of TWHON was itself more diverse

many of the varied and bizarre devastations visited upon former Ancient worlds occurred after the struggle proper, after the residents were themselves all or mostly dead

and what was left of the TWHON was far from sane, far, even, from a semblance of intelligible thought

torn apart by a civil war among different aspects of a single mind

shredded utterly bashing itself against the small but sharp holdouts of the Ancients

TWHON is an appropriate name for them, for the aspects were not so much named as {\em THE NAME}. Embodied aspects of thoughts, principles, ambition

But what is left is no longer a mind

D: don't suppose you've got history for what they were before and how they became that mind?

T: not really - that predates even their own knowledge, to some extent

it happened a long time ago

the Ancients certainly didn't know the details

part of the reason their libraries (some of which continue operation in the void between stars) are interested in watching groups such as the Andolians

pondering how a life form evolves from muck to Godhead

D: cool

T: but for an approximate date, let's say that the ancestors of the TWHON became the sole owners of the entire galaxy at least 1 billion years ago

what may or may not have co-existed prior to this is unknown

how many species they initially represented is unknown

what they became was very well known to the Ancients, and not really much at all known to the younger species

I've tried to keep details about the TWHON out of wide public distribution

D: which you've done very well ;-)

T: a developers-only section would be nice... neh, whatever

Some things are best not served until thoroughly cooked

D: hehehe, indeed

T: but, I'm happy you think it a worthwhile yarn that I'm spinning

D: ya, it sounds sweet :-)

T: how's the plot for Black Paralysis coming these days, btw?

D: the plot is just about as murky as it's every been ;-)

T: ok, well, I was jotting down some things in the timeline, and that may be of use/interest to you at some point

D: that may well be a very good idea :-)

T: so, basically, I see Black Paralysis as potentially being a really good setup piece for all the glory, suffering, and devastation that comes in Foundering Inheritance

with the title being an apt reference to our inability to move quickly outside of the jump network, which, by the time it takes place, is pretty well hemming everyone in (the Aera were just unlucky enough to be in a corner and find it out sooner than most others)

D: hey! that's like one of the many meanings I actually thought of :-)

T: namely, the discovery of the "on-ramps" out of our section of the galaxy

and out of our section of the jump network

which.... seem to be blocked off

this, of course, is of profound interest to the remnants of the resisting Aera

and sparks some turning points in Human-Rlaan relations

amid some power shifts internal to the Rlaan between anthro-philic and anthro-phobic elements

mostly opaque to the casual human observer

an opacity not to our benefit

D: okay, you've just summed up basically everything I had worked out so far (and then some) ... neato :-)

T: but ultimately not to theirs (but not until Foundering Inheritance)

Foundering Inheritance will feature the launching of the fleet of 10,000, the joint expeditionary force, a grand symbol of of unity, progress, and optimism 

it will also feature the near destruction of the entire Rlaan species

both of them ;-)

D: hehehe, howso?

T: there are multiple on-ramps

they have to be opened from both sides

the expeditionary fleet was sent into one

but the Rlaan had their own plans for one they didn't bother telling anyone about

unfortunately, while the expeditionary fleet went through something unlocked due to ages of neglect

the Rlaan opened something unlocked because the other side was waiting, and hoping for someone to unlock the far side

oops

D: heh, unlucky

T: insufficient precaution, in many ways

over-trust of their own

relative to the efforts at the joint gate

where cooperation bred sufficient mistrust

D: that doesn't stike me as an overly rlaan trait though (overtrust)

T: but of whom?

they trusted themselves

they were betrayed by themselves

unwittingly

hubris

D: trust in teh benign unknown then

they seem to have always approached new contact with caution

T: they didn't know they were making contact :)

D: ya, but surely they would have thought of the possibility?

T: oh, yes, they did

but that wasn't what screwed them

they didn't know they were making contact {\em while they were making contact}

D: oh

T: they didn't think far enough outside the box to understand what happened before it was rather late in the game - a somewhat more Rlaan trait

D: ya

T: but yes, bad luck as much as anything else

had it been the joint venture going there, it would have been disastrous, but less likely cataclysmic

they got greedy and paid for it dearly

as did most everyone else

pay for it dearly, that is

could have been even worse

except that it got worse ;-)

but in a way that helped those further away

D: very cryptic :-P

T: That which was overtaking the Rlaan attracted the notice of the remnants of TWHON

not that it knew what they were

not that the Rlaan really understood what it was when it came through

but the nano-plague remembered

it remembered very clearly

for the first time in a very long time

it fulfilled its intended purpose

D: ooooh, nice :-)

T: now, the conclusion of this part will intentionally leave a few loose ends, but make clear what groups are still around

it will therefore be somewhat surprising to view the situation at the beginning of Rimward of Eden

(if one has been following events)

because, in the long interim between the two, much seems to have been both gained and lost

the protagonist in Rimward of Eden is actually an Aeran youth

on a planet of subjugated Aera ruled over by humans, themselves clearly fallen from some greater height 

upon further inspection it will become apparent that interstellar travel no longer appears to be occurring

D: interesting

T: As the protagonist struggles for his own survival in a landscape of standing buildings and collapsed societies, powerful artifacts and abject poverty, he will end up seeking the aid of an ancient "demon"

(of course, nothing of the sort, but what do they know?)

whose aid will end up transporting them and others met along the way on a trip that opens their eyes (and ours) to what has become of our stretch of stars

filling in some of the pieces in-between, will be the story of the fleet of 10,000 which ends up stranded (relative to where it launched from) by the events concluding Foundering Inheritance

but this story takes place a long way away, and is a story of what happens in our neck of the woods only though inference, and the story of what happened somewhere else, a long time ago

D: :-)

if this were a set of novels, I would so own the series :-)

T: as a set of games, I don't envision all of them being well suited to being flight-sims, or at least not in any traditional sense

D: it seems to be most of them won't be

T: well, Icarus Descended would be best served Elite-like
moreso even than Upon the Coldest Sea, in that it {\em is} an often lawless, expanding frontier
much more "western"-esque

makes all the "space-is-for-libertarians" crowd happier... until they remember what it all turned into in VS - freedom begat monopolies, totalitarian states, fascist intrusions, and religious extremism as well as democracies, progress, and personal growth

There's also the story of the Fraternal War (aka. the war between the Andolians and Lightbearers) that I think has some good stories in it. From a Klk'k perspective, I'd go with "The Stars are also Pyres" as a good proto-title

that could play reasonably well as a mission-based flight sim

although, there are some ground elements that are more interesting

although it perhaps makes more sense to just experience them, rather than influence them

And then there's always the best fodder for an FPS - "Simon says die"

about the Simons

D: hehehe

T: did I tell you about the Simons?

D: nope :-)

T: think simon wiesenthal meets "the punisher"

From a Klk'k perspective

D: neat idea :-)

T: anywho, so those are the broadest outlines of where I see story points of interest in a ~9000 year spread of time in the VS future history

D: ya, it sounds really cool :-) you've certainly put a hell of a lot of thought into it :-)

T: well, I wish I'd put as many hours of writing into it ;-)

at least gaim archives conversations, so I can point and grunt when people ask again ;-)

D: lol

T: I think Rimward of Eden would be best as an adventure game of sorts

whatever the interface, that's the genre it seems most part of

even if the mechanics involve a first person perspective

I think, especially in a first person perspective, it might be interesting to be an Aera

D: definitely :-)

foundering inheritance doesn't really fit anywhere though in terms of comp games ... it'd be a great novel, but there's so much that just couldn't be conveyed in just one tyupe of game

T: well, maybe it doesn't make sense to just make it one game then :)

D: touche

T: but... given how long its taking us (admittedly at very-much-part-time-pace)

to do one installment, I'll worry about getting the universe in shape first

and then how we're going to present it

because there {\em are} several key points that do map to tangible genres

the other thing, is to look at one of the things I really want to do with VS:utcs

which is, to have a very flexible progression of "little story" in the scope of a much less flexible "big story"

which the character doesn't so much influence, as live in (the big story, that is)

Note that Deucalion is not the linchpin for 9000 years of VS history

D: course not :-)

utcs is just part of the story from his perspective

T: well, given that he hadn't been created yet, it'd be hard for him to be much influence on Icarus Descended ;-)

D: hehehe

T: and, unbeknownst to him, he did play some role, though through little fault of his own, in the timing of some events in utcs ;-)

which is to say, the role he played occurs befor utcs proper

and hence can't be screwed up by the player - then again, it'd be pretty hard to do, as the only required action is (more or less) his existing at all 

Deucalion is a remarkable being, in that he's actually an apology

Deucalion is a library's way of apologizing for killing an exploration crew some decades prior

His role in the timing of things entirely being an issue of getting the AI quorum pondering what built him, how, and why

and the effect that such ruminations had upon the rollout time for the "grandchildren"

sometimes, it is enough to know something is possible

that alone can make a difference, even if the particulars are beyond reach

a nudge, a narrowing of search, a path otherwise not taken, that leads to a slight speedup in something else, itself long in the making, long before Deucalion

D: cool :-)
